### 2021-06-11 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
Initiate R project repository

---
title: "msAggr_seurat-v1"
output: html_notebook
---

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/")
library(dplyr)
library(Seurat)
library(patchwork)
```
store session info
```{r}
sink("msAggr_seurat-v1.20210608")
sessionInfo()
sink()
```

# Define functions

```{r}
read_10XGenomics_data<-function(sample.list){
	data <- sapply(sample.list, FUN = function(x) {x = dirname(list.files(path = x, pattern = 'barcodes.tsv', full.names = TRUE, recursive = TRUE, no.. = TRUE))})
	names(data) <- sample.list
	return(data)
}
```
# A note about using SCTransform versus `ScaleData`
https://bioconductor.org/packages/3.10/workflows/vignettes/simpleSingleCell/inst/doc/batch.html#62_for_gene-based_analyses
>You can also normalize and scale data for the RNA assay. There are numerous resources on this, but Aaron Lun describes why the original log-normalized values should be used for DE and visualizations of expression quite well here:
>
>For gene-based procedures like differential expression (DE) analyses or gene network construction, it is desirable to use the original log-expression values or counts. The corrected values are only used to obtain cell-level results such as clusters or trajectories. Batch effects are handled explicitly using blocking terms or via a meta-analysis across batches. We do not use the corrected values directly in gene-based analyses, for various reasons:
>
>It is usually inappropriate to perform DE analyses on batch-corrected values, due to the failure to model the uncertainty of the correction. This usually results in loss of type I error control, i.e., more false positives than expected.
>
>The correction does not preserve the mean-variance relationship. Applications of common DE methods like edgeR or limma are unlikely to be valid.
>
>Batch correction may (correctly) remove biological differences between batches in the course of mapping all cells onto a common coordinate system. Returning to the uncorrected expression values provides an opportunity for detecting such differences if they are of interest. Conversely, if the batch correction made a mistake, the use of the uncorrected expression values provides an important sanity check.
>
>In addition, the normalized values in SCT and integrated assays don't necessary correspond to per-gene expression values anyway, rather containing residuals (in the case of the scale.data slot for each).

Mess with how to load 4 cell populations into single seurat object

SET SEED?????!!!!!

```{r}
data_files.list<-read_10XGenomics_data(data)
seurat.object.data<-Read10X(data_files.list)
seurat.object<- CreateSeuratObject(raw.data = seurat.object.data, min.cells = 3, min.genes = 200, project = projectName)
seurat.object<- create_percentMito_column(seurat.object)

```

```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident")
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident")
plot1 + plot2
```

remove low quality cells
require: nFeature_RNA between 200 and 2500 (inclusive)
--require: percent.mt <5--???
```{r}
print(paste("original object", seurat.object, sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 2500
												)
print(paste("new object", seurat.object, sep = " "))
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
```

Find variable features

```{r}
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
```

`selection.method	
How to choose top variable features. Choose one of :

vst: First, fits a line to the relationship of log(variance) and log(mean) using local polynomial regression (loess). Then standardizes the feature values using the observed mean and expected variance (given by the fitted line). Feature variance is then calculated on the standardized values after clipping to a maximum (see clip.max parameter).

mean.var.plot (mvp): First, uses a function to calculate average expression (mean.function) and dispersion (dispersion.function) for each feature. Next, divides features into num.bin (deafult 20) bins based on their average expression, and calculates z-scores for dispersion within each bin. The purpose of this is to identify variable features while controlling for the strong relationship between variability and average expression.

dispersion (disp): selects the genes with the highest dispersion values`

### 2021-06-11 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
continue seuratv4 pipeline development

@@
 ```{r setup}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/")
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
 library(dplyr)
@@
 
-# Define functions
-
-```{r}
-read_10XGenomics_data<-function(sample.list){
-	data <- sapply(sample.list, FUN = function(x) {x = dirname(list.files(path = x, pattern = 'barcodes.tsv', full.names = TRUE, recursive = TRUE, no.. = TRUE))})
-	names(data) <- sample.list
-	return(data)
-}
-```
 # A note about using SCTransform versus `ScaleData`
@@
 
+## Set global variables
+
 ```{r}
-data_files.list<-read_10XGenomics_data(data)
-seurat.object.data<-Read10X(data_files.list)
-seurat.object<- CreateSeuratObject(raw.data = seurat.object.data, min.cells = 3, min.genes = 200, project = projectName)
-seurat.object<- create_percentMito_column(seurat.object)
+projectName <- "msAggr"
+```
+
+```{r}
+source("msAggr_AnalysisCode/read_10XGenomics_data.R")
 
@@
 ```{r}
-seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
-VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
+setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
+data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
+seurat.object.data<-Read10X(data_file.list)
 ```
@@
 ```{r}
-plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident")
-plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident")
+seurat.object<- CreateSeuratObject(counts = seurat.object.data, min.cells = 3, min.genes = 200, project = projectName)
+```
+
+```{r}
+seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
+VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident')
+```
+
+```{r fig.width=4, fig.height=2}
+plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
+plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
 plot1 + plot2
@@
 remove low quality cells
-require: nFeature_RNA between 200 and 2500 (inclusive)
+require: nFeature_RNA between 200 and 4000 (inclusive)
 --require: percent.mt <5--???
+
 ```{r}
-print(paste("original object", seurat.object, sep = " "))
+print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
 seurat.object <- subset(seurat.object, 
 												subset = nFeature_RNA >=200 & 
-													nFeature_RNA <= 2500
+													nFeature_RNA <= 4000
 												)
-print(paste("new object", seurat.object, sep = " "))
+print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
 ```
 
-```{r}
-seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
-```
+## Normalization
 
-Find variable features
-
-```{r}
-seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
-```
+Struggling to wrap my head around this one. It seems that SCTransform is best for batch correction, but `NormalizeData` and `ScaleData` are best for DGE. Several vignettes have performed both
 
@@
 dispersion (disp): selects the genes with the highest dispersion values`
+
+```{r}
+seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
+```
+
+```{r}
+seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
+```
+
+Find variable features
+```{r fig.width = 5, fig.height = 2}
+seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
+top10 <- head(VariableFeatures(seurat.object), 10)
+plot1 <- VariableFeaturePlot(seurat.object)
+plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
+plot1 + plot2
+
+```
+
+Scale data (linear transformation)
+
+```{r}
+all.genes <- rownames(seurat.object)
+seurat.object <- ScaleData(seurat.object, features = all.genes)
+```
+
+## PCA
+
+linear dimensional reduction. Default are based on VariableFeatures, but can be changed
+
+```{r}
+seurat.object <- 
+```

### 2021-06-14 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
Seuratv4 pipeline continuance

@@
 library(patchwork)
+library(ggplot2)
 ```
@@
 
+Clean up to free memory
+
+```{r}
+remove(seurat.object.data)
+```
+
+Add mitochondrial metadata and plot some basic features
 ```{r}
@@
 
+### Save progress
+
+```{r}
+save.image(file = paste0(projectName, '.RData'))
+```
+
 ## PCA
@@
 ```{r}
-seurat.object <- 
+seurat.object <- RunPCA(seurat.object, features = VariableFeatures(object = seurat.object))
 ```
+
+```{r fig.width=4, fig.height=4}
+VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 2)
+```
+
+DimPlot colored by orig.ident
+```{r}
+DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
+```
+
+Let's put in a concerted effort to pick the right dimensionality using the newest software
+
+```{r}
+seurat.object <- JackStraw(seurat.object, num.replicate = 100, dims = 40) #runs ~50 min
+seurat.object <- ScoreJackStraw(seurat.object, dims = 1:40)
+save.image(paste0(projectName, ".RData"))
+```
+
+Draw dim.reduction plots
+
+```{r}
+# plot1 <- JackStrawPlot(seurat.object, dims = 1:50)
+```
+```{r, figures-side, fig.show='hold', out.width="50%"}
+ElbowPlot(seurat.object, ndims = 50)
+percent.variance(seurat.object@reductions$pca@stdev)
+
+```
+Number of PCs describing X% of variance
+
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
+paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
+paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
+paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
+
+```
+
+36 PCs describe 95% percent of variance, but one of the earlier critiques was that too many dimensions were included.
+Best way to deal with this is probably to try a couple of different ones.
+Let's start with 36 because I dare to rebel.
+
+```{r}
+seurat.object <- FindNeighbors(seurat.object, dims = 1:36)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+```
+
+Generate UMAP data
+```{r}
+seurat.object <- RunUMAP(seurat.object, dims = 1:36) # took < 1 min
+```
+
+Plot the results
+```{r}
+DimPlot(seurat.object, 
+				reduction = "umap"
+				) + ggtitle("msAggr dim36 res0.5")
+
+```
+
+```{r}
+DimPlot(seurat.object,
+				reduction = "umap", 
+				group.by = "orig.ident"
+				) + ggtitle("msAggt dim36 orig.ident")
+```
+
+```{r}
+saveRDS(seurat.object, file = "msAggr_dim36.rds")
+```
+
+Not a fan of how "swoopy" it looks. Could be to many dimensions? May as well try 90%
+
+```{r}
+seurat.object <- FindNeighbors(seurat.object, dims = 1:24)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1:24)
+```
+
+Plot the results
+```{r}
+DimPlot(seurat.object, 
+				reduction = "umap"
+				) + ggtitle("msAggr dim24 res0.5")
+
+```
+
+```{r}
+DimPlot(seurat.object,
+				reduction = "umap", 
+				group.by = "orig.ident"
+				) + ggtitle("msAggt dim24 orig.ident")
+```
+
+```{r}
+saveRDS(seurat.object, file = "msAggr_dim24.rds")
+```
+
+24 vs 36 dimensions doesn't seem to make that much difference RE overall relationships. Should probbaly create two separate objects and try a few plotting methods.

### 2021-06-16 msAggr_AnalysisCode/msAggr_sctransform-v1.Rmd
start work on SCTransform method

---
title: "msAggr_sctransform-v1"
output: html_notebook
---

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```
store session info
```{r}
sink("msAggr_sct-v1.20210615")
sessionInfo()
sink()
```

# A note about using SCTransform versus `ScaleData`
https://bioconductor.org/packages/3.10/workflows/vignettes/simpleSingleCell/inst/doc/batch.html#62_for_gene-based_analyses
>You can also normalize and scale data for the RNA assay. There are numerous resources on this, but Aaron Lun describes why the original log-normalized values should be used for DE and visualizations of expression quite well here:
>
>For gene-based procedures like differential expression (DE) analyses or gene network construction, it is desirable to use the original log-expression values or counts. The corrected values are only used to obtain cell-level results such as clusters or trajectories. Batch effects are handled explicitly using blocking terms or via a meta-analysis across batches. We do not use the corrected values directly in gene-based analyses, for various reasons:
>
>It is usually inappropriate to perform DE analyses on batch-corrected values, due to the failure to model the uncertainty of the correction. This usually results in loss of type I error control, i.e., more false positives than expected.
>
>The correction does not preserve the mean-variance relationship. Applications of common DE methods like edgeR or limma are unlikely to be valid.
>
>Batch correction may (correctly) remove biological differences between batches in the course of mapping all cells onto a common coordinate system. Returning to the uncorrected expression values provides an opportunity for detecting such differences if they are of interest. Conversely, if the batch correction made a mistake, the use of the uncorrected expression values provides an important sanity check.
>
>In addition, the normalized values in SCT and integrated assays don't necessary correspond to per-gene expression values anyway, rather containing residuals (in the case of the scale.data slot for each).

Load 4 populations into single seurat object

SET SEED?????!!!!!

# Set global variables
```{r}
projectName <- "msAggr_sct"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")

```

```{r}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
object.data <-Read10X(data_file.list)
```

```{r}
seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
```

Clean up to free memory

```{r}
remove(seurat.object.data)
```

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

```{r fig.width=4, fig.height=2}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```
remove low quality cells **
require: nFeature_RNA between 200 and 4000 (inclusive)
--require: percent.mt <5--???

\** In previous run it's possible that we'll need to limit percent.mt because of MEPm weirdness. Will have to see if the additional 2000 MEPm cells found by cellranger v6

```{r}
print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000
												)
print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- PercentageFeatureSet(seurat.object, pattern = '^mt-', col.name = "percent.mt")
```

## SCTransform

As previously mentioned, `SCTransform()` is best for batch correction, whereas `NormalizeData()`, `ScaleData()`, and `FindVariableFeatures()` are best for DGE. I'm struggling to make stable clusters, so I'm jumping ahead to `SCTransform()`.  
It does appear that `SCTransform()` data are stored in the SCT assay slot, which does not overwrite other normalization slots. It will become the default slot once populated.

```{r}
seurat.object <- SCTransform(seurat.object, 
														 method = "glmGamPoi", 
														 vars.to.regress = c("nCount_RNA", "nFeature_RNA")
														 ) # method = "glmGamPoi" speeds process
```

Tada! Now lets do some plotting and clustering!

linear dimensional reduction. Default are based on VariableFeatures, but can be changed

```{r}
seurat.object <- RunPCA(seurat.object, features = VariableFeatures(object = seurat.object), verbose = FALSE)
```

```{r fig.width=4, fig.height=4}
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 2)
```

DimPlot colored by orig.ident
```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r}
source("msAggr_AnalysisCode/PercentVariance.R")
```

```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```
# Variance described by dimensionality
Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

# Dim33

Cover 95% variability

```{r}
ndim <- 33
```

```{r}
seurat.object <- RunUMAP(seurat.object, dims = 1:ndim)
seurat.object <- FindNeighbors(seurat.object, dims = 1:ndim)
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```
Plotting...

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("SCT_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
											) + 
			ggtitle(paste0("msAggr sct dim31 res", gsub("SCT_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

```{r}
saveRDS(seurat.object, file = "msAggr_sctdim33.rds")
```

## Evaluate cluster stability
Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.

Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.

### Clustree
Previously my favourite has been Clustree, which gives a nice visual
NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.

```{r fig.height=5}
clustree(seurat.object, prefix = "SCT_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

Basically no stability difference compared to before.

# Dim22

Cover 90% variability

```{r}
ndim <- 22
```

```{r}
seurat.object <- RunUMAP(seurat.object, dims = 1:ndim)
seurat.object <- FindNeighbors(seurat.object, dims = 1:ndim)
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```
Plotting...

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("SCT_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
											) + 
			ggtitle(paste0("msAggr sct dim", ndim, "res", gsub("SCT_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

```{r}
saveRDS(seurat.object, file = paste0("msAggr_sctdim", ndim, ".rds"))
```

## Evaluate cluster stability
Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.

Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.

### Clustree
Previously my favourite has been Clustree, which gives a nice visual
NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.

```{r fig.height=5}
clustree(seurat.object, prefix = "SCT_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

Hard to tell if dim33 or dim22 is better, but if *feels* like dim22 has less cross-over between res0.5 and res1.0. Will continue at dim22, and see how things change between res0.5 and res1.0

```{r}
# Number of filtered cells left in each pop
sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
```
## MitoC in

```{r}
for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
	h$density = h$counts/sum(h$counts)*100
	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
}

### 2021-06-16 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
start work on SCTransform method

@@
 library(ggplot2)
+library(clustree)
 ```
@@
 projectName <- "msAggr"
+jackstraw.dim <- 40
 ```
@@
 seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
-VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident')
+VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
 ```
@@
 ```{r}
-seurat.object <- JackStraw(seurat.object, num.replicate = 100, dims = 40) #runs ~50 min
-seurat.object <- ScoreJackStraw(seurat.object, dims = 1:40)
+jackstraw.dim <- 40
+seurat.object <- JackStraw(seurat.object, num.replicate = 100, dims = jackstraw.dim) #runs ~50 min
+seurat.object <- ScoreJackStraw(seurat.object, dims = 1:jackstraw.dim)
 save.image(paste0(projectName, ".RData"))
@@
 ```{r}
-# plot1 <- JackStrawPlot(seurat.object, dims = 1:50)
+JackStrawPlot(seurat.object, dims = 25:36)
 ```
@@
 percent.variance(seurat.object@reductions$pca@stdev)
-
 ```
@@
 ```
+
+# KNN dimensional reduction
+## Dim36
 
@@
 
-Not a fan of how "swoopy" it looks. Could be to many dimensions? May as well try 90%
+## Dim24
+36 dimensions looks very "swoopy". might have asked for too many dimensions. Will try dim = 24, which accounts for 90% of variance
+
+```{r}
+# seurat.object <- FindNeighbors(seurat.object, dims = 1:24)
+# seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+# seurat.object <- RunUMAP(seurat.object, dims = 1:24)
+```
+
+Plot the results
+```{r}
+# DimPlot(seurat.object, 
+# 				reduction = "umap"
+# 				) + ggtitle("msAggr dim24 res0.5")
+
+```
+
+```{r}
+# DimPlot(seurat.object,
+# 				reduction = "umap", 
+# 				group.by = "orig.ident"
+# 				) + ggtitle("msAggt dim24 orig.ident")
+```
+
+```{r}
+# saveRDS(seurat.object, file = "msAggr_dim24.rds")
+```
+
+## KNN conclusion
+24 vs 36 dimensions doesn't seem to make that much difference RE overall relationships between cells.
+Should probbaly create two separate objects and try a few plotting methods, but will start by focusing on 36 dimensions.
+
+# Cluster analysis
+
+Will proceed with dim = 36 and do clustering analysis with a range of clusters. Later can do the tree-based over-clustering assessment
+
+```{r}
+seurat.object <- readRDS("msAggr_dim36.rds")
+```
+
+## Generate cluster profiles
+Picking range of resolutions
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+seurat.object <- RunUMAP(seurat.object, dims = 1:36)
+```
+
+Plot the clustering over different resolutions. Going to need a much better color palette, or try mixing in some different symbols
+
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
+											) + 
+			ggtitle(paste0("msAggr dim36 res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+```{r}
+# saveRDS(seurat.object, file = "msAggr_dim36.rds")
+```
+
+## Evaluate cluster stability
+Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
+
+Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
+
+### Clustree
+Previously my favourite has been Clustree, which gives a nice visual
+NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+These data suggest that node stability is aweful! Need to figure out if this is a dimensional reduction error or a clustering error.
+
+# Why are clusters so unstable?
+Differences could include:
+* cells in each population (cellranger v6 includes more cells than cellranger v1, especially in MEP)
+* dimensionality is incorrect
+* ScaleData didnt account for regression factors (e.g., "nCounts_RNA" or "nFeatures_RNA")
+* Incorrect normalization/scaling method
+* Clustering is too strict or not strict enough
+* neighborhood analysis used wrong parameters
+* Should include mitoC filter (there's a chunk of MEP w/ mitoC @ ~40%)
+* SCTransform accounts better for sources of variability
+
+```{r}
+# Number of filtered cells left in each pop
+sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
+```
+
+```{r}
+for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
+	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
+	h$density = h$counts/sum(h$counts)*100
+	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
+}
+```
+
+Looks like MEPm is the only sample with that huge MitoC % lump @ 40%. What do these cells look like, otherwise?
+
+```{r fig.width=5}
+VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
+				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
+```
+Save dim36 as is and try clustering analysis @ dim24
+```{r}
+saveRDS(seurat.object, file = "msAggr_AnalysisCode/msAggr_dim36.rds")
+```
+
+# Repeat clustering with dim24
+One possibility is that I've included too many dimensions. Will see if 90% increases stability.
 
@@
 
-Plot the results
-```{r}
-DimPlot(seurat.object, 
-				reduction = "umap"
-				) + ggtitle("msAggr dim24 res0.5")
-
-```
-
-```{r}
-DimPlot(seurat.object,
-				reduction = "umap", 
-				group.by = "orig.ident"
-				) + ggtitle("msAggt dim24 orig.ident")
-```
-
+Save object
 ```{r}
@@
 
-24 vs 36 dimensions doesn't seem to make that much difference RE overall relationships. Should probbaly create two separate objects and try a few plotting methods.
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
+											) + 
+			ggtitle(paste0("msAggr dim36 res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+## Evaluate cluster stability
+Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
+
+Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
+
+### Clustree
+Previously my favourite has been Clustree, which gives a nice visual
+NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+Think I'll explore regression factors using SCTransform in new document.

### 2021-06-22 CMPSubset.Rmd
Added CMP subset

---
title: "CMPSubset"
output: html_notebook
---

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
# library(clustree)
```
store session info
```{r}
sink("CMP-v1.20210616.txt")
sessionInfo()
sink()
```

# Set global variables
```{r}
projectName <- "CMP"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")
source("msAggr_AnalysisCode/PercentVariance.R")
source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
```

```{r}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
```

```{r}
seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
```

Clean up to free memory

```{r}
remove(object.data)
```

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

```{r}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```

We don't have to worry about comparing library depths, so we'll just do normalization/Scale data

remove low quality cells
require: nFeature_RNA between 200 and 4000 (inclusive)
require: percent.mt <=5

```{r}
print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000 & 
													percent.mt <= 5
												)
print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
```

Find variable features
```{r fig.width = 5, fig.height = 2}
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(seurat.object), 10)
plot1 <- VariableFeaturePlot(seurat.object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

```

Scale data (linear transformation)

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
```

```{r}
seurat.object <- RunPCA(seurat.object, features = VariableFeatures(seurat.object), ndims.print = 1:5, nfeatures.print = 5)
```

```{r}
DimPlot(seurat.object, reduction = "pca")
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```
Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

# Cluster cells

Let's take 90% of variance
```{r}
ndims <- 33
```

```{r}
seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
seurat.object <- FindClusters(seurat.object, resolution = 0.5)
```
## Generate UMAP
```{r}
seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
```
Plot UMAP

```{r}
DimPlot(seurat.object)
```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

```{r}
clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

Add new metadata column
```{r}
seurat.object@meta.data['clust.ID'] <- 0
head(seurat.object@meta.data)
```

now map new ids
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
```

do numbers make sense?

```{r}
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 10,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 11,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 17,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
```

make enough sense!

## Find old cells on UMAP

time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4

```{r}
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "clust.ID", 
				cols = c("gray", "orange", "blue", "red", "green"))
```

Is this success?? It may be!! At the very least they're clustering togehter...  
print out the picture
```{r}
png(filename = "ClusterIDs_on_newCMP.png", height = 800, width = 800)
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "clust.ID",
				cols = c("gray", "orange", "blue", "red", "green"), 
				pt.size = 1)
dev.off()
```

Let's see if we can get some gene expression profiles on these...
```{r, fig.height=10, fig.width=18}
gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
```

Write out picture w/ & w/o dots
```{r}
png(filename = "Fav_genes_on_newCMP.png", height = 800, width = 1600)
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0, cols = c("gray", "orange", "blue", "red", "green"))
dev.off()

png(filename = "Fav_genes_on_newCMP-dotted.png", height = 800, width = 1600)
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
dev.off()

```

# DGE

Let's do some cluster analyses and see if we can find these populations in our new analysis.
## Ideal resolution...? (Is this a thing?)

```{r}
color.palette <- c(
	"coral",
	"chartreuse4",
	"goldenrod1",
	"cadetblue1",
	"burlywood",
	"brown",
	"brown1",
	"blue",
	"blue4",
	"azure3",
	"aquamarine",
	"antiquewhite",
	"cadetblue",
	"gold3",
	"black",
	"darkgreen",
	"deeppink",
	"darkviolet",
	"darkturquoise",
	"darkslategray",
	"darksalmon",
	"darkorchid1",
	"darkolivegreen2",
	"forestgreen",
	"dodgerblue",
	"green",
	"lightpink",
	"lightcoral",
	"khaki1",
	"maroon",
	"peru",
	"lightseagreen",
	"lightsalmon",
	"plum",
	"moccasin",
	"tan",
	"tan1"
)
```

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = color.palette
											) + 
			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?

```{r}
temp <- seurat.object@meta.data
df <- data.frame(matrix(ncol = 4, nrow = length(levels(temp$RNA_snn_res.0.5))))
colnames(df) <- c("c3", "c17", "c10", "c11")
rownames(df) <- sort(as.numeric(levels(temp$RNA_snn_res.0.5)))
tot.clus <- nrow(temp[temp$RNA_snn_res.0.5 == "0",])
num.3 <- nrow(temp[(temp$RNA_snn_res.0.5 == "0") & (temp$clust.ID == "3"),])
pct.3 <- as.integer(num.3/tot.clus * 100)
df["0", "c3"]<- pct.3
```

Test: what percentage of each new clusterID matches one of the older clusters?
```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
		colnames(enrich.df) <- c(3, 17, 10, 11)
		rownames(enrich.df) <- new.clusters
		meta.df <- seurat.object@meta.data
		for(row.id in rownames(enrich.df)){
			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
			for(col.id in colnames(enrich.df)){
				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
				pct.x <- as.integer(num.x / tot.clus *100)
				# print(pct.x)
				enrich.df[row.id, col.id] <- pct.x
			}
		}
		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
		xlsx::write.xlsx(enrich.df, file = "PctOfNewClustersOverlappingOldClusters.xlsx", sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
		print(enrich.df)
	}
}

```

Used the exce doc to do some fancy conditional formatting. Old cluster 17 is pretty dispersed until you it resolution 2.5. Otherise, cells in old cluster 17 do not constitute more than 40% of any cells in the new clusters.  
As far as I can see, the two approaches are to do DGEof new CMP w/ resolution = 2.5, AND/OR do DGe using older cluster IDs. Sure seems to make sense to do both...

# DGE w/ resolution = 2.5
Strt with comparing all clusters against all other clusters
Write out cluster info

calculate `FindAllMarkers()` for different idents and save to new file
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	Idents(seurat.object) <- tested.ident
	all.markers <- FindAllMarkers(seurat.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_res2.5.xlsx"), 
									 sheetName = tested.ident, 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
}
```

Create `FindAllMarkers()` lists for GSEA
```{r}
Idents(seurat.object) <- "RNA_snn_res.2.5"
res.2.5.allmarkers <- FindAllMarkers(seurat.object)
```

Map HGNC symbols
```{r}
Mouse2HumanTable <- Mouse2Human(res.2.5.allmarkers$gene)

HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(res.2.5.allmarkers$gene, Mouse2HumanTable$MGI)])
head(res.2.5.allmarkers)
res.2.5.allmarkers$HGNC <- HGNC
tail(res.2.5.allmarkers)
sig.res.2.5 <- res.2.5.allmarkers[res.2.5.allmarkers$p_val_adj <= 0.05, ]
sig.res.2.5 <- sig.res.2.5[c("avg_log2FC", "HGNC", "cluster")]
sig.res.2.5 <- sig.res.2.5[!(sig.res.2.5$HGNC == "" | is.na(sig.res.2.5$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
sig.res.2.5 <- sig.res.2.5[]

```

```{r}
for(cluster in unique(sig.res.2.5$cluster)){
	print(paste("writing cluster", cluster))
	new.table <- sig.res.2.5[sig.res.2.5$cluster == cluster, c("HGNC", "avg_log2FC")]
	new.table <- new.table[order(-new.table$avg_log2FC), ]
	write.table(new.table, file = paste0("RankList_res2.5_findAll_hgnc/res.2.5cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )

}
```

calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_", gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
}
```

```{r}
for (cluster in sort(as.numeric(levels(seurat.object@meta.data$RNA_snn_res.2.5)))){
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_res2.5.xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
```

## Combine clusters that might represent old cluster ids

# DGE w/ metadata against clust.ID against "0"
reset ident as "clust.ID" and rerun `FindAllMarkers()`
```{r}
	Idents(seurat.object) <- "clust.ID"
	all.markers <- FindAllMarkers(seurat.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_clustID.xlsx"), 
									 sheetName = "clustID", 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
```

```{r}
# Idents(seurat.object) <- "clust.ID"
for (cluster in unique(seurat.object@meta.data$clust.ID)){
	print(cluster)
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_clustID.xlsx"), 
									 sheetName = paste0("oldclust", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}

```

# Cell cycle analysis
Just so we know what we're working with

```{r}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes
```

### 2021-06-23 CMPSubset.Rmd
Updated CMP; reorganize folders

@@
 
+# Distinguishing features of clusters
+Previously defined biomark genes based on PC contributions. Original list was based on *all* msAggr, but let's see how CMP subset does?
+```{r fig.height = 30, fig.width=6}
+VizDimLoadings(seurat.object, dims = 1:10, nfeatures = 30, reduction = "pca", ncol = 2)
+```
+
+```{r}
+pca.df <- seurat.object[["pca"]]
+pca.df <- as.data.frame(as.matrix(slot(object = pca.df, name = "feature.loadings")))
+print(seurat.object[["pca"]], dims = 2, nfeatures = 5)
+rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2, decreasing = TRUE)[1:5], ])
+rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2)[1:5], ])
+```
+
+now we can get a list of principal components!  
+first pull the list of oldAnalysis CMP top PC genes
+```{r}
+cmp.biomark <- read.table(file = "/Users/heustonef/Desktop/CMPSubpops/BioMark/ProbePanels/CMP_PCTopGenes.txt", sep = "\t", header = TRUE)
+biomark.cmptargets <- c()
+for(df.col in 1:ncol(cmp.biomark)){
+	biomark.cmptargets <- c(biomark.cmptargets, biomark[,df.col])
+}
+print(colnames(biomark))
+print(paste("total gene count:", length(biomark.cmptargets)))
+```
+
+Now get the list of current pc gene trgets (oldAnalysis used ndim = 1:6, so we'll start with that range)
+```{r}
+pc.list <- c("PC_1", "PC_2", "PC_3", "PC_4", "PC_5", "PC_6")
+pc.genes <- lapply(pc.list, function(x) rownames(pca.df[pca.df[[x]] %in% sort(pca.df[[x]], decreasing = TRUE)[1:30],])) #targeting roughly 180 genes like in biomark.cmptargets
+pc.genes <- unique(unlist(pc.genes))
+print(paste("total gene count:", length(pc.genes)))
+```
+
+Now compare the lists, I guess:
+
+```{r}
+# setdiff(x,y) gives you things in x not in y. setdiff(y,x) gives you things in y not in x
+setdiff(biomark.cmptargets, pc.genes)
+# print(paste("\n length:", length(setdiff(biomark.cmptargets, pc.genes))))
+writeLines(c("", "length:", length(setdiff(biomark.cmptargets, pc.genes))))
+```
+Umm, yeah that went kinda how I expected. Let's do this again, but for the actual biomark gene lists.
+```{r}
+biomark <- read.table(file = "/Users/heustonef/Desktop/CMPSubpops/BioMark/ProbePanels/BiomarkProbeList.txt", sep = "\t")
+biomark <- biomark[,1]
+setdiff(biomark, pc.genes)
+writeLines(c("", "length:", length(setdiff(biomark, pc.genes))))
+```
+
+What if we increase the number of pcs but decrease the depth of each? This might cover more of `biomark	`, which was originally developed using msAggr instead of only the CMP subset
+```{r}
+pc.list <- c("PC_1", "PC_2", "PC_3", "PC_4", "PC_5", "PC_6", "PC_7", "PC_8", "PC_9", "PC_10")
+pc.genes <- lapply(pc.list, function(x) rownames(pca.df[pca.df[[x]] %in% sort(pca.df[[x]], decreasing = TRUE)[1:20],]))
+pc.genes <- unique(unlist(pc.genes))
+print(paste("total gene count:", length(pc.genes)))
+
+```
+```{r}
+setdiff(biomark, pc.genes)
+writeLines(c("", "length:", length(setdiff(biomark, pc.genes))))
+```
+
+For comparison, let's just see how many of `biomark.cmptargets` were actually included in `biomark`
+```{r}
+setdiff(biomark.cmptargets, biomark)
+writeLines(c("", "length:", length(setdiff(biomark.cmptargets, pc.genes))))
+```
+```{r}
+length(biomark) - length(setdiff(biomark, biomark.cmptargets))
+```
+```{r}
+length(biomark) - length(setdiff(biomark, pc.genes))
+```
+So when you look at it like that, it's not actually that far off.
+
+What are the similarities?:
+```{r}
+setdiff(setdiff(biomark, biomark.cmptargets), setdiff(biomark, pc.genes))
+```
+These are genes from the 97probes not in the old CMP set that are also not in the new CMP set.
+Other than Itga2b (which is a failed probe anyway), nothing screams.
+Also we'd have thrown Flt3 and Cd34 for in anyway because they're requisite cell surface markers (also Flt3 surface marker is expensive but otherwise not noteworthy and not used in the current sorting strategy)
+
+What about cell surface marker expression?
+* Cd34
+* Cd16/32
+* Cd9
+* Cd41
+* Cd48
+* Sca1 (just throw that in for sh*&s and giggles)
+```{r, fig.height = 15, fig.width=10}
+surface.markers <- c("Cd34", "Fcgr3", "Fcgr2b", "Cd9", "Itga2b", "Cd48", "Ly6a")
+FeaturePlot(seurat.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
+```
+Save as png
+```{r}
+png(filename = "FeaturePlot_CMP_surfaceMarkers_clustIDfacet.png", height = 1600, width = 1600)
+FeaturePlot(seurat.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
+dev.off()
+```
+
 # Cell cycle analysis

### 2021-06-25 CMPSubset.Rmd
continued analyses on CMP and msAggr

@@
 store session info
-```{r}
-sink("CMP-v1.20210616.txt")
+```{r }
+# sink("CMP-v1.20210616.txt")
 sessionInfo()
-sink()
+# sink()
 ```
@@
 
-```{r}
+```{r echo = FALSE}
 all.genes <- rownames(seurat.object)
 seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
+```
+
+## Save raw object
+```{r}
+saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
 ```
@@
 ```{r}
-DimPlot(seurat.object, reduction = "pca")
+DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
 VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
@@
 
-# Cluster cells
-
-Let's take 90% of variance
-```{r}
-ndims <- 33
-```
-
-```{r}
-seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
-seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-```
-## Generate UMAP
-```{r}
-seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
-```
-Plot UMAP
-
-```{r}
-DimPlot(seurat.object)
-```
-
-## Add cluster IDs from Seurat v1
+# Add cluster IDs from Seurat v1
 
@@
 do numbers make sense?
-
 ```{r}
@@
 nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
+```
+
+### Color palette
+```{r}
+color.palette <- c(
+	"coral",
+	"chartreuse4",
+	"goldenrod1",
+	"cadetblue1",
+	"burlywood",
+	"brown",
+	"brown1",
+	"blue",
+	"blue4",
+	"azure3",
+	"aquamarine",
+	"antiquewhite",
+	"cadetblue",
+	"gold3",
+	"black",
+	"darkgreen",
+	"deeppink",
+	"darkviolet",
+	"darkturquoise",
+	"darkslategray",
+	"darksalmon",
+	"darkorchid1",
+	"darkolivegreen2",
+	"forestgreen",
+	"dodgerblue",
+	"green",
+	"lightpink",
+	"lightcoral",
+	"khaki1",
+	"maroon",
+	"peru",
+	"lightseagreen",
+	"lightsalmon",
+	"plum",
+	"moccasin",
+	"tan",
+	"tan1", 
+	"red", 
+	"purple",
+	"khaki4",
+	"black", 
+	"plum4"
+)
+```
+
+# Cluster cells
+
+Let's take 90% of variance
+```{r}
+ndims <- 33
+```
+
+```{r}
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+```
+## Generate UMAP
+```{r}
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+```
+Plot UMAP
+
+```{r}
+DimPlot(seurat.object)
 ```

### 2021-06-25 CMPSubset.Rmd
further CMP analysis

@@
 
-# Cluster cells
+# Total var 90%
+## Neighborhood and umap
+set total.var <- 90%
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 90))
 
-Let's take 90% of variance
-```{r}
-ndims <- 33
-```
-
-```{r}
 seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
 seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-```
-## Generate UMAP
-```{r}
 seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
 Plot UMAP
-
-```{r}
-DimPlot(seurat.object)
-```
-
-make enough sense!
-
-## Find old cells on UMAP
-
-time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
-
-```{r}
-DimPlot(seurat.object,
-				reduction = "umap",
-				group.by = "clust.ID", 
-				cols = c("gray", "orange", "blue", "red", "green"))
-```
-
-Is this success?? It may be!! At the very least they're clustering togehter...  
-print out the picture
-```{r}
-png(filename = "ClusterIDs_on_newCMP.png", height = 800, width = 800)
-DimPlot(seurat.object,
-				reduction = "umap",
-				group.by = "clust.ID",
-				cols = c("gray", "orange", "blue", "red", "green"), 
-				pt.size = 1)
-dev.off()
-```
-
-Let's see if we can get some gene expression profiles on these...
-```{r, fig.height=10, fig.width=18}
-gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
-VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
-```
-
-Write out picture w/ & w/o dots
-```{r}
-png(filename = "Fav_genes_on_newCMP.png", height = 800, width = 1600)
-VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0, cols = c("gray", "orange", "blue", "red", "green"))
-dev.off()
-
-png(filename = "Fav_genes_on_newCMP-dotted.png", height = 800, width = 1600)
-VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
-dev.off()
-
-```
-
-# DGE
-
-Let's do some cluster analyses and see if we can find these populations in our new analysis.
-## Ideal resolution...? (Is this a thing?)
-
-```{r}
-color.palette <- c(
-	"coral",
-	"chartreuse4",
-	"goldenrod1",
-	"cadetblue1",
-	"burlywood",
-	"brown",
-	"brown1",
-	"blue",
-	"blue4",
-	"azure3",
-	"aquamarine",
-	"antiquewhite",
-	"cadetblue",
-	"gold3",
-	"black",
-	"darkgreen",
-	"deeppink",
-	"darkviolet",
-	"darkturquoise",
-	"darkslategray",
-	"darksalmon",
-	"darkorchid1",
-	"darkolivegreen2",
-	"forestgreen",
-	"dodgerblue",
-	"green",
-	"lightpink",
-	"lightcoral",
-	"khaki1",
-	"maroon",
-	"peru",
-	"lightseagreen",
-	"lightsalmon",
-	"plum",
-	"moccasin",
-	"tan",
-	"tan1"
-)
-```
 
@@
 
+Test: what percentage of each new clusterID matches one of the older clusters?
 ```{r}
-temp <- seurat.object@meta.data
-df <- data.frame(matrix(ncol = 4, nrow = length(levels(temp$RNA_snn_res.0.5))))
-colnames(df) <- c("c3", "c17", "c10", "c11")
-rownames(df) <- sort(as.numeric(levels(temp$RNA_snn_res.0.5)))
-tot.clus <- nrow(temp[temp$RNA_snn_res.0.5 == "0",])
-num.3 <- nrow(temp[(temp$RNA_snn_res.0.5 == "0") & (temp$clust.ID == "3"),])
-pct.3 <- as.integer(num.3/tot.clus * 100)
-df["0", "c3"]<- pct.3
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
+		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
+		colnames(enrich.df) <- c(3, 17, 10, 11)
+		rownames(enrich.df) <- new.clusters
+		meta.df <- seurat.object@meta.data
+		for(row.id in rownames(enrich.df)){
+			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
+			for(col.id in colnames(enrich.df)){
+				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
+				pct.x <- as.integer(num.x / tot.clus *100)
+				# print(pct.x)
+				enrich.df[row.id, col.id] <- pct.x
+			}
+		}
+		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
+		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
+		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		print(enrich.df)
+	}
+}
+
 ```
+Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
+
+## Find old cells on UMAP
+
+time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
+
+```{r fig.width = 4}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "clust.ID", 
+				# split.by = "orig.ident",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+```{r fig.width = 4}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "orig.ident", 
+				split.by = "clust.ID",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+# Total var 95%
+## Neighborhood and umap
+set total.var <- 95%
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 95))
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+
+For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
 
@@
 		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
-		xlsx::write.xlsx(enrich.df, file = "PctOfNewClustersOverlappingOldClusters.xlsx", sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
 		print(enrich.df)
@@
 
+```
+Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
+
+## Find old cells on UMAP
+
+time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
+
+```{r fig.width = 2}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "clust.ID", 
+				pt.size = .1,
+				# split.by = "orig.ident",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+```{r fig.width = 4}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "orig.ident", 
+				split.by = "clust.ID",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+### Gene expression of old clustrs on new map
+Let's see if we can get some gene expression profiles on these...
+```{r, fig.height=10, fig.width=18}
+gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
+VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
 ```

### 2021-07-07 msAggr_seurat.Rmd

--- 2021-06-11 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
@@
 ---
-title: "msAggr_seurat-v1"
+title: "msAggr_seurat"
 output: html_notebook
@@
 
-Creating new pipeline using seurat v4.0.2 available 2021.06.08
+Creating new pipeline using seurat v4.0.2 available 2021.06.23  
+Important notes:
+
+* FILTERING on `percent.mt`, but NOT regressing on `percent.mt`
+* Regressing on `nCounts_RNA` and `nFeature_RNA`
 
@@
 library(patchwork)
+library(ggplot2)
+# library(clustree)
 ```
@@
 ```{r}
-sink("msAggr_seurat-v1.20210608")
+sink("msAggr_seurat-v1.20210623")
 sessionInfo()
@@
 source("msAggr_AnalysisCode/read_10XGenomics_data.R")
-
+source("msAggr_AnalysisCode/PercentVariance.R")
 ```
@@
 data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
-seurat.object.data<-Read10X(data_file.list)
+data.object<-Read10X(data_file.list)
 ```
@@
 ```{r}
-seurat.object<- CreateSeuratObject(counts = seurat.object.data, min.cells = 3, min.genes = 200, project = projectName)
+seurat.object<- CreateSeuratObject(counts = data.object, min.cells = 3, min.genes = 200, project = projectName)
 ```
 
+Clean up to free memory
+
+```{r}
+remove(data.object)
+```
+
+Add mitochondrial metadata and plot some basic features
 ```{r}
 seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
-VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident')
+VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
 ```
@@
 require: nFeature_RNA between 200 and 4000 (inclusive)
---require: percent.mt <5--???
+require: percent.mt <= 5
 
@@
 												subset = nFeature_RNA >=200 & 
-													nFeature_RNA <= 4000
+													nFeature_RNA <= 4000 & 
+													percent.mt <= 5
 												)
@@
 
-```{r}
-seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
-```
-
 Find variable features
@@
 plot1 + plot2
-
 ```
@@
 all.genes <- rownames(seurat.object)
-seurat.object <- ScaleData(seurat.object, features = all.genes)
+seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("nCount_RNA", "nFeature_RNA"))
+```
+
+### Save progress
+
+```{r}
+save.image(file = paste0(projectName, '.RData'))
 ```
@@
 ```{r}
-seurat.object <- 
+seurat.object <- RunPCA(seurat.object, features = VariableFeatures(object = seurat.object))
 ```
+Plot results
+```{r fig.width=4, fig.height=4}
+VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 2)
+```
+
+DimPlot colored by orig.ident
+```{r}
+DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
+```
+Let's put in a concerted effort to pick the right dimensionality using the newest software
+```{r}
+# jackstraw.dim <- 40
+# seurat.object <- JackStraw(seurat.object, num.replicate = 100, dims = jackstraw.dim) #runs ~50 min
+# seurat.object <- ScoreJackStraw(seurat.object, dims = 1:jackstraw.dim)
+# save.image(paste0(projectName, ".RData"))
+```
+Draw dim.reduction plots
+```{r}
+# JackStrawPlot(seurat.object, dims = 25:36)
+```
+```{r, figures-side, fig.show='hold', out.width="50%"}
+ElbowPlot(seurat.object, ndims = 50)
+percent.variance(seurat.object@reductions$pca@stdev)
+```
+Number of PCs describing X% of variance
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
+paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
+paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
+paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
+
+```
+
+## Add cluster IDs from Seurat v1
+
+Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
+Create column "clust.ID" and populate with 0's. Then import IDs for clusters
+```{r}
+clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
+clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
+clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
+clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
+clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+```
+
+Add new metadata column and map new ids
+```{r}
+seurat.object@meta.data['clust.ID'] <- 0
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
+```
+
+do numbers make sense (we don't expect the count to b exactly the same as the numbers in the original cluster)?
+```{r}
+nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 10,])
+nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 11,])
+nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 17,])
+nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
+```
+
+make enough sense!
+
+# DGE
+
+Let's do some cluster analyses and see if we can find these populations in our new analysis.
+## Ideal resolution...? (Is this a thing?)
+### Color palette
+```{r}
+color.palette <- c(
+	"coral",
+	"chartreuse4",
+	"goldenrod1",
+	"cadetblue1",
+	"burlywood",
+	"brown",
+	"brown1",
+	"blue",
+	"blue4",
+	"azure3",
+	"aquamarine",
+	"antiquewhite",
+	"cadetblue",
+	"gold3",
+	"black",
+	"darkgreen",
+	"deeppink",
+	"darkviolet",
+	"darkturquoise",
+	"darkslategray",
+	"darksalmon",
+	"darkorchid1",
+	"darkolivegreen2",
+	"forestgreen",
+	"dodgerblue",
+	"green",
+	"lightpink",
+	"lightcoral",
+	"khaki1",
+	"maroon",
+	"peru",
+	"lightseagreen",
+	"lightsalmon",
+	"plum",
+	"moccasin",
+	"tan",
+	"tan1", 
+	"red", 
+	"purple",
+	"khaki4",
+	"black", 
+	"plum4"
+)
+```
+
+# Total var 90%
+## Neighborhood and umap
+set total.var <- 90%
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 90))
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
+
+Test: what percentage of each new clusterID matches one of the older clusters?
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
+		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
+		colnames(enrich.df) <- c(3, 17, 10, 11)
+		rownames(enrich.df) <- new.clusters
+		meta.df <- seurat.object@meta.data
+		for(row.id in rownames(enrich.df)){
+			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
+			for(col.id in colnames(enrich.df)){
+				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
+				pct.x <- as.integer(num.x / tot.clus *100)
+				# print(pct.x)
+				enrich.df[row.id, col.id] <- pct.x
+			}
+		}
+		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
+		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
+		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		print(enrich.df)
+	}
+}
+
+```
+Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
+
+## Find old cells on UMAP
+
+time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
+
+```{r fig.width = 4}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "clust.ID", 
+				# split.by = "orig.ident",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+```{r fig.width = 4}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "orig.ident", 
+				split.by = "clust.ID",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+# Total var 95%
+## Neighborhood and umap
+set total.var <- 95%
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 95))
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+
+For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
+
+Test: what percentage of each new clusterID matches one of the older clusters?
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
+		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
+		colnames(enrich.df) <- c(3, 17, 10, 11)
+		rownames(enrich.df) <- new.clusters
+		meta.df <- seurat.object@meta.data
+		for(row.id in rownames(enrich.df)){
+			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
+			for(col.id in colnames(enrich.df)){
+				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
+				pct.x <- as.integer(num.x / tot.clus *100)
+				# print(pct.x)
+				enrich.df[row.id, col.id] <- pct.x
+			}
+		}
+		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
+		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
+		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		print(enrich.df)
+	}
+}
+
+```
+Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
+
+## Find old cells on UMAP
+
+time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
+
+```{r fig.width = 2}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "clust.ID", 
+				pt.size = .1,
+				# split.by = "orig.ident",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+```{r fig.width = 4}
+DimPlot(seurat.object,
+				reduction = "umap",
+				group.by = "orig.ident", 
+				split.by = "clust.ID",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+### Gene expression of old clustrs on new map
+Let's see if we can get some gene expression profiles on these...
+```{r, fig.height=10, fig.width=18}
+gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
+VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
+```
+
+## Evaluate cluster stability
+Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
+
+Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
+
+### Clustree
+Previously my favourite has been Clustree, which gives a nice visual
+NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+These data suggest that node stability is aweful! Need to figure out if this is a dimensional reduction error or a clustering error.
+
+# Why are clusters so unstable?
+Differences could include:
+* cells in each population (cellranger v6 includes more cells than cellranger v1, especially in MEP)
+* dimensionality is incorrect
+* ScaleData didnt account for regression factors (e.g., "nCounts_RNA" or "nFeatures_RNA")
+* Did not consider cell cycle
+* Incorrect normalization/scaling method
+* Clustering is too strict or not strict enough
+* neighborhood analysis used wrong parameters
+* Should include mitoC filter (there's a chunk of MEP w/ mitoC @ ~40%)
+* SCTransform accounts better for sources of variability
+
+```{r}
+# Number of filtered cells left in each pop
+sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
+```
+
+```{r}
+for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
+	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
+	h$density = h$counts/sum(h$counts)*100
+	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
+}
+```
+
+Looks like MEPm is the only sample with that huge MitoC % lump @ 40%. What do these cells look like, otherwise?
+
+```{r fig.width=5}
+VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
+				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
+```
+Save dim36 as is and try clustering analysis @ dim24
+```{r}
+saveRDS(seurat.object, file = "msAggr_AnalysisCode/msAggr_dim36.rds")
+```
+
+# Repeat clustering with dim24
+One possibility is that I've included too many dimensions. Will see if 90% increases stability.
+
+```{r}
+seurat.object <- FindNeighbors(seurat.object, dims = 1:24)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1:24)
+```
+
+Save object
+```{r}
+saveRDS(seurat.object, file = "msAggr_dim24.rds")
+```
+
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
+											) + 
+			ggtitle(paste0("msAggr dim36 res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+## Evaluate cluster stability
+Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
+
+Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
+
+### Clustree
+Previously my favourite has been Clustree, which gives a nice visual
+NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+Think I'll explore regression factors using SCTransform in new document.

### 2021-07-07 noMtThreshold/msAggr_sctransform-v1.Rmd

--- 2021-06-16 msAggr_AnalysisCode/msAggr_sctransform-v1.Rmd
@@
 title: "msAggr_sctransform-v1"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
@@
 library(ggplot2)
-library(clustree)
+# library(clustree)
 ```
@@
 
+########IS SCTRANSFORM MORE ROBUST FOR DIFFERENT SEQUENCING DEPTHS??#########
+
 Load 4 populations into single seurat object
@@
 ```{r}
-remove(seurat.object.data)
+remove(object.data)
 ```
@@
 seurat.object <- PercentageFeatureSet(seurat.object, pattern = '^mt-', col.name = "percent.mt")
+```
+
+## Save base object
+```{r}
+saveRDS(seurat.object, file = "msAggr_rawSeurat.rds")
 ```
@@
 ```{r}
+DimPlot(seurat.object, group.by = "orig.ident", reduction = "umap")
+```
+
+```{r}
 for (meta.col in colnames(seurat.object@meta.data)){
@@
 }
+```
+
+```{r}
+fav.genes <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1", "proc")
+```
+
+```{r}
+FeaturePlot(seurat.object, features = fav.genes, ncol = 3)
 ```
@@
 }
+```
+
+# Why are clusters so unstable?
+Differences could include:
+* cells in each population (cellranger v6 includes more cells than cellranger v1, especially in MEP)
+* --dimensionality is incorrect--
+* ScaleData didnt account for regression factors (e.g., "nCounts_RNA" or "nFeatures_RNA")
+* Incorrect normalization/scaling method
+* --Clustering is too strict or not strict enough--
+* neighborhood analysis used wrong parameters
+* Should include mitoC filter (there's a chunk of MEP w/ mitoC @ ~40%)
+* --SCTransform accounts better for sources of variability--
+
+# Regress out cell cycle
+One thing I've probably done in the past is regress out cell cycle factors. Did not do that here. Repeat analysis but include cell cycle regression.  
+Cell cycle scoring comes after `RunPCA()`
+
+Step 1: load in raw object
+```{r}
+seurat.object <- readRDS("msAggr_rawSeurat.rds")
+```
+
+Since `SCTransform()` is best for different correcting different sequencing depths, going to stick with that
+
+The [Cell Cycle Regression Vignette](https://satijalab.org/seurat/articles/cell_cycle_vignette.html) says to do `NormalizeData()`, `FindVariableFeatures()`, `ScaleData()`, `RunPCA()`, `CellCycleScoring()`, then `RunPCA()` with s.genes and g2m.genes.
+It seems to be more complicated with `SCTransform()` because you don't want to normalize normalized data, but you need normalized data to properly run `CellCycleScoring()`.
+Found discussion board and [this](https://github.com/satijalab/seurat/issues/1679#issuecomment-557781838) is the answer that makes the most sense to me
+
+```{r results='asis'}
+# normalize data with SCTransform()
+sample <- SCTransform(
+  sample,
+  assay = 'RNA',
+  new.assay.name = 'SCT',
+  vars.to.regress = c('percent.mt', 'nFeature_RNA', 'nCount_RNA')
+)
+
+# perform cell cycle analysis (make sure to specify the "assay" parameter
+sample <- CellCycleScoring(
+  sample,
+  s.features = s.genes,
+  g2m.features = g2m.genes,
+  assay = 'SCT',
+  set.ident = TRUE
+)
+
+# normalise again but this time including also the cell cycle scores
+sample <- SCTransform(
+  sample,
+  assay = 'RNA',
+  new.assay.name = 'SCT',
+  vars.to.regress = c('percent.mt', 'nFeature_RNA', 'nCount_RNA', 'S.Score', 'G2M.Score')
+)
+```
+
+so let's try this
+
+```{r}
+seurat.object <- SCTransform(seurat.object, assay = 'RNA', new.assay.name = "SCT", method = "glmGamPoi", vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
+```
+
+Cell cycle scoring
+```{r}
+seurat.object <- CellCycleScoring(seurat.object, s.features = cc.genes$s.genes, g2m.features = cc.genes$g2m.genes, assay = 'SCT', set.ident = FALSE)
+```

### 2021-08-12 CMPSubset.Rmd
Update for index sort parameters

@@
 ```{r}
-seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
+cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
 ```
@@
 ```{r}
-seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
-VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
+cmp.object[["percent.mt"]] <- PercentageFeatureSet(cmp.object, pattern = "^mt-")
+VlnPlot(cmp.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
 ```
@@
 ```{r}
-plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
-plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
+plot1 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
+plot2 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
 plot1 + plot2
@@
 ```{r}
-print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
-seurat.object <- subset(seurat.object, 
+print(paste("original object:", nrow(cmp.object@meta.data), "cells", sep = " "))
+cmp.object <- subset(cmp.object, 
 												subset = nFeature_RNA >=200 & 
@@
 												)
-print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
+print(paste("new object:", nrow(cmp.object@meta.data), "cells", sep = " "))
 ```
@@
 ```{r}
-seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
+cmp.object <- NormalizeData(cmp.object, normalization.method = "LogNormalize", scale.factor = 10000)
 ```
@@
 ```{r fig.width = 5, fig.height = 2}
-seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
-top10 <- head(VariableFeatures(seurat.object), 10)
-plot1 <- VariableFeaturePlot(seurat.object)
+cmp.object <- FindVariableFeatures(cmp.object, selection.method = "vst", nfeatures = 2000)
+top10 <- head(VariableFeatures(cmp.object), 10)
+plot1 <- VariableFeaturePlot(cmp.object)
 plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
@@
 ```{r echo = FALSE}
-all.genes <- rownames(seurat.object)
-seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
+all.genes <- rownames(cmp.object)
+cmp.object <- ScaleData(cmp.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
 ```
@@
 ```{r}
-saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
+saveRDS(cmp.object, file = paste0(projectName, "_raw.RDS"))
 ```
@@
 ```{r}
-seurat.object <- RunPCA(seurat.object, features = VariableFeatures(seurat.object), ndims.print = 1:5, nfeatures.print = 5)
+cmp.object <- RunPCA(cmp.object, features = VariableFeatures(cmp.object), ndims.print = 1:5, nfeatures.print = 5)
 ```
@@
 ```{r}
-DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
-VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
+DimPlot(cmp.object, reduction = "pca", group.by = "orig.ident")
+VizDimLoadings(cmp.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
 
@@
 ```{r, figures-side, fig.show='hold', out.width="50%"}
-ElbowPlot(seurat.object, ndims = 50)
-percent.variance(seurat.object@reductions$pca@stdev)
+ElbowPlot(cmp.object, ndims = 50)
+percent.variance(cmp.object@reductions$pca@stdev)
 ```
@@
 ```{r}
-tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
 paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
@@
 ```{r}
-seurat.object@meta.data['clust.ID'] <- 0
-head(seurat.object@meta.data)
+cmp.object@meta.data['clust.ID'] <- 0
+head(cmp.object@meta.data)
 ```
@@
 ```{r}
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
+cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust3.cells] <- 3
+cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust17.cells] <- 17
+cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust10.cells] <- 10
+cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust11.cells] <- 11
 ```
@@
 ```{r}
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 10,])
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 11,])
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 17,])
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
+nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 10,])
+nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 11,])
+nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 17,])
+nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 3,])
 ```
@@
 ```{r}
-tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
 ndims <- length(which(cumsum(tot.var) <= 90))
 
-seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
-seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
+cmp.object <- FindClusters(cmp.object, resolution = 0.5)
+cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
 
-saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 for(x in c(0.5, 1, 1.5, 2, 2.5)){
-	seurat.object <- FindClusters(seurat.object, resolution = x)
+	cmp.object <- FindClusters(cmp.object, resolution = x)
 }
@@
 ```{r}
-for (meta.col in colnames(seurat.object@meta.data)){
+for (meta.col in colnames(cmp.object@meta.data)){
 	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		myplot <- DimPlot(seurat.object, 
+		myplot <- DimPlot(cmp.object, 
 											group.by = meta.col,
@@
 
+for each resolution, number/percentage of cells in each cluster?
+
+```{r}
+tot.cells <- nrow(cmp.object@meta.data)
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
+		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
+		colnames(stats.df) <- c("num_cells", "pct_pop")
+		rownames(stats.df) <- new.clusters
+		meta.df <- cmp.object@meta.data
+		for(row.id in rownames(stats.df)){
+				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
+				pct.x <- as.integer(num.x / tot.cells *100)
+				# print(pct.x)
+				stats.df[row.id, "num_cells"] <- num.x
+				stats.df[row.id, "pct_pop"] <- pct.x
+		}
+		print(stats.df)
+	}
+}
+```
+
 For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
@@
 ```{r}
-for (meta.col in colnames(seurat.object@meta.data)){
+for (meta.col in colnames(cmp.object@meta.data)){
 	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
+		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
 		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
@@
 		rownames(enrich.df) <- new.clusters
-		meta.df <- seurat.object@meta.data
+		meta.df <- cmp.object@meta.data
 		for(row.id in rownames(enrich.df)){
@@
 ```{r fig.width = 4}
-DimPlot(seurat.object,
+DimPlot(cmp.object,
 				reduction = "umap",
@@
 ```{r}
-tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
 ndims <- length(which(cumsum(tot.var) <= 95))
 
-seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
-seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
+cmp.object <- FindClusters(cmp.object, resolution = 0.5)
+cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
 
-saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 ```{r fig.width = 2}
-DimPlot(seurat.object,
+DimPlot(cmp.object,
 				reduction = "umap",
@@
 gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
-VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
+VlnPlot(cmp.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
 ```
@@
 for(tested.ident in ident.list){
-	Idents(seurat.object) <- tested.ident
-	all.markers <- FindAllMarkers(seurat.object)
+	Idents(cmp.object) <- tested.ident
+	all.markers <- FindAllMarkers(cmp.object)
 	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
@@
 ```{r}
-Idents(seurat.object) <- "RNA_snn_res.2.5"
-res.2.5.allmarkers <- FindAllMarkers(seurat.object)
+Idents(cmp.object) <- "RNA_snn_res.2.5"
+res.2.5.allmarkers <- FindAllMarkers(cmp.object)
 ```
@@
 for(tested.ident in ident.list){
-	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
-	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+	for (cluster in sort(as.numeric(levels(cmp.object@meta.data[[tested.ident]])))){
+	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
 	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
@@
 ```{r}
-for (cluster in sort(as.numeric(levels(seurat.object@meta.data$RNA_snn_res.2.5)))){
-	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+for (cluster in sort(as.numeric(levels(cmp.object@meta.data$RNA_snn_res.2.5)))){
+	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
 	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
@@
 ```{r}
-	Idents(seurat.object) <- "clust.ID"
-	all.markers <- FindAllMarkers(seurat.object)
+	Idents(cmp.object) <- "clust.ID"
+	all.markers <- FindAllMarkers(cmp.object)
 	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
@@
 ```{r}
-# Idents(seurat.object) <- "clust.ID"
-for (cluster in unique(seurat.object@meta.data$clust.ID)){
+# Idents(cmp.object) <- "clust.ID"
+for (cluster in unique(cmp.object@meta.data$clust.ID)){
 	print(cluster)
-	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
 	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
@@
 ```{r fig.height = 30, fig.width=6}
-VizDimLoadings(seurat.object, dims = 1:10, nfeatures = 30, reduction = "pca", ncol = 2)
+VizDimLoadings(cmp.object, dims = 1:10, nfeatures = 30, reduction = "pca", ncol = 2)
 ```
@@
 ```{r}
-pca.df <- seurat.object[["pca"]]
+pca.df <- cmp.object[["pca"]]
 pca.df <- as.data.frame(as.matrix(slot(object = pca.df, name = "feature.loadings")))
-print(seurat.object[["pca"]], dims = 2, nfeatures = 5)
+print(cmp.object[["pca"]], dims = 2, nfeatures = 5)
 rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2, decreasing = TRUE)[1:5], ])
@@
 surface.markers <- c("Cd34", "Fcgr3", "Fcgr2b", "Cd9", "Itga2b", "Cd48", "Ly6a")
-FeaturePlot(seurat.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
+FeaturePlot(cmp.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
 ```
@@
 png(filename = "FeaturePlot_CMP_surfaceMarkers_clustIDfacet.png", height = 1600, width = 1600)
-FeaturePlot(seurat.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
+FeaturePlot(cmp.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
 dev.off()
@@
 ```
+
+# Compare @ hierarchcial clusteirng
+
+Do clustering using biomark RNAs as input
+```{r}
+# Read in BiomarkRNAs
+biomark.rnas <- read.table('/Users/heustonef/Desktop/10XGenomicsData/BiomarkRNAs.txt')
+biomark.rnas <- biomark.rnas$V1
+```
+
+use biomark RNAs to define dimensional reduction
+```{r}
+cmp.object <- readRDS("CMP_raw.RDS")
+cmp.object <- RunPCA(cmp.object, features = biomark.rnas, ndims.print = 1:5, , nfeatures.print = 5)
+ElbowPlot(cmp.object, ndims = 50)
+```
+
+Now run the clustering
+```{r}
+tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 90))
+
+cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
+cmp.object <- FindClusters(cmp.object, resolution = 0.5)
+cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
+
+```
+
+find the clusters
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	cmp.object <- FindClusters(cmp.object, resolution = x)
+}
+```
+
+Plot the umaps and cell cluster ids
+```{r}
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(cmp.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+### Calculate anticipated number of cells you'll find in each biomark cluster
+Get # cells in each cluster
+
+```{r}
+tot.cellcount <- nrow(cmp.object@meta.data)
+res05.list <- sort(unique(cmp.object@meta.data$RNA_snn_res.0.5), decreasing = FALSE)
+sapply(res05.list, 
+			 function(x){
+			 	print(
+			 		paste(
+			 			"cluster", x, "=", 
+			 			nrow(cmp.object@meta.data[cmp.object@meta.data$RNA_snn_res.0.5 == x,]), 
+			 			"cells or", 
+			 			round(nrow(cmp.object@meta.data[cmp.object@meta.data$RNA_snn_res.0.5 == x,])/tot.cellcount*100, digits = 2), 
+			 			"% of total"
+			 		)
+			 	)
+			 }
+			)
+```
+
+So we did the dimensional reduction based on the biomark RNAs, then did our UMAP nearest neighbor clustering.
+
+In the biomark hierarchcial clustering analysis I assayed 167 cells.
+The smallest cluster I detected had 3 cells, or 1.8% of total, and this is an uncomfortably small number of cells.
+Based on the UMAP calculations I would therefore expect to find 11 or 12 of the predicted 15 clusters.
+I found 12, and I don't really like that last one, so 11 or 12.
+Since I did the hierarchcial clustering yesterday and did this math today, we can say it was independent of these results and therefore totally legit.
+Yay!!

### 2021-08-13 CMPSubset.Rmd
updating analysis
Reorganizing repositories

@@
 ---
+
+# Updates
+
+## Update 2021.08.11
+Most thorough way to readdress differences between original and current analyses is to jsut redo the biomark.
+To that end, will restart analysis of *just* CMPm2, excluding other datasets (i.e., LSK, MEP, GMP).
+Will also keep CMP3K as test set/ bio rep 1
+
+# Notebook setup
 
@@
 ```
-store session info
-```{r }
-# sink("CMP-v1.20210616.txt")
-sessionInfo()
-# sink()
-```
 
@@
 jackstraw.dim <- 40
+```
+
+store session info
+```{r }
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
 ```

### 2021-08-13 noMtThreshold/msAggr_seurat_noMt-v1.Rmd
cleaning repositories

--- 2021-06-16 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
@@
 * ScaleData didnt account for regression factors (e.g., "nCounts_RNA" or "nFeatures_RNA")
+* Did not consider cell cycle
 * Incorrect normalization/scaling method

### 2021-08-13 CMP/CMPSubset.Rmd
is this cleaning or breaking
more reorganizing

--- 2021-08-13 CMPSubset.Rmd
@@
 To that end, will restart analysis of *just* CMPm2, excluding other datasets (i.e., LSK, MEP, GMP).
-Will also keep CMP3K as test set/ bio rep 1
+Will also keep
 
@@
 ```{r setup}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
 library(dplyr)
@@
 ```{r }
-sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
-sink(sessionInfo.filename)
+# sink("CMP-v1.20210616.txt")
 sessionInfo()
-sink()
+# sink()
 ```

### 2021-08-13 msAggr/msAggr_seurat.Rmd
fix or break fix or break

--- 2021-07-07 msAggr_seurat.Rmd
@@
 
-```{r setup}
+```{r setup, message=FALSE, warning=FALSE}
 knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
@@
 ndims <- length(which(cumsum(tot.var) <= 90))
+print(ndims)
 
@@
 
-saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+# saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 }
+```
+
+Generate statistics for each cluster/resolution combo
+```{r}
+current_res <- 'RNA_snn_res.0.5'
+
+cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
+for(id in cluster_ids){
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "LSKm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "CMPm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
+	counts_df[id, "MEPm"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
+	counts_df[id, "GMPm"] = cell_value
+
+}
+
 ```
@@
 		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
-		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		try{xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)}
 		print(enrich.df)
@@
 ndims <- length(which(cumsum(tot.var) <= 95))
+print(ndims)
 

### 2021-08-16 CMP/CMPSubset.Rmd
Looking for new Taqman probes; Identifying clustree optimal resolution

@@
 library(ggplot2)
-# library(clustree)
+library(clustree)
 ```
@@
 store session info
-```{r }
-# sink("CMP-v1.20210616.txt")
+```{r}
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
 sessionInfo()
-# sink()
+sink()
 ```
@@
 ```{r}
-source("msAggr_AnalysisCode/read_10XGenomics_data.R")
-source("msAggr_AnalysisCode/PercentVariance.R")
-source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
+source("../RFunctions/read_10XGenomics_data.R")
+source("../RFunctions/PercentVariance.R")
+source("../RFunctions/Mouse2Human_idconversion.R")
 ```
 
-```{r}
-setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
+```{r warning=FALSE}
+setwd("../../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
 data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
@@
 ```
-
-We don't have to worry about comparing library depths, so we'll just do normalization/Scale data
-
+## Filter data
 remove low quality cells
@@
 plot1 + plot2
-
 ```
 
-Scale data (linear transformation)
+## Scale data (linear transformation)
+We don't have to worry about comparing library depths, so we'll just do normalization/Scale data
 
@@
 ```{r}
-clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
+clust3.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
 clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
+clust17.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
 clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
+clust10.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
 clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
+clust11.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
 clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
@@
 nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 3,])
+```
+
+## Identify variable genes for new biomark
+```{r}
+length(cmp.object@assays$RNA@var.features)
+```
+```{r}
+top100.var <- cmp.object@assays$RNA@var.features[1:100]
+length(top100.var)
 ```
@@
 ```
+
+### Clustree
+what's the max resolution we can achieve while keepign clusters stable?
+```{r fig.width = 10, fig.height = 5}
+clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
+    scale_color_continuous(low = 'red3', high = 'white')
+```
+I think I'm liking res.1.0 from this
 
@@
 Create `FindAllMarkers()` lists for GSEA
-```{r}
+```{r }
 Idents(cmp.object) <- "RNA_snn_res.2.5"
@@
 
-Map HGNC symbols
+## Map HGNC symbols
 ```{r}
@@
 for(tested.ident in ident.list){
-	for (cluster in sort(as.numeric(levels(cmp.object@meta.data[[tested.ident]])))){
-	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
+	object.copy <- cmp.object
+	Idents(object.copy) <- tested.ident
+	print(paste("testing", tested.ident))
+	for (cluster in sort(as.numeric(levels(object.copy@meta.data[[tested.ident]])))){
+		print(paste("looking at cluster", cluster))
+		cluster.markers <- FindMarkers(object.copy, ident.1 = cluster)
+		try(
 			xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
@@
 											 append = TRUE)
+		)	
 	}
+	remove(object.copy)
 }
@@
 
-# Cell cycle analysis
-Just so we know what we're working with
-
-```{r}
-s.genes <- cc.genes.updated.2019$s.genes
-g2m.genes <- cc.genes.updated.2019$g2m.genes
-```
-
-# Compare @ hierarchcial clusteirng
+# Compare @g hierarchcial clusteirng
 

### 2021-08-16 CMP/CMPSubset.Rmd
continued clustree analysis

@@
 
-### Color palette
-```{r}
-color.palette <- c(
-	"coral",
-	"chartreuse4",
-	"goldenrod1",
-	"cadetblue1",
-	"burlywood",
-	"brown",
-	"brown1",
-	"blue",
-	"blue4",
-	"azure3",
-	"aquamarine",
-	"antiquewhite",
-	"cadetblue",
-	"gold3",
-	"black",
-	"darkgreen",
-	"deeppink",
-	"darkviolet",
-	"darkturquoise",
-	"darkslategray",
-	"darksalmon",
-	"darkorchid1",
-	"darkolivegreen2",
-	"forestgreen",
-	"dodgerblue",
-	"green",
-	"lightpink",
-	"lightcoral",
-	"khaki1",
-	"maroon",
-	"peru",
-	"lightseagreen",
-	"lightsalmon",
-	"plum",
-	"moccasin",
-	"tan",
-	"tan1", 
-	"red", 
-	"purple",
-	"khaki4",
-	"black", 
-	"plum4"
-)
-```
-
 # Total var 90%
@@
 ndims <- length(which(cumsum(tot.var) <= 90))
+
+cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
+cmp.object <- FindClusters(cmp.object, resolution = 0.5)
+cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
+
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	cmp.object <- FindClusters(cmp.object, resolution = x)
+}
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+
+```{r}
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(cmp.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+### Clustree
+what's the max resolution we can achieve while keepign clusters stable?
+```{r fig.width = 10, fig.height = 5}
+plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
+my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
+	scale_color_continuous(low = 'red3', high = 'white') + 
+	ggtitle(plot.title)
+png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
+plot(my.clustree)
+dev.off()
+
+```
+I think I'm liking res.1.0 from this. Although how much does this change if I use fewer PCs...
+
+for each resolution, number/percentage of cells in each cluster?
+
+```{r}
+tot.cells <- nrow(cmp.object@meta.data)
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
+		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
+		colnames(stats.df) <- c("num_cells", "pct_pop")
+		rownames(stats.df) <- new.clusters
+		meta.df <- cmp.object@meta.data
+		for(row.id in rownames(stats.df)){
+				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
+				pct.x <- as.integer(num.x / tot.cells *100)
+				# print(pct.x)
+				stats.df[row.id, "num_cells"] <- num.x
+				stats.df[row.id, "pct_pop"] <- pct.x
+		}
+		print(stats.df)
+	}
+}
+```
+
+For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
+
+Test: what percentage of each new clusterID matches one of the older clusters?
+```{r}
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
+		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
+		colnames(enrich.df) <- c(3, 17, 10, 11)
+		rownames(enrich.df) <- new.clusters
+		meta.df <- cmp.object@meta.data
+		for(row.id in rownames(enrich.df)){
+			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
+			for(col.id in colnames(enrich.df)){
+				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
+				pct.x <- as.integer(num.x / tot.clus *100)
+				# print(pct.x)
+				enrich.df[row.id, col.id] <- pct.x
+			}
+		}
+		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
+		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
+		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		print(enrich.df)
+	}
+}
+
+```
+Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
+
+## Find old cells on UMAP
+
+time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
+
+```{r fig.width = 4}
+DimPlot(cmp.object,
+				reduction = "umap",
+				group.by = "clust.ID", 
+				# split.by = "orig.ident",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+```{r fig.width = 4}
+DimPlot(cmp.object,
+				reduction = "umap",
+				group.by = "orig.ident", 
+				split.by = "clust.ID",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+# Total var 85%
+## Neighborhood and umap
+set total.var <- 85%
+```{r}
+tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 85))
 
@@
 ```{r fig.width = 10, fig.height = 5}
-clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
-    scale_color_continuous(low = 'red3', high = 'white')
+plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
+my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
+	scale_color_continuous(low = 'red3', high = 'white') + 
+	ggtitle(plot.title)
+png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
+plot(my.clustree)
+dev.off()
 ```
-I think I'm liking res.1.0 from this
 
@@
 # Total var 95%
+
 ## Neighborhood and umap
@@
 }
+```
+
+```{r}
+cmp.object <- readRDS("CMP_dim41.RDS")
+```
+
+```{r fig.width = 10, fig.height = 5}
+plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
+my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
+	scale_color_continuous(low = 'red3', high = 'white') + 
+	ggtitle(plot.title)
+png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
+plot(my.clustree)
+dev.off()
 ```

### 2021-08-16 RFunctions/ColorPalette.R
create color palette function

@@ -1,0 +1,44 @@
+color.palette <- c(
+	"coral",
+	"chartreuse4",
+	"goldenrod1",
+	"cadetblue1",
+	"burlywood",
+	"brown",
+	"brown1",
+	"blue",
+	"blue4",
+	"azure3",
+	"aquamarine",
+	"antiquewhite",
+	"cadetblue",
+	"gold3",
+	"black",
+	"darkgreen",
+	"deeppink",
+	"darkviolet",
+	"darkturquoise",
+	"darkslategray",
+	"darksalmon",
+	"darkorchid1",
+	"darkolivegreen2",
+	"forestgreen",
+	"dodgerblue",
+	"green",
+	"lightpink",
+	"lightcoral",
+	"khaki1",
+	"maroon",
+	"peru",
+	"lightseagreen",
+	"lightsalmon",
+	"plum",
+	"moccasin",
+	"tan",
+	"tan1", 
+	"red", 
+	"purple",
+	"khaki4",
+	"black", 
+	"plum4"
+)

### 2021-08-16 CMP/CMPSubset.Rmd
Generate Biomark candidates for 85% var (dim18), res1 and res1.5

@@
 source("../RFunctions/Mouse2Human_idconversion.R")
+source ("../RFunctions/ColorPalette.R")
 ```
@@
 
-## Identify variable genes for new biomark
-```{r}
-length(cmp.object@assays$RNA@var.features)
-```
-```{r}
-top100.var <- cmp.object@assays$RNA@var.features[1:100]
-length(top100.var)
-```
-
 # Total var 90%
@@
 
-saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 	cmp.object <- FindClusters(cmp.object, resolution = x)
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+}
+```
+
+```{r}
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(cmp.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+### Clustree
+what's the max resolution we can achieve while keepign clusters stable?
+```{r fig.width = 10, fig.height = 5}
+plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
+my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
+	scale_color_continuous(low = 'red3', high = 'white') + 
+	ggtitle(plot.title)
+png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
+plot(my.clustree)
+dev.off()
+```
+
+for each resolution, number/percentage of cells in each cluster?
+
+```{r}
+tot.cells <- nrow(cmp.object@meta.data)
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
+		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
+		colnames(stats.df) <- c("num_cells", "pct_pop")
+		rownames(stats.df) <- new.clusters
+		meta.df <- cmp.object@meta.data
+		for(row.id in rownames(stats.df)){
+				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
+				pct.x <- as.integer(num.x / tot.cells *100)
+				# print(pct.x)
+				stats.df[row.id, "num_cells"] <- num.x
+				stats.df[row.id, "pct_pop"] <- pct.x
+		}
+		print(stats.df)
+	}
+}
+```
+
+## Identify variable genes for new biomark
+```{r}
+length(cmp.object@assays$RNA@var.features)
+```
+
+### Gene profiles of clusters
+set ident at res = 1 and get markers
+```{r}
+Idents(cmp.object) <- "RNA_snn_res.1"
+cmp.allmarkers.res1 <- FindAllMarkers(cmp.object)
+cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
+```
+
+```{r}
+cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
+cmp.top100markers.res1 <- cmp.top100markers.res1[cmp.top100markers.res1$p_val_adj <= 0.05, ]
+for(cluster in unique(cmp.top100markers.res1$cluster)){
+	num.hits <- nrow(cmp.top100markers.res1[cmp.top100markers.res1$cluster == cluster, ])
+	print(paste("cluster", cluster, "has", num.hits, "genes"))
+}
+```
+
+Find top.var	genes in cmp.top100markers.res1
+```{r}
+top.var <- cmp.object@assays$RNA@var.features[1:150]
+# top.var
+n.vargenes <- c()
+for(gene in top.var){
+	if(gene %in% cmp.top100markers.res1$gene){
+		n.vargenes <- c(n.vargenes, gene)
+	}
+}
+print(length(n.vargenes))
+top.var[!(top.var %in% n.vargenes)]
+```
+
+Lets try to export these, with genes on rows and clusters on columns, and xpression value in data table
+```{r}
+var.df <- cmp.top100markers.res1[cmp.top100markers.res1$gene %in% n.vargenes,]
+var.df <- subset.data.frame(var.df, select = c("gene", "cluster", "avg_log2FC"))
+var.df <- reshape2::dcast(var.df, gene~cluster, value.var = "avg_log2FC")
+xlsx::write.xlsx(var.df, file = paste0(projectName, "_dim", ndims, "_candididateBiomarkGenes.xlsx"), sheetName = "res1", append = TRUE, row.names = FALSE)
+
+```
+
+repeat for top.var	genes in cmp.top100markers.res1.5
+```{r}
+top.var <- cmp.object@assays$RNA@var.features[1:150]
+Idents(cmp.object) <- "RNA_snn_res.1.5"
+cmp.allmarkers.res1.5 <- FindAllMarkers(cmp.object)
+cmp.top100markers.res1.5 <- cmp.allmarkers.res1.5 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
+n.vargenes <- c()
+for(gene in top.var){
+	if(gene %in% cmp.top100markers.res1.5$gene){
+		n.vargenes <- c(n.vargenes, gene)
+	}
+}
+print(length(n.vargenes))
+top.var[!(top.var %in% n.vargenes)]
+```
+
+Lets try to export these, with genes on rows and clusters on columns, and xpression value in data table
+```{r}
+var.df <- cmp.top100markers.res1.5[cmp.top100markers.res1.5$gene %in% n.vargenes,]
+var.df <- subset.data.frame(var.df, select = c("gene", "cluster", "avg_log2FC"))
+var.df <- reshape2::dcast(var.df, gene~cluster)
+xlsx::write.xlsx(var.df, file = paste0(projectName, "_dim", ndims, "_candididateBiomarkGenes.xlsx"), sheetName = "res1.5", append = TRUE, row.names = FALSE)
+
+```
+For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
+
+Test: what percentage of each new clusterID matches one of the older clusters?
+```{r}
+for (meta.col in colnames(cmp.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
+		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
+		colnames(enrich.df) <- c(3, 17, 10, 11)
+		rownames(enrich.df) <- new.clusters
+		meta.df <- cmp.object@meta.data
+		for(row.id in rownames(enrich.df)){
+			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
+			for(col.id in colnames(enrich.df)){
+				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
+				pct.x <- as.integer(num.x / tot.clus *100)
+				# print(pct.x)
+				enrich.df[row.id, col.id] <- pct.x
+			}
+		}
+		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
+		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
+		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
+		print(enrich.df)
+	}
+}
+
+```
+Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
+
+## Find old cells on UMAP
+
+time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
+
+```{r fig.width = 4}
+DimPlot(cmp.object,
+				reduction = "umap",
+				group.by = "clust.ID", 
+				# split.by = "orig.ident",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+```{r fig.width = 4}
+DimPlot(cmp.object,
+				reduction = "umap",
+				group.by = "orig.ident", 
+				split.by = "clust.ID",
+				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+# Total var 80%
+## Neighborhood and umap
+set total.var <- 80%
+```{r}
+tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 80))
+
+cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
+cmp.object <- FindClusters(cmp.object, resolution = 0.5)
+cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
+
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	cmp.object <- FindClusters(cmp.object, resolution = x)
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 }
@@
 }
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```

### 2021-08-19 CMP/CMPSubset.Rmd
Add dim25 biomark targets

@@
 ndims <- length(which(cumsum(tot.var) <= 85))
-
+```
+
+```{r}
 cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
@@
 cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
-
 ```
@@
 	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		plot.title <- paste0(projectName, "dim", ndims, "res", gsub("RNA_snn_res", "", meta.col))
 		myplot <- DimPlot(cmp.object, 
@@
 											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+			ggtitle(plot.title)
 		plot(myplot)
+		png(filename = paste0(plot.title, ".png"), height = 800, width = 800)
+		plot(DimPlot(cmp.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette, 
+											pt.size = 1.5
+											) + 
+			ggtitle(plot.title)
+			)
+		dev.off()
 	}
@@
 
+Here's the list of genes, plus some controls
+```{r}
+res1.biomark.genes <- c("Aif1",
+"Aqp1",
+"Birc5",
+"Ccl3",
+"Ccr2",
+"Cd74",
+"Cdc20",
+"Cenpf",
+"Cst3",
+"Ctsh",
+"Dntt",
+"Elane",
+"Ermap",
+"F13a1",
+"Fcer1g",
+"Gm15915",
+"Gm17590",
+"H2-Aa",
+"H2-Eb1",
+"H2afx",
+"H2afy",
+"Hist1h2ac",
+"Hmgb2",
+"Hp",
+"Ighm",
+"Irf8",
+"Lgals3",
+"Lmo4",
+"Ms4a2",
+"Mt1",
+"Plac8",
+"Prtn3",
+"Rap1b",
+"Rgs1",
+"Vwf",
+"Wfdc17",
+"Csrp3",
+"Hist1h2ae",
+"Ifitm1",
+"Lgals1",
+"Tmsb4x",
+"Arl6ip1",
+"Car2",
+"Ccl9",
+"Ccnb2",
+"Cd9",
+"Cenpa",
+"Cpa3",
+"Fos",
+"Hist1h2ap",
+"Ly6c2",
+"Mpo",
+"Pclaf",
+"Slpi",
+"Top2a",
+"Ube2c",
+"Ly86",
+"Hist1h2bc",
+"Pf4",
+"Apoe",
+"Ctsg",
+"Car1",
+"Hmmr", 
+"Gapdh", 
+"B2m", 
+"Actb", 
+"Hprt1",
+"Gusb")
+```
+
+make vln plots per cluster
+```{r}
+for(gene in res1.biomark.genes){
+	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
+	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
+	png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
+	plot(myplot)
+	dev.off()
+
+}
+```
+
+## Res 1.5
+
 repeat for top.var	genes in cmp.top100markers.res1.5
@@
 				cols = c("gray", "orange", "blue", "red", "green"),)
+```
+
+### Gene expression of old clustrs on new map
+Let's see if we can get some gene expression profiles on these...
+```{r, fig.height=10, fig.width=18}
+gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
+VlnPlot(cmp.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
 ```

### 2021-08-20 CMP/CMPSubset.Rmd
Res1 dge analysis for biomark probes

@@
 
+## Update 2021.08.19
+Pursuing Dim25 (85% of variance explained) res1 and res1.5 for biomark targets
+
 ## Update 2021.08.11
@@
 
+## Load libraries
 ```{r setup}
@@
 
-# Set global variables
+## Set global variables
 ```{r}
@@
 
-store session info
+## Store session info
 ```{r}
@@
 
+## Load local scripts
 ```{r}
@@
 
+# Read CMPm2
 ```{r warning=FALSE}
@@
 what's the max resolution we can achieve while keepign clusters stable?
-```{r fig.width = 10, fig.height = 5}
+```{r fig.width = 15, fig.height = 5}
 plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
@@
 	ggtitle(plot.title)
+plot(my.clustree)
 png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
@@
 Idents(cmp.object) <- "RNA_snn_res.1"
+```
+
+```{r}
 cmp.allmarkers.res1 <- FindAllMarkers(cmp.object)
@@
 cmp.top100markers.res1 <- cmp.top100markers.res1[cmp.top100markers.res1$p_val_adj <= 0.05, ]
-for(cluster in unique(cmp.top100markers.res1$cluster)){
+for(cluster in sort(as.numeric(unique(cmp.top100markers.res1$cluster)))){
 	num.hits <- nrow(cmp.top100markers.res1[cmp.top100markers.res1$cluster == cluster, ])
 	print(paste("cluster", cluster, "has", num.hits, "genes"))
+		cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
+		try(
+			xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+											 file = paste0(projectName, "_dim", ndims, "_FindMarkersTop100.xlsx"), 
+											 sheetName = paste0("clst", cluster), 
+											 col.names = TRUE, 
+											 row.names = TRUE, 
+											 append = TRUE)
+		)	
 	}
@@
 												"Car1",
-"Hmmr", 
-"Gapdh", 
-"B2m", 
-"Actb", 
-"Hprt1",
-"Gusb")
+												"Hmmr")
+hkgenes <- c("Gapdh", "B2m", "Hprt", "Pgk1", "Rplp2", "Pgk1", "Ubc", "Ywhaz", "Ppia", "Pum1", "Psmc4", "Elf1", "Mrpl19")
+surface.markers<- c("Cd34", 
+										"Kit",
+										"Cd48", 
+										"Ly6a",
+										"Ly6e",
+										"Cd9",
+										"Itga2b",
+										"Ly86",
+										"Itga6",
+										"Cd55",
+										"Slamf1",
+										"Flt3"
+										)
 ```
@@
 	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
-	png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
+	png(filename = paste0("VlnPlots/", plot.title, ".png"), height = 800, width = 1600)
 	plot(myplot)
 	dev.off()
-
-}
+}
+for(gene in hkgenes){
+	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
+	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
+	png(filename = paste0("VlnPlots/", plot.title, "-hkgene.png"), height = 800, width = 1600)
+	plot(myplot)
+	dev.off()
+}
+for(gene in surface.markers){
+	plot.title <- paste0(projectName, "dim", ndims, "res1_", gene)
+	myplot<- VlnPlot(cmp.object, features = gene, pt.size = 0.01) + ggtitle(plot.title)
+	png(filename = paste0("VlnPlots/", plot.title, "-surfMkr.png"), height = 800, width = 1600)
+	plot(myplot)
+	dev.off()
+}
+```
+
+Plot candidate control genes
+```{r fig.height=20, fig.width=10}
+# VlnPlot(cmp.object, features = hkgenes, pt.size = 0.001, same.y.lims = FALSE, ncol = 2)
+png(filename = paste0(projectName, "dim", ndims, "res1_TaqmanControls.png"), height = 1600, width = 2400)
+VlnPlot(cmp.object, features = hkgenes, pt.size = 0.01, same.y.lims = FALSE, ncol = 5)
+dev.off()
 ```

### 2021-09-13 pseudobulk_vs_bulk/pseudobulk_CMP.Rmd
create pseudobulk branch

---
title: "pseudobulk_CMP"
output: html_notebook
---

# Purpose
Determine _ranked spearman correlation_ between single cell RNASeq (scRNASeq) and bulk RNASeq (bRNASeq)

## Method
1. Perform pseudo-bulk analysis from scRNASeq of population (i.e., CMP)
2. Download bRNASeq of same population (has 2 replicates)
3. Ranked spearman correlation between bRNASeq replicates
4. Ranked spearman correlation between each bRNASeq replicate and pseudobulk scRNASeq sample
5. Ranked spearman correlation between combined bRNASeq replicates and pseudobulk scRNASeq sample

# Notebook setup

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

## Load libraries
```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(SingleCellExperiment)
library(Matrix.utils)
library(DESeq2)
library(edgeR)
library(csaw)
library(tximport)
```

## Set global variables
```{r}
projectName <- "pseudobulk_CMP"
```

## Store session info
```{r}
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

## Load local scripts
```{r}
# source("../RFunctions/read_10XGenomics_data.R")
# source("../RFunctions/PercentVariance.R")
source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

# Load scRNASeq object

## Dim 25
Using CMP.rds object of dim=25, describing 85% of variation
```{r}
# CMP seurat object generated in "CMPSubset.Rmd"
seurat.object <- readRDS("CMP_dim25.RDS")
```
```{r fig.height=5, fig.width=15}
p <- FeaturePlot(seurat.object, features = c("Fli1", "Klf1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
plot(p)
png(filename = "CMP_Fli1Klf1Ratio.png", width = 2400, height = 400)
plot(p)
dev.off()
```

```{r fig.height=5, fig.width=15}
p <- FeaturePlot(seurat.object, features = c("Gata2", "Gata1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
plot(p)
png(filename = "CMP_Gata2Gata1Ratio.png", width = 2400, height = 400)
plot(p)
dev.off()
```

## Calculate pseudobulk gene expression

__ Based on tutorial @ https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html __
```{r}
# Extract raw counts and metadata to create SingleCellExperiment object

counts <- seurat.object@assays$RNA@counts

metadata <- seurat.object@meta.data

# Set up metadata as desired for aggregation and DE analysis
metadata$cluster_id <- factor(seurat.object@active.ident)

# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = counts), 
                           colData = metadata)

```

Investigate the object
```{r}
print(dim(counts(sce)))
print(dim(colData(sce)))
print(head(colData(sce)))
```

Note that QC was performed during Seurat analysis

```{r}
# sce <- perCellQCMetrics(sce)
```
## Aggregate counts
### Aggregate top-level counts

```{r}
groups <- colData(sce)[, "orig.ident"]
aggr.counts <- aggregate.Matrix(t(counts(sce)), groupings = groups, fun = "sum")

dim(aggr.counts)
```

```{r}
sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)

```

# Load bRNASeq data

```{r}
# import rsem.genes.results files

setwd("../../bRNASeq/")
file_list <- list.files(pattern = "^CMP", recursive = TRUE, include.dirs = TRUE)
names(file_list) <- lapply(file_list, function(x) gsub('.genes.results', '', x))
file_list
```

```{r}
mstxi <- tximport(files = paste0("../../bRNASeq/", file_list), type = 'rsem', txIn = FALSE, txOut = FALSE, geneIdCol = 3, )
colnames(mstxi$counts) <- gsub("^.*/", "", names(file_list))
```

```{r}
rna.counts <- merge(mstxi$counts, t(aggr.counts), by = "row.names", all = TRUE)
rna.counts[is.na(rna.counts)]<-0
rownames(rna.counts) <- rna.counts$Row.names
rna.counts <- rna.counts[, 2:ncol(rna.counts)]
head(rna.counts)
```

```{r}
dgelist_groups <- factor(c(unlist(stringr::str_extract_all(gsub("/.*", "", names(file_list)), pattern = "[a-zA-Z]+")), "SingleCell"))
levels(dgelist_groups)
```

```{r}
dgelist <- DGEList(counts = rna.counts, group = dgelist_groups)
dgelist
apply(dgelist$counts, 2, sum)
keep <- rowSums(cpm(dgelist) > 10) >=2
dgelist <- dgelist[keep,]
```

### Calculate library size
```{r}
dgelist$samples$lib.size <- colSums(dgelist$counts)
dgelist <- calcNormFactors(dgelist)
```

### Calculate dispersion

designMat
```{r}
designMat <- model.matrix(~ 0 + dgelist$samples$group)
colnames(designMat) <- levels(dgelist$samples$group)
designMat
```

estimate GLM dispersion and apply
```{r}
dgelist <- estimateGLMCommonDisp(dgelist, designMat)
dgelist <- estimateGLMTrendedDisp(dgelist, designMat, method = 'bin.spline')
dgelist <- estimateGLMTagwiseDisp(dgelist, designMat)
```

```{r}
fit <- glmQLFit(dgelist, designMat, robust = TRUE)
plotQLDisp(fit)
```

```{r}
qlf <- glmQLFTest(fit, coef = )
summary(decideTests(qlf))
```

#### rld on  individual samples

```{r}
rld <- rlog(round(dgelist$counts), blind = TRUE)
rld <- as.data.frame(rld)
rld.spear <- cor(rld, method = "spearman")
rld.melt <- reshape2::melt(rld)
colnames(rld.melt) <- c("sample", "rld")
```

```{r}
rld.spear
```

```{r}
ggplot(data = rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
```

```{r}
ggplot(data = rld, aes(x = CMPm2, y = CMP1010)) + geom_point()
```

#### rld on pooled samples

```{r}
pool.counts <- dgelist$counts
pool.counts <- as.data.frame(pool.counts)
pool.counts$ScriptSeq <- (pool.counts$CMP1009 + pool.counts$CMP1010)/2
pool.counts$TruSeq <- (pool.counts$CMP443 + pool.counts$CMP448)/2
pool.counts <- subset(pool.counts, select = c(ScriptSeq, TruSeq, CMPm2))
pool.counts <- as.matrix(pool.counts)
```

```{r}
pool.rld <- rlog(round((pool.counts)), blind = TRUE)
pool.rld.spear <- cor(pool.rld, method = "spearman")
pool.rld.melt <- reshape2::melt(pool.rld)
colnames(pool.rld.melt) <- c("sample", "rld")
```

```{r}
pool.rld.spear
```

```{r}
ggplot(data = pool.rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
```

```{r}
ggplot(data = pool.rld.melt, aes(x = CMPm2, y = ScriptSeq)) + geom_point()
```

Combine replicates from CPM table

```{r}
rna.cpm <- as.data.frame(cpm(dgelist, log = TRUE, normalized.lib.sizes = TRUE))
rna.cpm$ScriptSeq <- (rna.cpm$CMP1009 + rna.cpm$CMP1010)/2
rna.cpm$TruSeq <- (rna.cpm$CMP443 + rna.cpm$CMP448)/2
rna.cpm <- subset(rna.cpm, select = c(ScriptSeq, TruSeq, CMPm2))
```

plot correlation and scatter grams
```{r}
cpm.spear <- cor(rna.cpm, method = "spearman")
cpm.melt <- reshape2::melt(rld)
colnames(cpm.melt) <- c("sample", "rld")
```

```{r}
cpm.spear
```

## Create DESeq object from single cell data
```{r}
sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)
```

### 2021-09-13 pseudobulk_vs_bulk/pseudobulk_CMP.Rmd
appended TPM vs CPM correlation tutorial

@@
 projectName <- "pseudobulk_CMP"
+comparison_pop <- "CMP"
 ```
@@
 ```
+
+# Single cell/Bulk RNA-Seq concordance tutorial
+Following a new (tutorial) [https://jdblischak.github.io/singlecell-qtl/sc-vs-bulk.html].
+
+Calculate cpm from single cell umi matrix
+```{r}
+sc.mtx <- data.frame(rowSums(seurat.object@assays$RNA@counts))
+sc.mtx <- sc.mtx/colSums(sc.mtx)*1000000
+```
+Tutorial states that only quantity directly cmoprable between bulk and scRNA-Seq is relative abundance (cites Pachter 2001).
+
+read cmp bulk TPM
+```{r}
+# generate table with individual TPMs
+bulk.path <- "../../bRNASeq/TotalSeq/"
+file_list <- list.files(path = bulk.path, pattern = comparison_pop)
+bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]
+for(file in file_list){
+	print(file)
+	bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
+}
+
+#calculate average tpm
+bulk.pop[comparison_pop] <- rowMeans(x = bulk.pop[, c(2, 3)])
+
+# clean and check df
+bulk.pop <- bulk.pop[, -c(2, 3)]
+head(bulk.pop)
+```
+```{r}
+#calc log_rho's
+bulk.rho <- log((bulk.pop[comparison_pop] + 1), 2) - log(colSums((bulk.pop[comparison_pop])+ 1), 2)
+rownames(bulk.rho) <- bulk.pop$gene_id
+colnames(bulk.rho) <- "bulk"
+sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
+colnames(sc.rho) <- "sc"
+```
+
+check that all values are numeric
+```{r}
+print(is.numeric(bulk.rho[,1]))
+print(is.numeric(sc.rho[,1]))
+```
+## create concordance df
+### set NA = 0
+```{r}
+concordance.NAasZero <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
+concordance.NAasZero[is.na(concordance.NAasZero)] <- 0
+rownames(concordance.NAasZero) <- concordance.NAasZero$Row.names
+concordance.NAasZero <- concordance.NAasZero[, c(2, 3)]
+```
+#### plot data
+```{r}
+ggplot(data = concordance.NAasZero, aes(x = bulk, y = sc)) + 
+	geom_point()  +
+	ggtitle(paste0("concordance in ", comparison_pop))
+```
+```{r}
+cor(concordance.NAasZero, method = "spearman")
+```
+
+### drop NA
+```{r}
+concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
+concordance.df <- na.omit(concordance.df)
+rownames(concordance.df) <- concordance.df$Row.names
+concordance.df <- concordance.df[, c(2, 3)]
+```
+```{r}
+spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
+spearman.coeff
+```
+#### plot data
+```{r}
+ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
+	geom_point()  +
+	ggtitle(paste0("concordance in ", comparison_pop, " spearman coeff = ", round(spearman.coeff, digits = 3)))
+```

### 2021-09-13 pseudobulk_vs_bulk/pseudobulk_allPops.Rmd
Initiate new notebook to cycle through all 4 populations

--- 2021-09-13 pseudobulk_vs_bulk/pseudobulk_CMP.Rmd
@@
 5. Ranked spearman correlation between combined bRNASeq replicates and pseudobulk scRNASeq sample
+
+## Update
+### 2021.09.13
+Finally got a method working for CMP vs.
+TotalSeq CMP (see pseudobulk_CMP.RMD).
+Will now clean up notebook and comparisons for both TotalSeq and ScriptSeq, and for all LSK, CMP, MEP, and GMP populations.
 
@@
 ```{r}
-projectName <- "pseudobulk_CMP"
+projectName <- "pseudobulk_allPops"
 comparison_pop <- "CMP"

### 2021-09-13 CMP/CMPSubset.Rmd
Creating new generate seurat object script and updating some other things

--- 2021-08-16 CMP/CMPSubset.Rmd
@@
 library(ggplot2)
-library(clustree)
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/CMP/")
+# library(clustree)
 ```
@@
 store session info
-```{r}
-sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
-sink(sessionInfo.filename)
+```{r }
+# sink("CMP-v1.20210616.txt")
 sessionInfo()
-sink()
+# sink()
 ```
@@
 ```{r}
-source("../RFunctions/read_10XGenomics_data.R")
-source("../RFunctions/PercentVariance.R")
-source("../RFunctions/Mouse2Human_idconversion.R")
-source ("../RFunctions/ColorPalette.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
 ```
@@
 ```{r warning=FALSE}
-setwd("../../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
+setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
 data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
@@
 ```
-## Filter data
+
+We don't have to worry about comparing library depths, so we'll just do normalization/Scale data
+
 remove low quality cells
@@
 plot1 + plot2
+
 ```
 
-## Scale data (linear transformation)
-We don't have to worry about comparing library depths, so we'll just do normalization/Scale data
+Scale data (linear transformation)
 
@@
 ```{r}
-clust3.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
+clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
 clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust17.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
+clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
 clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust10.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
+clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
 clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust11.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
+clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
 clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
@@
 
+### Color palette
+```{r}
+color.palette <- c(
+	"coral",
+	"chartreuse4",
+	"goldenrod1",
+	"cadetblue1",
+	"burlywood",
+	"brown",
+	"brown1",
+	"blue",
+	"blue4",
+	"azure3",
+	"aquamarine",
+	"antiquewhite",
+	"cadetblue",
+	"gold3",
+	"black",
+	"darkgreen",
+	"deeppink",
+	"darkviolet",
+	"darkturquoise",
+	"darkslategray",
+	"darksalmon",
+	"darkorchid1",
+	"darkolivegreen2",
+	"forestgreen",
+	"dodgerblue",
+	"green",
+	"lightpink",
+	"lightcoral",
+	"khaki1",
+	"maroon",
+	"peru",
+	"lightseagreen",
+	"lightsalmon",
+	"plum",
+	"moccasin",
+	"tan",
+	"tan1", 
+	"red", 
+	"purple",
+	"khaki4",
+	"black", 
+	"plum4"
+)
+```
+
 # Total var 90%
@@
 
+saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 	cmp.object <- FindClusters(cmp.object, resolution = x)
-}
-saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
-```
-
-```{r}
-for (meta.col in colnames(cmp.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		myplot <- DimPlot(cmp.object, 
-											group.by = meta.col,
-											reduction = "umap", 
-											cols = color.palette
-											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
-		plot(myplot)
-	}
-}
-```
-
-### Clustree
-what's the max resolution we can achieve while keepign clusters stable?
-```{r fig.width = 10, fig.height = 5}
-plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
-my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
-	scale_color_continuous(low = 'red3', high = 'white') + 
-	ggtitle(plot.title)
-png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
-plot(my.clustree)
-dev.off()
-
-```
-I think I'm liking res.1.0 from this. Although how much does this change if I use fewer PCs...
-
-for each resolution, number/percentage of cells in each cluster?
-
-```{r}
-tot.cells <- nrow(cmp.object@meta.data)
-for (meta.col in colnames(cmp.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
-		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
-		colnames(stats.df) <- c("num_cells", "pct_pop")
-		rownames(stats.df) <- new.clusters
-		meta.df <- cmp.object@meta.data
-		for(row.id in rownames(stats.df)){
-				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
-				pct.x <- as.integer(num.x / tot.cells *100)
-				# print(pct.x)
-				stats.df[row.id, "num_cells"] <- num.x
-				stats.df[row.id, "pct_pop"] <- pct.x
-		}
-		print(stats.df)
-	}
-}
-```
-
-For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
-
-Test: what percentage of each new clusterID matches one of the older clusters?
-```{r}
-for (meta.col in colnames(cmp.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
-		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
-		colnames(enrich.df) <- c(3, 17, 10, 11)
-		rownames(enrich.df) <- new.clusters
-		meta.df <- cmp.object@meta.data
-		for(row.id in rownames(enrich.df)){
-			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
-			for(col.id in colnames(enrich.df)){
-				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
-				pct.x <- as.integer(num.x / tot.clus *100)
-				# print(pct.x)
-				enrich.df[row.id, col.id] <- pct.x
-			}
-		}
-		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
-		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
-		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
-		print(enrich.df)
-	}
-}
-
-```
-Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
-
-## Find old cells on UMAP
-
-time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
-
-```{r fig.width = 4}
-DimPlot(cmp.object,
-				reduction = "umap",
-				group.by = "clust.ID", 
-				# split.by = "orig.ident",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-```{r fig.width = 4}
-DimPlot(cmp.object,
-				reduction = "umap",
-				group.by = "orig.ident", 
-				split.by = "clust.ID",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-
-# Total var 85%
-## Neighborhood and umap
-set total.var <- 85%
-```{r}
-tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 85))
-
-cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
-cmp.object <- FindClusters(cmp.object, resolution = 0.5)
-cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
-
-```
-Plot UMAP
-
-```{r}
-for(x in c(0.5, 1, 1.5, 2, 2.5)){
-	cmp.object <- FindClusters(cmp.object, resolution = x)
-saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 }
@@
 }
-```
-
-### Clustree
-what's the max resolution we can achieve while keepign clusters stable?
-```{r fig.width = 10, fig.height = 5}
-plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
-my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
-	scale_color_continuous(low = 'red3', high = 'white') + 
-	ggtitle(plot.title)
-png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
-plot(my.clustree)
-dev.off()
-```
-
-for each resolution, number/percentage of cells in each cluster?
-
-```{r}
-tot.cells <- nrow(cmp.object@meta.data)
-for (meta.col in colnames(cmp.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
-		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
-		colnames(stats.df) <- c("num_cells", "pct_pop")
-		rownames(stats.df) <- new.clusters
-		meta.df <- cmp.object@meta.data
-		for(row.id in rownames(stats.df)){
-				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
-				pct.x <- as.integer(num.x / tot.cells *100)
-				# print(pct.x)
-				stats.df[row.id, "num_cells"] <- num.x
-				stats.df[row.id, "pct_pop"] <- pct.x
-		}
-		print(stats.df)
-	}
-}
-```
-
-## Identify variable genes for new biomark
-```{r}
-length(cmp.object@assays$RNA@var.features)
-```
-
-### Gene profiles of clusters
-set ident at res = 1 and get markers
-```{r}
-Idents(cmp.object) <- "RNA_snn_res.1"
-cmp.allmarkers.res1 <- FindAllMarkers(cmp.object)
-cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
-```
-
-```{r}
-cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
-cmp.top100markers.res1 <- cmp.top100markers.res1[cmp.top100markers.res1$p_val_adj <= 0.05, ]
-for(cluster in unique(cmp.top100markers.res1$cluster)){
-	num.hits <- nrow(cmp.top100markers.res1[cmp.top100markers.res1$cluster == cluster, ])
-	print(paste("cluster", cluster, "has", num.hits, "genes"))
-}
-```
-
-Find top.var	genes in cmp.top100markers.res1
-```{r}
-top.var <- cmp.object@assays$RNA@var.features[1:150]
-# top.var
-n.vargenes <- c()
-for(gene in top.var){
-	if(gene %in% cmp.top100markers.res1$gene){
-		n.vargenes <- c(n.vargenes, gene)
-	}
-}
-print(length(n.vargenes))
-top.var[!(top.var %in% n.vargenes)]
-```
-
-Lets try to export these, with genes on rows and clusters on columns, and xpression value in data table
-```{r}
-var.df <- cmp.top100markers.res1[cmp.top100markers.res1$gene %in% n.vargenes,]
-var.df <- subset.data.frame(var.df, select = c("gene", "cluster", "avg_log2FC"))
-var.df <- reshape2::dcast(var.df, gene~cluster, value.var = "avg_log2FC")
-xlsx::write.xlsx(var.df, file = paste0(projectName, "_dim", ndims, "_candididateBiomarkGenes.xlsx"), sheetName = "res1", append = TRUE, row.names = FALSE)
-
-```
-
-repeat for top.var	genes in cmp.top100markers.res1.5
-```{r}
-top.var <- cmp.object@assays$RNA@var.features[1:150]
-Idents(cmp.object) <- "RNA_snn_res.1.5"
-cmp.allmarkers.res1.5 <- FindAllMarkers(cmp.object)
-cmp.top100markers.res1.5 <- cmp.allmarkers.res1.5 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
-n.vargenes <- c()
-for(gene in top.var){
-	if(gene %in% cmp.top100markers.res1.5$gene){
-		n.vargenes <- c(n.vargenes, gene)
-	}
-}
-print(length(n.vargenes))
-top.var[!(top.var %in% n.vargenes)]
-```
-
-Lets try to export these, with genes on rows and clusters on columns, and xpression value in data table
-```{r}
-var.df <- cmp.top100markers.res1.5[cmp.top100markers.res1.5$gene %in% n.vargenes,]
-var.df <- subset.data.frame(var.df, select = c("gene", "cluster", "avg_log2FC"))
-var.df <- reshape2::dcast(var.df, gene~cluster)
-xlsx::write.xlsx(var.df, file = paste0(projectName, "_dim", ndims, "_candididateBiomarkGenes.xlsx"), sheetName = "res1.5", append = TRUE, row.names = FALSE)
-
-```
-For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
-
-Test: what percentage of each new clusterID matches one of the older clusters?
-```{r}
-for (meta.col in colnames(cmp.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
-		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
-		colnames(enrich.df) <- c(3, 17, 10, 11)
-		rownames(enrich.df) <- new.clusters
-		meta.df <- cmp.object@meta.data
-		for(row.id in rownames(enrich.df)){
-			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
-			for(col.id in colnames(enrich.df)){
-				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
-				pct.x <- as.integer(num.x / tot.clus *100)
-				# print(pct.x)
-				enrich.df[row.id, col.id] <- pct.x
-			}
-		}
-		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
-		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
-		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
-		print(enrich.df)
-	}
-}
-
-```
-Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
-
-## Find old cells on UMAP
-
-time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
-
-```{r fig.width = 4}
-DimPlot(cmp.object,
-				reduction = "umap",
-				group.by = "clust.ID", 
-				# split.by = "orig.ident",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-```{r fig.width = 4}
-DimPlot(cmp.object,
-				reduction = "umap",
-				group.by = "orig.ident", 
-				split.by = "clust.ID",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-
-# Total var 80%
-## Neighborhood and umap
-set total.var <- 80%
-```{r}
-tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 80))
-
-cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
-cmp.object <- FindClusters(cmp.object, resolution = 0.5)
-cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)
-
-```
-Plot UMAP
-
-```{r}
-for(x in c(0.5, 1, 1.5, 2, 2.5)){
-	cmp.object <- FindClusters(cmp.object, resolution = x)
-saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
-}
-```
-
-```{r}
-for (meta.col in colnames(cmp.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		myplot <- DimPlot(cmp.object, 
-											group.by = meta.col,
-											reduction = "umap", 
-											cols = color.palette
-											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
-		plot(myplot)
-	}
-}
-```
-
-### Clustree
-what's the max resolution we can achieve while keepign clusters stable?
-```{r fig.width = 10, fig.height = 5}
-plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
-my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
-	scale_color_continuous(low = 'red3', high = 'white') + 
-	ggtitle(plot.title)
-png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
-plot(my.clustree)
-dev.off()
 ```
@@
 # Total var 95%
-
 ## Neighborhood and umap
@@
 }
-saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
-```
-
-```{r}
-cmp.object <- readRDS("CMP_dim41.RDS")
-```
-
-```{r fig.width = 10, fig.height = 5}
-plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
-my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
-	scale_color_continuous(low = 'red3', high = 'white') + 
-	ggtitle(plot.title)
-png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
-plot(my.clustree)
-dev.off()
 ```
@@
 Create `FindAllMarkers()` lists for GSEA
-```{r }
+```{r}
 Idents(cmp.object) <- "RNA_snn_res.2.5"
@@
 
-## Map HGNC symbols
+Map HGNC symbols
 ```{r}
@@
 for(tested.ident in ident.list){
-	object.copy <- cmp.object
-	Idents(object.copy) <- tested.ident
-	print(paste("testing", tested.ident))
-	for (cluster in sort(as.numeric(levels(object.copy@meta.data[[tested.ident]])))){
-		print(paste("looking at cluster", cluster))
-		cluster.markers <- FindMarkers(object.copy, ident.1 = cluster)
-		try(
+	for (cluster in sort(as.numeric(levels(cmp.object@meta.data[[tested.ident]])))){
+	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
 	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
@@
 									 append = TRUE)
-		)	
 }
-	remove(object.copy)
 }
@@
 
-# Compare @g hierarchcial clusteirng
+# Cell cycle analysis
+Just so we know what we're working with
+
+```{r}
+s.genes <- cc.genes.updated.2019$s.genes
+g2m.genes <- cc.genes.updated.2019$g2m.genes
+```
+
+# Compare @ hierarchcial clusteirng
 

### 2021-09-13 pseudobulk_vs_bulk/pseudobulk_allPops.Rmd
Creating new generate seurat object script and updating some other things

--- 2021-09-13 pseudobulk_vs_bulk/pseudobulk_CMP.Rmd
@@
 
+## Update
+### 2021.09.13
+Finally got a method working for CMP vs.
+TotalSeq CMP (see pseudobulk_CMP.RMD).
+Will now clean up notebook and comparisons for both TotalSeq and ScriptSeq, and for all LSK, CMP, MEP, and GMP populations.
+Following a new [tutorial][https://jdblischak.github.io/singlecell-qtl/sc-vs-bulk.html].
+
 # Notebook setup
-
-Creating new pipeline using seurat v4.0.2 available 2021.06.08
-
-Load libraries required for Seuratv4
 
@@
 library(Seurat)
-library(patchwork)
 library(ggplot2)
-library(SingleCellExperiment)
-library(Matrix.utils)
-library(DESeq2)
-library(edgeR)
-library(csaw)
-library(tximport)
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/pseudobulk_vs_bulk/")
 ```
@@
 ```{r}
-projectName <- "pseudobulk_CMP"
+projectName <- "pseudobulk_allPops"
+comparison.pop <- "CMP"
+export.graphs <- TRUE
 ```
@@
 # source("../RFunctions/PercentVariance.R")
-source("../RFunctions/Mouse2Human_idconversion.R")
+# source("../RFunctions/Mouse2Human_idconversion.R")
 source ("../RFunctions/ColorPalette.R")
@@
 # Load scRNASeq object
-
-## Dim 25
-Using CMP.rds object of dim=25, describing 85% of variation
+##Load raw object
+Has feature and mt filters applied
 ```{r}
-# CMP seurat object generated in "CMPSubset.Rmd"
-seurat.object <- readRDS("CMP_dim25.RDS")
-```
-```{r fig.height=5, fig.width=15}
-p <- FeaturePlot(seurat.object, features = c("Fli1", "Klf1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
-plot(p)
-png(filename = "CMP_Fli1Klf1Ratio.png", width = 2400, height = 400)
-plot(p)
-dev.off()
+seurat.object <- readRDS(paste0("~/Desktop/10XGenomicsData/msAggr_scRNASeq/raw_seuratObjects/", comparison.pop, "_raw.rds"))
 ```
 
-```{r fig.height=5, fig.width=15}
-p <- FeaturePlot(seurat.object, features = c("Gata2", "Gata1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
-plot(p)
-png(filename = "CMP_Gata2Gata1Ratio.png", width = 2400, height = 400)
-plot(p)
-dev.off()
+# Calculate cpm 
+from single cell umi matrix
+```{r}
+sc.mtx <- data.frame(rowSums(seurat.object@assays$RNA@counts))
+sc.mtx <- sc.mtx/colSums(sc.mtx)*1000000
+```
+_Tutorial states that only quantity directly cmoprable between bulk and scRNA-Seq is relative abundance (cites Pachter 2001)._
+
+# Load bRNASeq data
+## read cmp bulk TPM
+```{r}
+bulk.subfolders <- c("../../bRNASeq/TotalSeq/", "../../bRNASeq/ScriptSeq/")
+bulk.path <- bulk.subfolders[1]
+file_list <- list.files(path = bulk.path, pattern = comparison.pop)
+bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]
+
+for(file in file_list){
+	print(file)
+	bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
+}
+
+#calculate average tpm
+bulk.pop[comparison.pop] <- rowMeans(x = bulk.pop[, c(2, 3)])
+
+# clean and check df
+bulk.pop <- bulk.pop[, -c(2, 3)]
+head(bulk.pop)
+```
+```{r}
+#calc log_rho's
+bulk.rho <- log((bulk.pop[comparison.pop] + 1), 2) - log(colSums((bulk.pop[comparison.pop])+ 1), 2)
+rownames(bulk.rho) <- bulk.pop$gene_id
+colnames(bulk.rho) <- "bulk"
+sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
+colnames(sc.rho) <- "sc"
 ```
 
-## Calculate pseudobulk gene expression
-
-__ Based on tutorial @ https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html __
+check that all values are numeric
 ```{r}
-# Extract raw counts and metadata to create SingleCellExperiment object
-
-counts <- seurat.object@assays$RNA@counts
-
-metadata <- seurat.object@meta.data
-
-# Set up metadata as desired for aggregation and DE analysis
-metadata$cluster_id <- factor(seurat.object@active.ident)
-
-# Create single cell experiment object
-sce <- SingleCellExperiment(assays = list(counts = counts), 
-                           colData = metadata)
-
+print(is.numeric(bulk.rho[,1]))
+print(is.numeric(sc.rho[,1]))
+```
+## create concordance df
+```{r}
+# drop NAs
+concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
+concordance.df <- na.omit(concordance.df)
+rownames(concordance.df) <- concordance.df$Row.names
+concordance.df <- concordance.df[, c(2, 3)]
+```
+```{r}
+spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
+spearman.coeff
+```
+## plot data
+```{r}
+cor.plot <- ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
+	geom_point()  +
+	ggtitle(paste0("concordance in ", comparison.pop, " spearman coeff = ", round(spearman.coeff, digits = 3)))
+plot(cor.plot)
 ```
 
-Investigate the object
+## print images
 ```{r}
-print(dim(counts(sce)))
-print(dim(colData(sce)))
-print(head(colData(sce)))
+if(export.graphs == TRUE){
+	png(filename = paste0(comparison.pop, "_spearmanPlot.png"), height = 480, width = 800)
+	plot(cor.plot)
+	dev.off()
+} else{
+	print("Export graphs set to FALSE")
+}
 ```
 
-Note that QC was performed during Seurat analysis
+# Automate through bRNASeq data
+```{r}
+bulk.folders <- c("TotalSeq", "ScriptSeq")
 
-```{r}
-# sce <- perCellQCMetrics(sce)
+for(sequencing.folder in bulk.folders){
+	bulk.subfolders <- c(paste0("~/Desktop/10XGenomicsData/bRNASeq/", sequencing.folder, "/"))
+
+	for(bulk.path in bulk.subfolders){
+		bulk.path <- bulk.subfolders[1]
+		file_list <- list.files(path = bulk.path, pattern = comparison.pop)
+		bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]
+
+		for(file in file_list){
+			print(file)
+			bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
+		}
+
+		#calculate average tpm
+		bulk.pop[comparison.pop] <- rowMeans(x = bulk.pop[, c(2, 3)])
+
+		# clean and check df
+		bulk.pop <- bulk.pop[, -c(2, 3)]
+		head(bulk.pop)
+
+		#calc log_rho's
+		bulk.rho <- log((bulk.pop[comparison.pop] + 1), 2) - log(colSums((bulk.pop[comparison.pop])+ 1), 2)
+		rownames(bulk.rho) <- bulk.pop$gene_id
+		colnames(bulk.rho) <- "bulk"
+		sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
+		colnames(sc.rho) <- "sc"
+
+		# check all entries are numeric
+		print(is.numeric(bulk.rho[,1]))
+		print(is.numeric(sc.rho[,1]))
+
+		# drop NAs
+		concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
+		concordance.df <- na.omit(concordance.df)
+		rownames(concordance.df) <- concordance.df$Row.names
+		concordance.df <- concordance.df[, c(2, 3)]
+
+		# spearman plot
+		spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
+		print(spearman.coeff)
+
+		cor.plot <- ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
+			geom_point()  +
+			ggtitle(paste0("concordance in ", comparison.pop, " vs ", sequencing.folder, ": spearman coeff = ", round(spearman.coeff, digits = 3)))
+		plot(cor.plot)
+
+		if(export.graphs == TRUE){
+			png(filename = paste0(comparison.pop, "_vs", sequencing.folder,"-spearmanPlot.png"), height = 480, width = 800)
+			plot(cor.plot)
+			dev.off()
+		} else{
+			print("Export graphs set to FALSE")
+		}
+	}
+
+}
 ```
-## Aggregate counts
-### Aggregate top-level counts
-
-```{r}
-groups <- colData(sce)[, "orig.ident"]
-aggr.counts <- aggregate.Matrix(t(counts(sce)), groupings = groups, fun = "sum")
-
-dim(aggr.counts)
-```
-
-```{r}
-sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)
-
-```
-
-# Load bRNASeq data
-
-```{r}
-# import rsem.genes.results files
-
-setwd("../../bRNASeq/")
-file_list <- list.files(pattern = "^CMP", recursive = TRUE, include.dirs = TRUE)
-names(file_list) <- lapply(file_list, function(x) gsub('.genes.results', '', x))
-file_list
-```
-
-```{r}
-mstxi <- tximport(files = paste0("../../bRNASeq/", file_list), type = 'rsem', txIn = FALSE, txOut = FALSE, geneIdCol = 3, )
-colnames(mstxi$counts) <- gsub("^.*/", "", names(file_list))
-```
-
-```{r}
-rna.counts <- merge(mstxi$counts, t(aggr.counts), by = "row.names", all = TRUE)
-rna.counts[is.na(rna.counts)]<-0
-rownames(rna.counts) <- rna.counts$Row.names
-rna.counts <- rna.counts[, 2:ncol(rna.counts)]
-head(rna.counts)
-```
-
-```{r}
-dgelist_groups <- factor(c(unlist(stringr::str_extract_all(gsub("/.*", "", names(file_list)), pattern = "[a-zA-Z]+")), "SingleCell"))
-levels(dgelist_groups)
-```
-
-```{r}
-dgelist <- DGEList(counts = rna.counts, group = dgelist_groups)
-dgelist
-apply(dgelist$counts, 2, sum)
-keep <- rowSums(cpm(dgelist) > 10) >=2
-dgelist <- dgelist[keep,]
-```
-
-### Calculate library size
-```{r}
-dgelist$samples$lib.size <- colSums(dgelist$counts)
-dgelist <- calcNormFactors(dgelist)
-```
-
-### Calculate dispersion
-
-designMat
-```{r}
-designMat <- model.matrix(~ 0 + dgelist$samples$group)
-colnames(designMat) <- levels(dgelist$samples$group)
-designMat
-```
-
-estimate GLM dispersion and apply
-```{r}
-dgelist <- estimateGLMCommonDisp(dgelist, designMat)
-dgelist <- estimateGLMTrendedDisp(dgelist, designMat, method = 'bin.spline')
-dgelist <- estimateGLMTagwiseDisp(dgelist, designMat)
-```
-
-```{r}
-fit <- glmQLFit(dgelist, designMat, robust = TRUE)
-plotQLDisp(fit)
-```
-
-```{r}
-qlf <- glmQLFTest(fit, coef = )
-summary(decideTests(qlf))
-```
-
-#### rld on  individual samples
-
-```{r}
-rld <- rlog(round(dgelist$counts), blind = TRUE)
-rld <- as.data.frame(rld)
-rld.spear <- cor(rld, method = "spearman")
-rld.melt <- reshape2::melt(rld)
-colnames(rld.melt) <- c("sample", "rld")
-```
-
-```{r}
-rld.spear
-```
-
-```{r}
-ggplot(data = rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
-```
-
-```{r}
-ggplot(data = rld, aes(x = CMPm2, y = CMP1010)) + geom_point()
-```
-
-#### rld on pooled samples
-
-```{r}
-pool.counts <- dgelist$counts
-pool.counts <- as.data.frame(pool.counts)
-pool.counts$ScriptSeq <- (pool.counts$CMP1009 + pool.counts$CMP1010)/2
-pool.counts$TruSeq <- (pool.counts$CMP443 + pool.counts$CMP448)/2
-pool.counts <- subset(pool.counts, select = c(ScriptSeq, TruSeq, CMPm2))
-pool.counts <- as.matrix(pool.counts)
-```
-
-```{r}
-pool.rld <- rlog(round((pool.counts)), blind = TRUE)
-pool.rld.spear <- cor(pool.rld, method = "spearman")
-pool.rld.melt <- reshape2::melt(pool.rld)
-colnames(pool.rld.melt) <- c("sample", "rld")
-```
-
-```{r}
-pool.rld.spear
-```
-
-```{r}
-ggplot(data = pool.rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
-```
-
-```{r}
-ggplot(data = pool.rld.melt, aes(x = CMPm2, y = ScriptSeq)) + geom_point()
-```
-
-Combine replicates from CPM table
-
-```{r}
-rna.cpm <- as.data.frame(cpm(dgelist, log = TRUE, normalized.lib.sizes = TRUE))
-rna.cpm$ScriptSeq <- (rna.cpm$CMP1009 + rna.cpm$CMP1010)/2
-rna.cpm$TruSeq <- (rna.cpm$CMP443 + rna.cpm$CMP448)/2
-rna.cpm <- subset(rna.cpm, select = c(ScriptSeq, TruSeq, CMPm2))
-```
-
-plot correlation and scatter grams
-```{r}
-cpm.spear <- cor(rna.cpm, method = "spearman")
-cpm.melt <- reshape2::melt(rld)
-colnames(cpm.melt) <- c("sample", "rld")
-```
-
-```{r}
-cpm.spear
-```
-
-## Create DESeq object from single cell data
-```{r}
-sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)
-```

### 2021-09-13 raw_seuratObjects/GenerateRawSeuratObjects.Rmd
Creating new generate seurat object script and updating some other things

---
title: "Create Raw Seurat Objects"
output: html_notebook
---

# Purpose
Purpose of this notebook is simply to generate raw seurat objects for analysis in other programs

# Notebook setup

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup	}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/cellRanger/")
```

# Set global variables
```{r}
projectName <- "GenerateRawSeuratObject"
```

store session info
```{r }
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
```

```{r warning=FALSE}
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)

# Clean up to free memory

remove(object.data)

# Add mitochondrial metadata and plot some basic features

seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )

plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2

# remove low quality cells
# require: nFeature_RNA between 200 and 4000 (inclusive)
# require: percent.mt <=5

print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000 & 
													percent.mt <= 5
												)
print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))

# Save raw object

saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```

### 2021-09-14 pseudobulk_vs_bulk/pseudobulk_allPops.Rmd
start automating pseudobulk analysis

@@
 
+```{r setup}
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/pseudobulk_vs_bulk/")
+```
+
 ## Load libraries
-```{r setup}
+```{r}
 library(dplyr)
@@
 library(ggplot2)
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/pseudobulk_vs_bulk/")
 ```

### 2021-09-14 RFunctions/read_10XGenomics_data.R

@@ -1,5 +1,5 @@
 read_10XGenomics_data<-function(sample.list){
-	data <- sapply(sample.list, FUN = function(x) {x = dirname(list.files(path = x, pattern = 'barcodes.tsv', full.names = TRUE, recursive = TRUE, no.. = TRUE))})
+	data <- sapply(sample.list, FUN = function(x) {x = dirname(list.files(path = x, pattern = '*barcode*', full.names = TRUE, recursive = TRUE, no.. = TRUE))})
 	names(data) <- sample.list
 	return(data)
 }

### 2021-09-15 pseudobulk_vs_bulk/pseudobulk_automate.R
add pseudobulk_automate.R

@@ -1,0 +1,119 @@
+## Load libraries
+library(dplyr)
+library(Seurat)
+library(ggplot2)
+
+## Set global variables
+projectName <- "pseudobulk_allPops"
+cell.list <- c("LSKm2", "CMPm2", "CMPm", "MEPm", "GMPm")
+bulk.list <- c("LSK", "CMP", "CFUE", "CFUMK", "ERY", "GMP", "MK", "MEP")
+export.graphs <- TRUE
+all.comparisons <- TRUE
+
+## Store session info
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
+
+## Load local scripts
+source ("../RFunctions/ColorPalette.R")
+
+spearman.scriptseq <- data.frame(matrix(NA, nrow = length(bulk.list), ncol = length(cell.list)))
+rownames(spearman.scriptseq) <- bulk.list
+colnames(spearman.scriptseq) <- cell.list
+head(spearman.scriptseq)
+
+spearman.totalseq <- spearman.scriptseq
+
+
+for (cell.pop in cell.list){
+	# Load scRNASeq object
+	seurat.object <- readRDS(paste0("~/Desktop/10XGenomicsData/cellRanger/", cell.pop, "_raw.RDS"))
+	# Calculate cpm 
+	sc.mtx <- data.frame(rowSums(seurat.object@assays$RNA@counts))
+	sc.mtx <- sc.mtx/colSums(sc.mtx)*1000000
+	
+	if(all.comparisons == TRUE){
+		bulk.pops <- bulk.list
+	}else{
+		bulk.pops <- substr(cell.pop, 1, 3)
+	}
+	print(bulk.pops)	
+	for(comparison.pop in bulk.pops){
+		# Automate through bRNASeq data
+		bulk.folders <- c("TotalSeq", "ScriptSeq")
+		
+		for(sequencing.folder in bulk.folders){
+			bulk.subfolders <- c(paste0("~/Desktop/10XGenomicsData/bRNASeq/", sequencing.folder, "/"))
+			
+			for(bulk.path in bulk.subfolders){
+				bulk.path <- bulk.subfolders[1]
+				file_list <- list.files(path = bulk.path, pattern = comparison.pop)
+				bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]
+				
+				for(file in file_list){
[…]

### 2021-09-16 msAggr/msAggr_seurat.Rmd
completing code for pseudobulk analysis

@@
 # library(clustree)
-```
-store session info
-```{r}
-sink("msAggr_seurat-v1.20210623")
-sessionInfo()
-sink()
 ```
@@
 jackstraw.dim <- 40
+```
+
+## Store session info
+```{r}
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
 ```

### 2021-09-16 RFunctions/ColorPalette.R

@@ -1,0 +1,44 @@
+color.palette <- c(
+  "coral",
+  "chartreuse4",
+  "goldenrod1",
+  "cadetblue1",
+  "burlywood",
+  "brown",
+  "brown1",
+  "blue",
+  "blue4",
+  "azure3",
+  "aquamarine",
+  "antiquewhite",
+  "cadetblue",
+  "gold3",
+  "black",
+  "darkgreen",
+  "deeppink",
+  "darkviolet",
+  "darkturquoise",
+  "darkslategray",
+  "darksalmon",
+  "darkorchid1",
+  "darkolivegreen2",
+  "forestgreen",
+  "dodgerblue",
+  "green",
+  "lightpink",
+  "lightcoral",
+  "khaki1",
+  "maroon",
+  "peru",
+  "lightseagreen",
+  "lightsalmon",
+  "plum",
+  "moccasin",
+  "tan",
+  "tan1", 
+  "red", 
+  "purple",
+  "khaki4",
+  "black", 
+  "plum4"
+)

### 2021-09-16 pseudobulk_vs_bulk/pseudobulk_automate.R

@@ -2,12 +2,18 @@
 library(dplyr)
 library(Seurat)
 library(ggplot2)
+library(xlsx)
 
 ## Set global variables
 projectName <- "pseudobulk_allPops"
 cell.list <- c("LSKm2", "CMPm2", "CMPm", "MEPm", "GMPm")
 bulk.list <- c("LSK", "CMP", "CFUE", "CFUMK", "ERY", "GMP", "MK", "MEP")
-export.graphs <- TRUE
+bulk.folders <- c("TotalSeq", "ScriptSeq")
+count.progress <- 0
+
+## Set function options
+plot.graphs <- FALSE
+export.graphs <- FALSE
 all.comparisons <- TRUE
 
 ## Store session info
@@ -16,16 +22,18 @@ sink(sessionInfo.filename)
 sessionInfo()
 sink()
 
-## Load local scripts
-source ("../RFunctions/ColorPalette.R")
-
+## Create empty spearman coefficient dataframes
 spearman.scriptseq <- data.frame(matrix(NA, nrow = length(bulk.list), ncol = length(cell.list)))
 rownames(spearman.scriptseq) <- bulk.list
 colnames(spearman.scriptseq) <- cell.list
-head(spearman.scriptseq)
-
 spearman.totalseq <- spearman.scriptseq
 
+if(plot.graphs == FALSE){
+	print("Export graphs set to FALSE")
+}
+if(export.graphs == FALSE){
+	print("Export graphs set to FALSE")
+}
 
 for (cell.pop in cell.list){
 	# Load scRNASeq object
@@ -39,10 +47,8 @@ for (cell.pop in cell.list){
 	}else{
 		bulk.pops <- substr(cell.pop, 1, 3)
 	}
-	print(bulk.pops)	
 	for(comparison.pop in bulk.pops){
 		# Automate through bRNASeq data
-		bulk.folders <- c("TotalSeq", "ScriptSeq")
 		
 		for(sequencing.folder in bulk.folders){
 			bulk.subfolders <- c(paste0("~/Desktop/10XGenomicsData/bRNASeq/", sequencing.folder, "/"))
@@ -87,33 +93,43 @@ for (cell.pop in cell.list){
 				
 				# spearman plot
 				spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
+				# spearman.coeff <- round(spearman.coeff, 3)
 				if(sequencing.folder == "TotalSeq"){
[…]

### 2021-09-17 msAggr_seurat/msAggr_seurat.Rmd
complete non-SCTransform msAggr analysis to ID optimal cluster dim and resolution

--- 2021-09-16 msAggr/msAggr_seurat.Rmd
@@
 ```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/msAggr_seurat/")
 library(dplyr)
@@
 ```{r}
-projectName <- "msAggr"
+projectName <- "msAggr_seurat"
 jackstraw.dim <- 40
@@
 ```{r}
-source("msAggr_AnalysisCode/read_10XGenomics_data.R")
-source("msAggr_AnalysisCode/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
 ```
 
-```{r}
-setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
+```{r warning=FALSE}
+setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
 data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
@@
 ```{r}
-save.image(file = paste0(projectName, '.RData'))
+# save.image(file = paste0(projectName, '.RData'))
+saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
 ```
@@
 tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
-paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
-paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
-paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
+paste0("Num pcs for 80% variance: ", length(which(cumsum(tot.var) <= 80)))
+paste0("Num pcs for 85% variance: ", length(which(cumsum(tot.var) <= 85)))
+paste0("Num pcs for 90% variance: ", length(which(cumsum(tot.var) <= 90)))
+paste0("Num pcs for 95% variance: ", length(which(cumsum(tot.var) <= 95)))
 
@@
 
-## Add cluster IDs from Seurat v1
+Percent of PCs describing X% of variance (transcribed from above cell because I don't know how to freeze results)
 
-Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
-Create column "clust.ID" and populate with 0's. Then import IDs for clusters
-```{r}
-clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
-clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
-clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
-clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
-clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-```
+Num pcs for 80% variance: 12
+Num pcs for 85% variance: 18
+Num pcs for 90% variance: 26
+Num pcs for 95% variance: 37
 
-Add new metadata column and map new ids
-```{r}
-seurat.object@meta.data['clust.ID'] <- 0
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
-```
+# ID clusters based on different variances
+## Describe 80% of variance with 12 dimensions
 
-do numbers make sense (we don't expect the count to b exactly the same as the numbers in the original cluster)?
-```{r}
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 10,])
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 11,])
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 17,])
-nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
-```
-
-make enough sense!
-
-# DGE
-
-Let's do some cluster analyses and see if we can find these populations in our new analysis.
-## Ideal resolution...? (Is this a thing?)
-### Color palette
-```{r}
-color.palette <- c(
-	"coral",
-	"chartreuse4",
-	"goldenrod1",
-	"cadetblue1",
-	"burlywood",
-	"brown",
-	"brown1",
-	"blue",
-	"blue4",
-	"azure3",
-	"aquamarine",
-	"antiquewhite",
-	"cadetblue",
-	"gold3",
-	"black",
-	"darkgreen",
-	"deeppink",
-	"darkviolet",
-	"darkturquoise",
-	"darkslategray",
-	"darksalmon",
-	"darkorchid1",
-	"darkolivegreen2",
-	"forestgreen",
-	"dodgerblue",
-	"green",
-	"lightpink",
-	"lightcoral",
-	"khaki1",
-	"maroon",
-	"peru",
-	"lightseagreen",
-	"lightsalmon",
-	"plum",
-	"moccasin",
-	"tan",
-	"tan1", 
-	"red", 
-	"purple",
-	"khaki4",
-	"black", 
-	"plum4"
-)
-```
-
-# Total var 90%
-## Neighborhood and umap
-set total.var <- 90%
+### Neighborhood and umap
+set total.var <- 80%
 ```{r}
 tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 90))
-print(ndims)
+ndims <- length(which(cumsum(tot.var) <= 80))
 
@@
 
-# saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 }
+```
+
+```{r}
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
 ```
@@
 
-For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
-
-Test: what percentage of each new clusterID matches one of the older clusters?
-```{r}
-for (meta.col in colnames(seurat.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
-		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
-		colnames(enrich.df) <- c(3, 17, 10, 11)
-		rownames(enrich.df) <- new.clusters
-		meta.df <- seurat.object@meta.data
-		for(row.id in rownames(enrich.df)){
-			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
-			for(col.id in colnames(enrich.df)){
-				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
-				pct.x <- as.integer(num.x / tot.clus *100)
-				# print(pct.x)
-				enrich.df[row.id, col.id] <- pct.x
-			}
-		}
-		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
-		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
-		try{xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)}
-		print(enrich.df)
-	}
-}
-
-```
-Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
-
-## Find old cells on UMAP
-
-time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
-
-```{r fig.width = 4}
-DimPlot(seurat.object,
-				reduction = "umap",
-				group.by = "clust.ID", 
-				# split.by = "orig.ident",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-```{r fig.width = 4}
-DimPlot(seurat.object,
-				reduction = "umap",
-				group.by = "orig.ident", 
-				split.by = "clust.ID",
-				cols = c("gray", "orange", "blue", "red", "green"),)
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
 ```
 
-# Total var 95%
-## Neighborhood and umap
-set total.var <- 95%
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+## Describe 85% of variance with 18 dimensions
+
+### Neighborhood and umap
+set total.var <- 85%
 ```{r}
 tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 95))
-print(ndims)
+ndims <- length(which(cumsum(tot.var) <= 85))
 
@@
 
-For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?
-
-Test: what percentage of each new clusterID matches one of the older clusters?
 ```{r}
@@
 	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
-		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
-		colnames(enrich.df) <- c(3, 17, 10, 11)
-		rownames(enrich.df) <- new.clusters
-		meta.df <- seurat.object@meta.data
-		for(row.id in rownames(enrich.df)){
-			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
-			for(col.id in colnames(enrich.df)){
-				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
-				pct.x <- as.integer(num.x / tot.clus *100)
-				# print(pct.x)
-				enrich.df[row.id, col.id] <- pct.x
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
 	}
 }
-		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
-		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
-		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
-		print(enrich.df)
-	}
+```
+
+```{r}
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+
+Generate statistics for each cluster/resolution combo
+```{r}
+current_res <- 'RNA_snn_res.0.5'
+
+cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
+for(id in cluster_ids){
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "LSKm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "CMPm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
+	counts_df[id, "MEPm"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
+	counts_df[id, "GMPm"] = cell_value
+
 }
@@
 ```
-Absolutely terrible overlap, no enrichment of any of these across the new clustering algorithm. Maybe should try 95% variation covered
 
-## Find old cells on UMAP
-
-time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4
-
-```{r fig.width = 2}
-DimPlot(seurat.object,
-				reduction = "umap",
-				group.by = "clust.ID", 
-				pt.size = .1,
-				# split.by = "orig.ident",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-```{r fig.width = 4}
-DimPlot(seurat.object,
-				reduction = "umap",
-				group.by = "orig.ident", 
-				split.by = "clust.ID",
-				cols = c("gray", "orange", "blue", "red", "green"),)
-```
-
-### Gene expression of old clustrs on new map
-Let's see if we can get some gene expression profiles on these...
-```{r, fig.height=10, fig.width=18}
-gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
-VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
-```
-
-## Evaluate cluster stability
+### Evaluate cluster stability
 Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
@@
 Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
-
-### Clustree
 Previously my favourite has been Clustree, which gives a nice visual
@@
 
-These data suggest that node stability is aweful! Need to figure out if this is a dimensional reduction error or a clustering error.
+## Describe 90% of variance with 26 dimensions
 
-# Why are clusters so unstable?
-Differences could include:
+### Neighborhood and umap
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 90))
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+```{r}
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+
+Generate statistics for each cluster/resolution combo
+```{r}
+current_res <- 'RNA_snn_res.0.5'
+
+cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
+for(id in cluster_ids){
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "LSKm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "CMPm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
+	counts_df[id, "MEPm"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
+	counts_df[id, "GMPm"] = cell_value
+
+}
+
+```
+
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_dim", ndims, "-clustree.png"), height = 800, width = 1600)
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
+```
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+# Describe 95% of variance with 37 dimensions
+
+### Neighborhood and umap
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 95))
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											cols = color.palette
+											) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+		plot(myplot)
+	}
+}
+```
+
+```{r}
+saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
+```
+
+Generate statistics for each cluster/resolution combo
+```{r}
+current_res <- 'RNA_snn_res.0.5'
+
+cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
+for(id in cluster_ids){
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "LSKm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "CMPm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
+	counts_df[id, "MEPm"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
+	counts_df[id, "GMPm"] = cell_value
+
+}
+
+```
+
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+# DGE
+I think clusters from 90% variance at 0.5 and 1.0 resolution are most stable.
+We'll do some statistics and DGE on that.
+Will also need to go back and play with SCTransform, since these are multiple cell types from multiple lanes.
+
+```{r}
+seurat.obje
+```
+
+# Notes on cluster stability
+Cluster stability could be influenced by:
 * cells in each population (cellranger v6 includes more cells than cellranger v1, especially in MEP)
@@
 ```
-Save dim36 as is and try clustering analysis @ dim24
-```{r}
-saveRDS(seurat.object, file = "msAggr_AnalysisCode/msAggr_dim36.rds")
-```
-
-# Repeat clustering with dim24
-One possibility is that I've included too many dimensions. Will see if 90% increases stability.
-
-```{r}
-seurat.object <- FindNeighbors(seurat.object, dims = 1:24)
-seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-seurat.object <- RunUMAP(seurat.object, dims = 1:24)
-```
-
-Save object
-```{r}
-saveRDS(seurat.object, file = "msAggr_dim24.rds")
-```
-
-```{r}
-for (meta.col in colnames(seurat.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		myplot <- DimPlot(seurat.object, 
-											group.by = meta.col,
-											reduction = "umap", 
-											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
-											) + 
-			ggtitle(paste0("msAggr dim36 res", gsub("RNA_snn_res", "", meta.col) ))
-		plot(myplot)
-	}
-}
-```
-
-## Evaluate cluster stability
-Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
-
-Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
-
-### Clustree
-Previously my favourite has been Clustree, which gives a nice visual
-NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-Think I'll explore regression factors using SCTransform in new document.

### 2021-09-17 RFunctions/ColorPalette.R

@@ -1,44 +1,50 @@
 color.palette <- c(
-  "coral",
-  "chartreuse4",
-  "goldenrod1",
-  "cadetblue1",
-  "burlywood",
+  "darkolivegreen2",
+  "green",
+  "forestgreen",
+  "darkgreen",
+  "darkslategray",
+  "tan",
+  "black",
+  "indianred1",
+  "maroon",
+  "salmon4",
   "brown",
-  "brown1",
+  "firebrick3",
+  "red", 
+  "aquamarine",
+  "cadetblue1",
+  "dodgerblue",
   "blue",
   "blue4",
+  "slategray2",
   "azure3",
-  "aquamarine",
+  "dimgrey", 
   "antiquewhite",
-  "cadetblue",
-  "gold3",
-  "black",
-  "darkgreen",
-  "deeppink",
-  "darkviolet",
-  "darkturquoise",
-  "darkslategray",
-  "darksalmon",
-  "darkorchid1",
-  "darkolivegreen2",
-  "forestgreen",
-  "dodgerblue",
-  "green",
-  "lightpink",
-  "lightcoral",
-  "khaki1",
-  "maroon",
-  "peru",
-  "lightseagreen",
-  "lightsalmon",
-  "plum",
   "moccasin",
-  "tan",
-  "tan1", 
-  "red", 
-  "purple",
+  "khaki1",
+  "cadetblue",
+  "lightseagreen",
+  "darkturquoise",
   "khaki4",
-  "black", 
-  "plum4"
+  "deeppink",
+  "magenta",
+  "darkorchid1",
+  "mediumpurple1",
+  "plum",
+  "plum4",
+  "darkviolet",
+  "lightpink",
+  "darksalmon",
+  "lightcoral",
+  "coral",
+  "darkorange",
+  "tan1", 
+  "lightsalmon",
+  "goldenrod1",
+  "peru",
+  "gold3",
+  "black" 
 )
+
+# show_col(color.palette)

### 2021-09-23 CMP/CMPSubset.Rmd
Adding individual population work and continuing aggregated population work

@@
 
-store session info
-```{r }
-# sink("CMP-v1.20210616.txt")
+## Store session info
+```{r}
+sink(paste0(projectName, "_sessionInfo.txt"))
 sessionInfo()
-# sink()
+sink()
 ```
@@
 Yay!!
+
+# Export for Biomark clustering
+
+```{r}
+biomark.genes <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/MergeBiomarkWith10X/geneIDs.txt", header = FALSE)
+biomark.genes <- unlist(biomark.genes)
+```
+```{r}
+seurat.xprsn <- seurat.object@assays$RNA@scale.data
+seurat.xprsn <- seurat.xprsn[row.names(seurat.xprsn) %in% biomark.genes,]
+seurat.xprsn <- merge(seurat.xprsn, seurat.object@meta.data$RNA_snn_res.1, by = rownames(seurat.xprsn))
+```

### 2021-09-23 IndividualPops/CMPSubpop.Rmd
Adding individual population work and continuing aggregated population work

--- 2021-09-17 msAggr_seurat/msAggr_seurat.Rmd
@@
 ---
-title: "msAggr_seurat"
-output: html_notebook
+title: "CMP_seurat"
+output: 
+  html_notebook: 
+    fig_height: 6
 ---
@@
 
-Load libraries required for Seuratv4
-
+# Setup
+## Load libraries required for Seuratv4
 ```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/msAggr_seurat/")
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/")
 library(dplyr)
@@
 library(clustree)
+```
+## Source functions
+```{r}
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
+```
+
+## Set global variables
+
+```{r}
+projectName <- "CMPSubpop"
+jackstraw.dim <- 40
+```
+
+## Store session info
+```{r}
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
 ```
@@
 
-## Set global variables
-
-```{r}
-projectName <- "msAggr_seurat"
-jackstraw.dim <- 40
-```
-
-## Store session info
-```{r}
-sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
-sink(sessionInfo.filename)
-sessionInfo()
-sink()
-```
-
-```{r}
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
-```
-
 ```{r warning=FALSE}
 setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
-data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
+data_file.list <- read_10XGenomics_data(sample.list = c("CMPm2"))
 data.object<-Read10X(data_file.list)
@@
 
-Generate statistics for each cluster/resolution combo
-```{r}
-current_res <- 'RNA_snn_res.0.5'
-
-cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
-counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
-rownames(counts_df) <- cluster_ids
-colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
-for(id in cluster_ids){
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "LSKm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "CMPm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
-	counts_df[id, "MEPm"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
-	counts_df[id, "GMPm"] = cell_value
-
-}
-
-```
-
 ### Evaluate cluster stability
@@
 
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-## Describe 85% of variance with 18 dimensions
+## Describe 85% of variance
 
@@
 
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-## Describe 90% of variance with 26 dimensions
+## Describe 90% of variance
 
@@
 for (meta.col in colnames(seurat.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
+	if(grepl(pattern = "RNA_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
 		myplot <- DimPlot(seurat.object, 
@@
 											reduction = "umap", 
-											cols = color.palette
-											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col) ))
 		plot(myplot)
+		png(filename = paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
 	}
@@
 saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
-```
-
-Generate statistics for each cluster/resolution combo
-```{r}
-current_res <- 'RNA_snn_res.0.5'
-
-cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
-counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
-rownames(counts_df) <- cluster_ids
-colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
-for(id in cluster_ids){
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "LSKm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "CMPm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
-	counts_df[id, "MEPm"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
-	counts_df[id, "GMPm"] = cell_value
-
-}
-
 ```
@@
 ```
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-# Describe 95% of variance with 37 dimensions
+## Describe 95% of variance
 
@@
 
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
+# Export for Biomark
+## Import biomark probe list
+```{r}
+seurat.object <- readRDS("CMPSubpop_dim18.RDS")
 ```
 
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
+```{r}
+probe.list <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/BiomarkRNAs.txt", header = FALSE)
+probe.list <- unlist(as.list(probe.list), use.names = FALSE)
 ```
 
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
+```{r}
+xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))) %in% probe.list, ])
+head(xprsn.mtx[1:5, 1:5])
+```
+```{r}
+biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
+```
+```{r}
+write.table(biomark.export, file = paste0(projectName, "_ScaledData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
 ```
@@
 # DGE
-I think clusters from 90% variance at 0.5 and 1.0 resolution are most stable.
+I think clusters from 80% variance at 0.5 and 1.0 resolution are most stable.
 We'll do some statistics and DGE on that.
 Will also need to go back and play with SCTransform, since these are multiple cell types from multiple lanes.
+## Load favourite dim reduction file
+```{r}
+rds.file <- ""
+seurat.object <- readRDS(rds.file)
+ndims <- as.numeric(gsub("[^0-9]", "", stringr::str_split(rds.file, "_")[[1]][3]))
+```
+
+## Set favourite resolution
+```{r}
+object.res <- ".0.5"
+Idents(seurat.object) <- paste0("RNA_snn_res", object.res)
+length(levels(seurat.object@active.ident))
+```
+
+## count number of filtered cells left from each population
+```{r}
+# Number of filtered cells left in each pop
+sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
+```
 
 ```{r}
-seurat.obje
+par(mfrow = c(2, 2))
+for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
+	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
+	h$density = h$counts/sum(h$counts)*100
+	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
+}
+par(mfrow = c(1,1))
+```
+
+```{r fig.width=5}
+VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
+				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
+```
+
+## DGE w/ resolution = 0.5
+Strt with comparing all clusters against all other clusters and write out cluster info
+calculate `FindAllMarkers()` for different idents and save to new file
+```{r}
+# ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+
+for(tested.ident in ident.list){
+	Idents(seurat.object) <- tested.ident
+	all.markers <- FindAllMarkers(seurat.object)
+	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+									 file = paste0(projectName, "_FindALLMarkers_dim",ndims, "_allres.xlsx"), 
+									 sheetName = tested.ident, 
+									 col.names = TRUE, 
+									 row.names = FALSE, 
+									 append = TRUE)
+}
+```
+
+## Create `FindAllMarkers()` lists for GSEA
+```{r}
+object.res.allmarkers <- FindAllMarkers(seurat.object)
+```
+
+## Map HGNC symbols
+```{r}
+Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
+
+HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
+head(object.res.allmarkers)
+object.res.allmarkers$HGNC <- HGNC
+tail(object.res.allmarkers)
+sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
+sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
+sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
+sig.res <- sig.res[]
+
+```
+
+```{r}
+for(cluster in unique(sig.res$cluster)){
+	print(paste("writing cluster", cluster))
+	new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
+	new.table <- new.table[order(-new.table$avg_log2FC), ]
+	dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE)
+	write.table(new.table, file = paste0("RankList_res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+
+}
+```
+
+calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
+```{r}
+# ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+for(tested.ident in ident.list){
+	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
+		cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+		xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+										 file = paste0(projectName, "_FindMarkers_dim", ndims, gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
+										 sheetName = paste0("clst", cluster), 
+										 col.names = TRUE, 
+										 row.names = TRUE, 
+										 append = TRUE)
+	}
+}
+```
+
+```{r}
+for (cluster in sort(as.numeric(levels(seurat.object@meta.data[paste0("RNA_snn_res", object.res)])))){
+	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+									 file = paste0(projectName, "_FindMarkers_dim", ndims, "res", object.res, ".xlsx"), 
+									 sheetName = paste0("clst", cluster), 
+									 col.names = TRUE, 
+									 row.names = TRUE, 
+									 append = TRUE)
+}
 ```
@@
 * SCTransform accounts better for sources of variability
-
-```{r}
-# Number of filtered cells left in each pop
-sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
-```
-
-```{r}
-for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
-	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
-	h$density = h$counts/sum(h$counts)*100
-	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
-}
-```
-
-Looks like MEPm is the only sample with that huge MitoC % lump @ 40%. What do these cells look like, otherwise?
-
-```{r fig.width=5}
-VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
-				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
-```

### 2021-09-23 IndividualPops/LSKSubpop.Rmd
Adding individual population work and continuing aggregated population work

--- 2021-09-23 IndividualPops/CMPSubpop.Rmd
@@
 ---
-title: "CMP_seurat"
+title: "msAggr_seurat"
 output: 
@@
 
-## Set global variables
-
-```{r}
-projectName <- "CMPSubpop"
-jackstraw.dim <- 40
-```
-
-## Store session info
-```{r}
-sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
-sink(sessionInfo.filename)
-sessionInfo()
-sink()
-```
-
 # A note about using SCTransform versus `ScaleData`
@@
 >In addition, the normalized values in SCT and integrated assays don't necessary correspond to per-gene expression values anyway, rather containing residuals (in the case of the scale.data slot for each).
+## Set global variables
 
-Mess with how to load 4 cell populations into single seurat object
+```{r}
+projectName <- "LSKSubpop"
+jackstraw.dim <- 40
+```
 
-SET SEED?????!!!!!
+## Store session info
+```{r}
+sink(paste0(projectName, "_sessionInfo.txt"))
+sessionInfo()
+sink()
+```
 
@@
 setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
-data_file.list <- read_10XGenomics_data(sample.list = c("CMPm2"))
+data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2"))
 data.object<-Read10X(data_file.list)
@@
 
-## Describe 85% of variance
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+
+## Describe 85% of variance with 18 dimensions
 
@@
 
+Generate statistics for each cluster/resolution combo
+```{r}
+current_res <- 'RNA_snn_res.0.5'
+
+cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
+for(id in cluster_ids){
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "LSKm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "CMPm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
+	counts_df[id, "MEPm"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
+	counts_df[id, "GMPm"] = cell_value
+
+}
+
+```
+
 ### Evaluate cluster stability
@@
 
-## Describe 90% of variance
+## Describe 90% of variance with 26 dimensions
 
@@
 
+Generate statistics for each cluster/resolution combo
+```{r}
+current_res <- 'RNA_snn_res.1'
+
+cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 9))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm", "TotInClust", "PctClusLSK", "PctClusCMP", "PctClusMEP", "PctClusGMP")
+for(id in cluster_ids){
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+	counts_df[id, "LSKm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "CMPm2"),])
+	counts_df[id, "CMPm2"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
+	counts_df[id, "MEPm"] = cell_value
+	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
+																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
+	counts_df[id, "GMPm"] = cell_value
+
+}
+counts_df["TotInClust"] <- rowSums(counts_df, na.rm = TRUE)
+counts_df$PctClusLSK <- round(counts_df$LSKm2/counts_df$TotInClust, 3)*100
+counts_df$PctClusCMP <- round(counts_df$CMPm2/counts_df$TotInClust, 3)*100
+counts_df$PctClusMEP <- round(counts_df$MEPm/counts_df$TotInClust, 3)*100
+counts_df$PctClusGMP <- round(counts_df$GMPm/counts_df$TotInClust, 3)*100
+xlsx::write.xlsx(x = counts_df, 
+								 file = paste0(projectName, "CellCountPerClust",ndims, "res.1.xlsx"), 
+								 sheetName = "res1", 
+								 col.names = TRUE, 
+								 row.names = TRUE, 
+								 append = TRUE)
+
+counts_df
+
+```
+
 ### Evaluate cluster stability
@@
 ```
-## Describe 95% of variance
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+```
+## Describe 95% of variance with 37 dimensions
 
@@
 
-# Export for Biomark
-## Import biomark probe list
-```{r}
-seurat.object <- readRDS("CMPSubpop_dim18.RDS")
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
 ```
 
-```{r}
-probe.list <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/BiomarkRNAs.txt", header = FALSE)
-probe.list <- unlist(as.list(probe.list), use.names = FALSE)
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
 ```
 
-```{r}
-xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))) %in% probe.list, ])
-head(xprsn.mtx[1:5, 1:5])
-```
-```{r}
-biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
-```
-```{r}
-write.table(biomark.export, file = paste0(projectName, "_ScaledData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
 ```
@@
 # DGE
-I think clusters from 80% variance at 0.5 and 1.0 resolution are most stable.
+I think clusters from 90% variance at 0.5 and 1.0 resolution are most stable.
 We'll do some statistics and DGE on that.
@@
 ```{r}
-rds.file <- ""
+rds.file <- "msAggr_seurat_dim26.RDS"
 seurat.object <- readRDS(rds.file)

### 2021-09-23 msAggr_seurat/msAggr_seurat.Rmd
Adding individual population work and continuing aggregated population work

@@
 title: "msAggr_seurat"
-output: html_notebook
+output: 
+  html_notebook: 
+    fig_height: 6
 ---
@@
 
-Load libraries required for Seuratv4
-
+# Setup
+## Load libraries required for Seuratv4
 ```{r setup, message=FALSE, warning=FALSE}
@@
 library(clustree)
+```
+## Source functions
+```{r}
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
 ```
@@
 sink()
-```
-
-```{r}
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
 ```
@@
 ```{r}
-current_res <- 'RNA_snn_res.0.5'
+current_res <- 'RNA_snn_res.1'
 
 cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
-counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 9))
 rownames(counts_df) <- cluster_ids
-colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
+colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm", "TotInClust", "PctClusLSK", "PctClusCMP", "PctClusMEP", "PctClusGMP")
 for(id in cluster_ids){
@@
 	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
+																						 	(seurat.object@meta.data$orig.ident == "CMPm2"),])
 	counts_df[id, "CMPm2"] = cell_value
@@
 }
+counts_df["TotInClust"] <- rowSums(counts_df, na.rm = TRUE)
+counts_df$PctClusLSK <- round(counts_df$LSKm2/counts_df$TotInClust, 3)*100
+counts_df$PctClusCMP <- round(counts_df$CMPm2/counts_df$TotInClust, 3)*100
+counts_df$PctClusMEP <- round(counts_df$MEPm/counts_df$TotInClust, 3)*100
+counts_df$PctClusGMP <- round(counts_df$GMPm/counts_df$TotInClust, 3)*100
+xlsx::write.xlsx(x = counts_df, 
+								 file = paste0(projectName, "CellCountPerClust",ndims, "res.1.xlsx"), 
+								 sheetName = "res1", 
+								 col.names = TRUE, 
+								 row.names = TRUE, 
+								 append = TRUE)
+
+counts_df
 
@@
 ```
-# Describe 95% of variance with 37 dimensions
+## Describe 95% of variance with 37 dimensions
 
@@
 Will also need to go back and play with SCTransform, since these are multiple cell types from multiple lanes.
+## Load favourite dim reduction file
+```{r}
+rds.file <- "msAggr_seurat_dim26.RDS"
+seurat.object <- readRDS(rds.file)
+ndims <- as.numeric(gsub("[^0-9]", "", stringr::str_split(rds.file, "_")[[1]][3]))
+```
+
+## Set favourite resolution
+```{r}
+object.res <- ".0.5"
+Idents(seurat.object) <- paste0("RNA_snn_res", object.res)
+length(levels(seurat.object@active.ident))
+```
+
+## count number of filtered cells left from each population
+```{r}
+# Number of filtered cells left in each pop
+sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
+```
 
 ```{r}
-seurat.obje
+par(mfrow = c(2, 2))
+for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
+	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
+	h$density = h$counts/sum(h$counts)*100
+	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
+}
+par(mfrow = c(1,1))
+```
+
+```{r fig.width=5}
+VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
+				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
+```
+
+## DGE w/ resolution = 0.5
+Strt with comparing all clusters against all other clusters and write out cluster info
+calculate `FindAllMarkers()` for different idents and save to new file
+```{r}
+# ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+
+for(tested.ident in ident.list){
+	Idents(seurat.object) <- tested.ident
+	all.markers <- FindAllMarkers(seurat.object)
+	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+									 file = paste0(projectName, "_FindALLMarkers_dim",ndims, "_allres.xlsx"), 
+									 sheetName = tested.ident, 
+									 col.names = TRUE, 
+									 row.names = FALSE, 
+									 append = TRUE)
+}
+```
+
+## Create `FindAllMarkers()` lists for GSEA
+```{r}
+object.res.allmarkers <- FindAllMarkers(seurat.object)
+```
+
+## Map HGNC symbols
+```{r}
+Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
+
+HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
+head(object.res.allmarkers)
+object.res.allmarkers$HGNC <- HGNC
+tail(object.res.allmarkers)
+sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
+sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
+sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
+sig.res <- sig.res[]
+
+```
+
+```{r}
+for(cluster in unique(sig.res$cluster)){
+	print(paste("writing cluster", cluster))
+	new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
+	new.table <- new.table[order(-new.table$avg_log2FC), ]
+	dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE)
+	write.table(new.table, file = paste0("RankList_res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+
+}
+```
+
+calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
+```{r}
+# ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+for(tested.ident in ident.list){
+	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
+		cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+		xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+										 file = paste0(projectName, "_FindMarkers_dim", ndims, gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
+										 sheetName = paste0("clst", cluster), 
+										 col.names = TRUE, 
+										 row.names = TRUE, 
+										 append = TRUE)
+	}
+}
+```
+
+```{r}
+for (cluster in sort(as.numeric(levels(seurat.object@meta.data[paste0("RNA_snn_res", object.res)])))){
+	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+									 file = paste0(projectName, "_FindMarkers_dim", ndims, "res", object.res, ".xlsx"), 
+									 sheetName = paste0("clst", cluster), 
+									 col.names = TRUE, 
+									 row.names = TRUE, 
+									 append = TRUE)
+}
 ```

### 2021-09-23 RFunctions/ColorPalette.R

@@ -1,50 +1,50 @@
 color.palette <- c(
-  "darkolivegreen2",
-  "green",
-  "forestgreen",
-  "darkgreen",
-  "darkslategray",
-  "tan",
-  "black",
-  "indianred1",
-  "maroon",
-  "salmon4",
-  "brown",
-  "firebrick3",
   "red", 
-  "aquamarine",
-  "cadetblue1",
-  "dodgerblue",
   "blue",
-  "blue4",
-  "slategray2",
+  "tan1", 
+  "green",
   "azure3",
-  "dimgrey", 
-  "antiquewhite",
-  "moccasin",
-  "khaki1",
-  "cadetblue",
-  "lightseagreen",
-  "darkturquoise",
-  "khaki4",
   "deeppink",
   "magenta",
-  "darkorchid1",
-  "mediumpurple1",
-  "plum",
-  "plum4",
   "darkviolet",
-  "lightpink",
-  "darksalmon",
-  "lightcoral",
-  "coral",
-  "darkorange",
-  "tan1", 
-  "lightsalmon",
-  "goldenrod1",
+  "tan",
+  "aquamarine",
+  "black",
+  "maroon",
+  "cadetblue",
   "peru",
+  "lightpink",
+  "plum4",
+  "blue4",
   "gold3",
+  "mediumpurple1",
+  "khaki1",
+  "firebrick3",
+  "dodgerblue",
+  "forestgreen",
+  "indianred1",
+  "darkorange",
+  "dimgrey", 
+  "antiquewhite",
+  "coral",
+  "moccasin",
+  "lightseagreen",
+  "darkturquoise",
+  "darkgreen",
+  "darkolivegreen2",
+  "lightcoral",
+  "khaki4",
+  "darkorchid1",
+  "plum",
+  "cadetblue1",
+  "darksalmon",
+  "lightsalmon",
+  "darkslategray",
+  "salmon4",
+  "brown",
+  "goldenrod1",
+  "slategray2",
   "black" 
 )
 
-# show_col(color.palette)
+# scales::show_col(color.palette)

### 2021-09-23 IndividualPops/CMPSubpop.Rmd
creating single script to streamline individual pop analyses

@@
 output:
-  html_notebook: 
-    fig_height: 6
+  html_document:
+    df_print: paged
 ---
-
 Creating new pipeline using seurat v4.0.2 available 2021.06.23  
@@
 ```{r}
-write.table(biomark.export, file = paste0(projectName, "_ScaledData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+write.table(biomark.export, file = paste0(projectName, "_Counts-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
 ```

### 2021-09-23 IndividualPops/Individual_SubpopAnalyses.Rmd
creating single script to streamline individual pop analyses

--- 2021-09-17 msAggr_seurat/msAggr_seurat.Rmd
@@
 ---
-title: "msAggr_seurat"
-output: html_notebook
+title: "Individual_Pop_Analysis"
+output:
+  html_document:
+    df_print: paged
 ---
-
 Creating new pipeline using seurat v4.0.2 available 2021.06.23  
@@
 
-Load libraries required for Seuratv4
-
+# Setup
+## Load libraries required for Seuratv4
 ```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/msAggr_seurat/")
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/")
+knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
 library(dplyr)
@@
 library(clustree)
+```
+## Source functions
+```{r}
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
+```
+
+## Set global variables
+```{r}
+pop.id <- "GMP"
+folder.appendix <- "m"
+fav.var <- 80
+object.res <- ".1"
+ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+```
+
+```{r}
+folder.id <- paste0(pop.id, folder.appendix)
+projectName <- paste0(pop.id, "Subpop")
+```
+
+## Store session info
+```{r}
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
 ```
@@
 
-## Set global variables
-
-```{r}
-projectName <- "msAggr_seurat"
-jackstraw.dim <- 40
-```
-
-## Store session info
-```{r}
-sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
-sink(sessionInfo.filename)
-sessionInfo()
-sink()
-```
-
-```{r}
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
-```
-
-```{r warning=FALSE}
+```{r warning=FALSE, message=FALSE}
 setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
-data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
+data_file.list <- read_10XGenomics_data(sample.list = c(folder.id))
 data.object<-Read10X(data_file.list)
@@
 
-```{r fig.width=4, fig.height=2}
+```{r fig.width = 6, fig.height=4}
 plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
@@
 Find variable features
-```{r fig.width = 5, fig.height = 2}
+```{r fig.width = 12, fig.height = 5}
 seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
@@
 plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
-plot1 + plot2
+try(plot1 + plot2)
 ```
@@
 
-Percent of PCs describing X% of variance (transcribed from above cell because I don't know how to freeze results)
-
-Num pcs for 80% variance: 12
-Num pcs for 85% variance: 18
-Num pcs for 90% variance: 26
-Num pcs for 95% variance: 37
-
 # ID clusters based on different variances
@@
 tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 80))
+pct.var <- 80
+ndims <- length(which(cumsum(tot.var) <= pct.var))
+print(ndims)
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+
+```{r}
+for (meta.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = "RNA_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col) ))
+		plot(myplot)
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
+	}
+}
+```
+
+```{r}
+saveRDS(seurat.object, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
+```
+
+```{r}
+
+```
+
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
+```
+
+## Describe 85% of variance
+
+### Neighborhood and umap
+set total.var <- 85%
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+pct.var <- 85
+ndims <- length(which(cumsum(tot.var) <= pct.var))
 
@@
 											reduction = "umap", 
-											cols = color.palette
-											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col) ))
 		plot(myplot)
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(seurat.object, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
 	}
@@
 ```{r}
-saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
-```
-
-Generate statistics for each cluster/resolution combo
-```{r}
-current_res <- 'RNA_snn_res.0.5'
-
-cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
-counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
-rownames(counts_df) <- cluster_ids
-colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
-for(id in cluster_ids){
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "LSKm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "CMPm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
-	counts_df[id, "MEPm"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
-	counts_df[id, "GMPm"] = cell_value
-
-}
-
+saveRDS(seurat.object, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
 ```
@@
 ### Evaluate cluster stability
+Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
+
+Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
+Previously my favourite has been Clustree, which gives a nice visual
+NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
+
 ```{r fig.height=5}
@@
 	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
 ```
 
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-## Describe 85% of variance with 18 dimensions
+## Describe 90% of variance
 
 ### Neighborhood and umap
-set total.var <- 85%
 ```{r}
 tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 85))
+pct.var <- 90
+ndims <- length(which(cumsum(tot.var) <= pct.var))
 
@@
 ### Evaluate cluster stability
-Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.
-
-Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.
-Previously my favourite has been Clustree, which gives a nice visual
-NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.
-
 ```{r fig.height=5}
@@
 	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
 ```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-## Describe 90% of variance with 26 dimensions
+## Describe 95% of variance
 
@@
 tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 90))
+pct.var <- 95
+ndims <- length(which(cumsum(tot.var) <= pct.var))
 
@@
 	scale_color_continuous(low = 'red3', high = 'white')
-png(filename = paste0(projectName, "_dim", ndims, "-clustree.png"), height = 800, width = 1600)
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
 clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") +
@@
 ```
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
+
+# Export for Biomark
+## Import biomark probe list
+```{r}
+seurat.object <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
 ```
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-# Describe 95% of variance with 37 dimensions
-
-### Neighborhood and umap
-```{r}
-tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-ndims <- length(which(cumsum(tot.var) <= 95))
-
-seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
-seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
-
-```
-Plot UMAP
 
 ```{r}
-for(x in c(0.5, 1, 1.5, 2, 2.5)){
-	seurat.object <- FindClusters(seurat.object, resolution = x)
+probe.list <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/BiomarkRNAs.txt", header = FALSE)
+probe.list <- unlist(as.list(probe.list), use.names = FALSE)
+```
+
+```{r}
+xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))) %in% probe.list, ])
+head(xprsn.mtx[1:5, 1:5])
+```
+```{r}
+biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
+```
+```{r}
+write.table(biomark.export, file = paste0(projectName, "_Counts-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+```
+
+# DGE
+I think clusters from 80% variance at 0.5 and 1.0 resolution are most stable.
+We'll do some statistics and DGE on that.
+Will also need to go back and play with SCTransform, since these are multiple cell types from multiple lanes.
+## Load favourite dim reduction file
+```{r}
+seurat.object <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
+```
+
+## Set favourite resolution
+```{r}
+
+Idents(seurat.object) <- paste0("RNA_snn_res", object.res)
+length(levels(seurat.object@active.ident))
+```
+
+## count number of filtered cells left from each population
+```{r}
+# Number of filtered cells left in each pop
+# sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
+```
+
+```{r}
+# par(mfrow = c(2, 2))
+# for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
+# 	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
+# 	h$density = h$counts/sum(h$counts)*100
+# 	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
+# }
+# par(mfrow = c(1,1))
+```
+
+```{r fig.width=5}
+# VlnPlot(subset(seurat.object, subset = orig.ident == "MEPm"), 
+# 				features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 1, pt.size = 0, fill.by = 'ident', flip = TRUE)
+```
+
+## DGE w/ object
+Strt with comparing all clusters against all other clusters and write out cluster info
+calculate `FindAllMarkers()` for different idents and save to new file
+```{r}
+# ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+
+for(tested.ident in ident.list){
+	Idents(seurat.object) <- tested.ident
+	all.markers <- FindAllMarkers(seurat.object)
+	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+									 file = paste0(projectName, "_FindALLMarkers_pctVar", pct.var, "_allres.xlsx"), 
+									 sheetName = tested.ident, 
+									 col.names = TRUE, 
+									 row.names = FALSE, 
+									 append = TRUE)
 }
@@
 
+## Create `FindAllMarkers()` lists for GSEA
 ```{r}
-for (meta.col in colnames(seurat.object@meta.data)){
-	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
-		myplot <- DimPlot(seurat.object, 
-											group.by = meta.col,
-											reduction = "umap", 
-											cols = color.palette
-											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
-		plot(myplot)
+object.res.allmarkers <- FindAllMarkers(seurat.object)
+```
+
+## Map HGNC symbols
+```{r}
+Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
+
+HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
+head(object.res.allmarkers)
+object.res.allmarkers$HGNC <- HGNC
+tail(object.res.allmarkers)
+sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
+sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
+sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
+sig.res <- sig.res[]
+
+```
+
+```{r}
+for(cluster in unique(sig.res$cluster)){
+	print(paste("writing cluster", cluster))
+	new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
+	new.table <- new.table[order(-new.table$avg_log2FC), ]
+	dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE)
+	try(write.table(new.table, file = paste0(projectName, "RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
+
+}
+```
+
+calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
+```{r}
+# ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+for(tested.ident in ident.list){
+	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
+		cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+		xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+										 file = paste0(projectName, "_FindMarkers_pctVar", pct.var, gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
+										 sheetName = paste0("clst", cluster), 
+										 col.names = TRUE, 
+										 row.names = TRUE, 
+										 append = TRUE)
 	}
@@
 ```{r}
-saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
-```
-
-Generate statistics for each cluster/resolution combo
-```{r}
-current_res <- 'RNA_snn_res.0.5'
-
-cluster_ids <- sort(unique(seurat.object@meta.data[,current_res]))
-counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
-rownames(counts_df) <- cluster_ids
-colnames(counts_df) <- c("LSKm2", "CMPm2", "MEPm", "GMPm")
-for(id in cluster_ids){
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "LSKm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "LSKm2"),])
-	counts_df[id, "CMPm2"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "MEPm"),])
-	counts_df[id, "MEPm"] = cell_value
-	cell_value <- nrow(seurat.object@meta.data[(seurat.object@meta.data[current_res] == id) & 
-																						 	(seurat.object@meta.data$orig.ident == "GMPm"),])
-	counts_df[id, "GMPm"] = cell_value
-
+for (cluster in sort(as.numeric(levels(seurat.object@meta.data[paste0("RNA_snn_res", object.res)])))){
+	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
+	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
+									 file = paste0(projectName, "_FindMarkers_pctVar", pct.var, "res", object.res, ".xlsx"), 
+									 sheetName = paste0("clst", cluster), 
+									 col.names = TRUE, 
+									 row.names = TRUE, 
+									 append = TRUE)
 }
-
-```
-
-### Evaluate cluster stability
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'scale.data', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-```{r fig.height=5}
-clustree(seurat.object, prefix = "RNA_snn_res.", expres = 'counts', node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-```
-
-# DGE
-I think clusters from 90% variance at 0.5 and 1.0 resolution are most stable.
-We'll do some statistics and DGE on that.
-Will also need to go back and play with SCTransform, since these are multiple cell types from multiple lanes.
-
-```{r}
-seurat.obje
 ```

### 2021-09-24 IndividualPops/CMPSubpop.Rmd
streamlining individualPops analyses

@@
 
-## DGE w/ resolution = 0.5
+## DGE
 Strt with comparing all clusters against all other clusters and write out cluster info

### 2021-09-24 IndividualPops/Individual_SubpopAnalyses.Rmd
streamlining individualPops analyses

@@
 ```{r}
-pop.id <- "GMP"
-folder.appendix <- "m"
+pop.id <- "CMP"
+folder.appendix <- "m2"
 fav.var <- 80
@@
 xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))) %in% probe.list, ])
-head(xprsn.mtx[1:5, 1:5])
-```
-```{r}
 biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
-```
-```{r}
 write.table(biomark.export, file = paste0(projectName, "_Counts-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+
+xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@data))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@data))) %in% probe.list, ])
+biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
+write.table(biomark.export, file = paste0(projectName, "_Data-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+
+xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@scale.data))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@scale.data))) %in% probe.list, ])
+biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
+write.table(biomark.export, file = paste0(projectName, "_ScaleData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+
 ```

### 2021-10-01 IndividualPops/CMPSubpop.Rmd
added code to export dataframe from seurat assay slots

@@
 ```{r}
+setdiff(probe.list, rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))))
+setdiff(c("Kit", "Il3ra"), rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))))
+
+```
+
+```{r}
 xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))) %in% probe.list, ])
+colnames(xprsn.mtx)[colnames(xprsn.mtx) == "Cd117"] <- "Kit"
+colnames(xprsn.mtx)[colnames(xprsn.mtx) == "Cd123"] <- "Il3ra"
 head(xprsn.mtx[1:5, 1:5])

### 2021-10-01 IndividualPops/Individual_SubpopAnalyses.Rmd
added code to export dataframe from seurat assay slots

--- 2021-09-23 IndividualPops/Individual_SubpopAnalyses.Rmd
@@
 ```{r}
-pop.id <- "GMP"
-folder.appendix <- "m"
-fav.var <- 80
+pop.id <- "CMP"
+folder.appendix <- "m2"
+fav.var <- 90
 object.res <- ".1"
 ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
-```
-
-```{r}
 folder.id <- paste0(pop.id, folder.appendix)
@@
 
+# Hierarchcial clustering
+```{r}
+seurat.object <- SetIdent(seurat.object, value = "RNA_snn_res.0.5")
+seurat.object <- BuildClusterTree(seurat.object)
+PlotClusterTree(seurat.object, direction = "downwards")
+```
+
 # Export for Biomark
@@
 seurat.object <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
+seurat.object <- SetIdent(seurat.object, value = paste0("RNA_snn_res", object.res))
 ```
@@
 probe.list <- unlist(as.list(probe.list), use.names = FALSE)
+probe.list <- c(probe.list, "Kit", "Il3ra")
+```
+## downsample
+__NOTE: do I need to do the downsampling before the corrections? Probably not, because I'm exporting counts and not corrected data
+```{r}
+object.downsample <- subset(biomark.object, cells = sample(Cells(biomark.object), 300))
 ```
 
+## generate and export counts matrix
 ```{r}
-xprsn.mtx <- t(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))[rownames(as.data.frame(as.matrix(seurat.object@assays$RNA@counts))) %in% probe.list, ])
-head(xprsn.mtx[1:5, 1:5])
-```
-```{r}
-biomark.export <- merge(xprsn.mtx, seurat.object@meta.data["RNA_snn_res.1"], by = 0)
-```
-```{r}
+xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "counts"))))
+object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
+# biomark.export <- xprsn.mtx
+biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
+colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
+colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
 write.table(biomark.export, file = paste0(projectName, "_Counts-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+
+xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "data"))))
+object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
+# biomark.export <- xprsn.mtx
+biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
+colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
+colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
+write.table(biomark.export, file = paste0(projectName, "_Data-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+
+xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "scale.data"))))
+object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
+# biomark.export <- xprsn.mtx
+biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
+colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
+colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
+write.table(biomark.export, file = paste0(projectName, "_ScaleData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
+
 ```

### 2021-10-05 IndividualPops/Individual_SubpopAnalyses.Rmd
changes made to individual subpop analysis on main branch by mistake

@@
 
-# Hierarchcial clustering
-```{r}
-seurat.object <- SetIdent(seurat.object, value = "RNA_snn_res.0.5")
-seurat.object <- BuildClusterTree(seurat.object)
-PlotClusterTree(seurat.object, direction = "downwards")
-```
-
 # Export for Biomark
@@
 
+```
+
+## Heatmap cluster
+```{r}
+seurat.object <- SetIdent(seurat.object, value = "RNA_snn_res.0.5")
+seurat.markers <- FindAllMarkers(seurat.object)
+```
+```{r}
+# backup.seurat.markers <- seurat.markers
+seurat.markers <- backup.seurat.markers
+```
+
+```{r warning=FALSE}
+# seurat.markers.dplyr <- seurat.markers %>% group_by(cluster) %>% arrange(desc(avg_log2FC))
+# biomark.intersect <- seurat.markers[rownames(seurat.markers) %in% probe.list, ]
+# biomark.intersect <- biomark.intersect[biomark.intersect$p_val_adj < 0.05, ]
+# biomark.intersect <- biomark.intersect[order(biomark.intersect$cluster, -biomark.intersect$avg_log2FC),]
+```
+```{r}
+# seurat.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) -> biomark.intersect
+biomark.intersect <- seurat.markers[seurat.markers$gene %in% probe.list,]
+biomark.intersect %>% group_by(cluster) %>% arrange(desc(avg_log2FC)) -> biomark.intersect
+DoHeatmap(seurat.object, features = biomark.intersect$gene)
+png(filename = paste0(projectName, "_pctVar", fav.var, "res.0.5.png"), height = 1600, width = 1600)
+DoHeatmap(seurat.object, features = biomark.intersect$gene)
+dev.off()
+```
+
+```{r}
+biomark.order <- c("Ifitm1", "Lgals3", "Txnip", "Ube2c", "Cst3", "Lmo4", "Psap", "S100a8", "Fos", "Rbp2", "Kit", "Klf6", "AA467197", "Cd34","Ly6c2", "Hist1h2ap", "Ly6a", "Pbx1", "Apoe", "Cd44", "Emb", "Actb", "Csrp3", "Lsp1", "Ms4a2", "Cpa3", "Gp1bb", "Mt2" ,"Ccl3", "Cebpe", "H2-Aa", "B2m", "Mpl", "Ctla2a", "Cd69", "Cd53", "Flt3", "Egr1", "Nrgn", "Aif1", "Jun", "Cd9", "Irf8", "Hp", "Prtn3", "Vamp5", "Cd27", "Birc5", "Mpo", "Vwf", "Cnrip1", "Itga2b", "Cd74", "Hlf", "Ms4a6c", "Kng2", "Jund", "Lyz2", "Il3ra", "Cd48", "Cd82", "Ccl9", "Gata1", "Tsc22d1", "Ly86", "Mecom", "S100a11", "Itgb3", "Gata2", "Sell", "Dhrs3", "Cd52", "Pf4", "Tsmb4x", "Klf1", "Tmod1", "Cd177", "Dntt")
+setdiff(biomark.order, seurat.markers$gene)
+```
+
+```{r}
+biomark.intersect <- seurat.markers[seurat.markers$gene %in% biomark.order,]
+# biomark.intersect$gene <- factor(biomark.intersect$gene, levels = biomark.order)
+biomark.intersect %>% group_by(cluster) %>% arrange(factor(biomark.intersect$gene, levels = biomark.order)) -> biomark.intersect
+biomark.heatmap <- DoHeatmap(seurat.object, features = biomark.intersect$gene)
+```
+```{r}
+png(filename = paste0(projectName, "_pctVar", fav.var, "res.0.5-biomarkorder.png"), height = 1600, width = 1600)
+plot(biomark.heatmap)
+dev.off()
+```
+
+## Hierarchcial clustering
+```{r}
+# seurat.object <- SetIdent(seurat.object, value = "RNA_snn_res.0.5")
+seurat.object <- BuildClusterTree(seurat.object)
+PlotClusterTree(seurat.object, direction = "downwards")
 ```

### 2021-10-05 MergeBiomarkWith10X/MergeBiomark167With10X.Rmd
initiate Bioimark branch

---
title: "Mergebiomark167With10X"
output:
  html_document:
    df_print: paged
---

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/MergeBiomarkWith10X/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
```

# Set global variables
```{r}
projectName <- "Mergebiomark167With10X"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")
source("msAggr_AnalysisCode/PercentVariance.R")
source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
```

store session info
```{r }
sink(paste0(projectName, ".txt"))
sessionInfo()
sink()
```

```{r}
color.palette <- c(
	"coral",
	"chartreuse4",
	"goldenrod1",
	"cadetblue1",
	"burlywood",
	"brown",
	"brown1",
	"blue",
	"blue4",
	"azure3",
	"aquamarine",
	"antiquewhite",
	"cadetblue",
	"gold3",
	"black",
	"darkgreen",
	"deeppink",
	"darkviolet",
	"darkturquoise",
	"darkslategray",
	"darksalmon",
	"darkorchid1",
	"darkolivegreen2",
	"forestgreen",
	"dodgerblue",
	"green",
	"lightpink",
	"lightcoral",
	"khaki1",
	"maroon",
	"peru",
	"lightseagreen",
	"lightsalmon",
	"plum",
	"moccasin",
	"tan",
	"tan1", 
	"red", 
	"purple",
	"khaki4",
	"black", 
	"plum4"
)
```

# Create seurat object

## Load biomark data
Try to read in Biomark Data

Will need to include:
cell list
probe list
expression matrix

```{r}
biomark.data <- read.table("biomark167_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0
biomark.data.T <- data.table::transpose(biomark.data)
colnames(biomark.data.T) <- rownames(biomark.data)
rownames(biomark.data.T) <- colnames(biomark.data)
# fix non-MGI gene names
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
```

Pull in original cellID info

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r}
# move fluorophor markers to metadata column, remove for expression dataset
fluor.ids <- colnames(biomark.data[grep("pos", colnames(biomark.data))])
fluor.table <- biomark.data[fluor.ids]
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) %in% fluor.ids),]
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
colnames(biomark.meta) <- c("wid", "pid")

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}

# Match ids from original hierarchical clustering assignments to cells retained in Biomark167
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

# match normalized fluorescene 
i1 <- match(row.names(biomark.meta), row.names(fluor.table))
biomark.meta[names(fluor.table)[1:ncol(fluor.table)]] <- lapply(fluor.table[1:ncol(fluor.table)], `[`, i1)

```

```{r}
# N.B., can't find alternate name for Kng2, so remove
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
biomark.data.T <- as.sparse(biomark.data.T)
biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

## Load SC data

```{r warning=FALSE}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
```

Remove non-biomark genes from cmp.object

```{r}
biomark.genes <- rownames(biomark@assays$RNA@data)
cmp.object <- subset(cmp.object, features = biomark.genes)
cmp.object@meta.data["wid"] <- '10X'
cmp.object@meta.data["pid"] <- "p14"
setdiff(rownames(biomark@assays$RNA@data), rownames(cmp.object@assays$RNA@data))
```

Clean up to free memory

```{r}
remove(object.data, biomark.data, biomark.meta)
```

## merge objects
```{r}
seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
seurat.object@meta.data$orig.ident <- as.factor(seurat.object@meta.data$orig.ident)
seurat.object@meta.data$pid <- as.factor(seurat.object@meta.data$pid)
seurat.object@meta.data$wid <- as.factor(seurat.object@meta.data$wid)
seurat.object
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
```

Scale data (linear transformation)

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("orig.ident", "nFeature_RNA", "pid", "nCount_RNA"))
```

### Save merged object
```{r}
# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r}
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

```{r}
clust3.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

## map cluster ids from Seuratv1
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
seurat.object@meta.data$clust.ID <- as.factor(seurat.object@meta.data$clust.ID)
```

## Lets see which cells correspond to our old clusters

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r fig.width=15, fig.height=5}
DimPlot(seurat.object, split.by = "clust.ID", group.by = "orig.ident", pt.size = 2)
```

## Split object for anchor-based integration

```{r}
cmp.subset <- SplitObject(seurat.object, split.by = "orig.ident")$`10X`
DimPlot(cmp.subset, group.by = "clust.ID")
```

## Follow integration protocol

https://satijalab.org/seurat/archive/v3.0/integration.html

```{r}
tech.list <- SplitObject(seurat.object, split.by = "orig.ident")

for(i in 1:length(tech.list)){
	tech.list[[i]] <- NormalizeData(tech.list[[i]], verbose = FALSE)
	tech.list[[i]] <- FindVariableFeatures(tech.list[[i]], selection.method = "vst", nfeatures = 10, verbose = FALSE)
}
```

## identify anchors
```{r}
# reference.list <- tech.list["10X"]
tech.anchors <- FindIntegrationAnchors(object.list = tech.list, dims = 1:30, anchor.features = all.genes, k.anchor = 20, reduction = 'rpca', k.filter = NA)
# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 5)
```
## tech.integrated PCA

```{r}
DefaultAssay(tech.integrated) <- "integrated"
tech.integrated <- ScaleData(tech.integrated)
tech.integrated <- RunPCA(tech.integrated, npcs = 30)
DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident")
```

so it looks like anchor-based integration is much nicer! goign to stick with this approach.

Problem now is, how do we find whether Biomark overlays on 10X

```{r fig.height = 5, fig.width=15}
DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "pca")
```

```{r fig.height=10, fig.width=10}
png(filename = "tech.integrated_biomark167Fluor.png", height = 1600, width = 1600)
FeaturePlot(seurat.object, features = c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos"), split.by = "clust.ID", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 1)
dev.off()

```

## tech.integrated umap
```{r}
tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
ndims <- length(which(cumsum(tot.var) <= 90))
```

```{r}
tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims, reduction = "pca")
tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims, reduction = "pca")
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	tech.integrated <- FindClusters(tech.integrated, resolution = x)
}
```

```{r}
for (meta.col in colnames(tech.integrated@meta.data)){
	if(grepl(pattern = ("integrated_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											cols = color.palette
											) + 
			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("integrated_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

Switch ident
```{r}
Idents(tech.integrated) <- "integrated_snn_res.0.5"
```

```{r fig.width = 15, fig.height=20}
DimPlot(tech.integrated, group.by = "clust.ID", pt.size = 2, split.by = "ident", ncol = 4, cols = color.palette, shape.by = "orig.ident") + ggmin::theme_powerpoint()
```

```{r fig.width=20}
DimPlot(tech.integrated, group.by = "orig.ident", pt.size = 2, ncol = 2)
```

```{r fig.width=10, fig.height=20}
DimPlot(tech.integrated, group.by = "orig.ident", pt.size = 2, ncol = 1, split.by = "clust.ID", cols = c("red", "black")) + ggmin::theme_powerpoint()
```

```{r}
png("tech.integrated_biomark167UMAP.res.0.5.png", width = 1600, height = 1600)
DimPlot(tech.integrated, split.by = "ident", group.by = "orig.ident", pt.size = 2, ncol = 5, ) + ggmin::theme_powerpoint()
dev.off()
```

```{r}
png(filename = "tech.integrated_biomark167FeatureFluor.png", height = 1600, width =1600)
FeaturePlot(tech.integrated, features = c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos"), reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 3, pt.size = 2)
dev.off()
```

```{r}
for(i in c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos")){
	my_title <- paste0("tech.integrated_biomark167Fluor_", i, ".png")
	print(my_title)
png(filename = my_title, height = 1000, width = 6200)
my_plot <- FeaturePlot(tech.integrated, features = i, split.by = "ident", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 5, pt.size = 3)
plot(my_plot)
dev.off()
}
```

```{r}
# for(i in unique(tech.integrated@meta.data$clust.ID)){
	# my_title <- paste0("tech.integrated_biomark167clustID_", i, ".png")
	print(my_title)
png(filename = "tech.integrated_biomark167clustID.png", height = 1000, width = 6200)
my_plot <- DimPlot(tech.integrated, pt.size = 6, group.by = "clust.ID", split.by = "ident", shape.by = "orig.ident") + ggmin::theme_powerpoint()
plot(my_plot)
dev.off()
# }
```

# Biomark Only
## Load biomark data

```{r}
biomark.data <- read.table("biomark167_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0
biomark.data.T <- data.table::transpose(biomark.data)
colnames(biomark.data.T) <- rownames(biomark.data)
rownames(biomark.data.T) <- colnames(biomark.data)
# fix non-MGI gene names
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
```

Pull in original cellID info

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
```

```{r}
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
colnames(biomark.meta) <- c("wid", "pid")

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

```

```{r}
# N.B., can't find alternate name for Kng2, so remove
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
biomark.data.T <- as.sparse(biomark.data.T)
biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

```{r}
biomark <- NormalizeData(biomark, normalization.method = "LogNormalize", scale.factor = 10000)
biomark <- FindVariableFeatures(biomark, selection.method = "vst", nfeatures = 10)
all.genes <- rownames(biomark)
biomark <- ScaleData(biomark, features = all.genes, vars.to.regress = c("nFeature_RNA", "pid", "nCounts_RNA"))
```

### Save merged object
```{r}
# saveRDS(biomark, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
biomark <- RunPCA(biomark, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(biomark, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(biomark, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r}
ElbowPlot(biomark, ndims = 50)
percent.variance(biomark@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(biomark@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

## Lets see which cells correspond to our old clusters

```{r}
DimPlot(biomark, reduction = "pca", group.by = "orig.ident") + facet_wrap(~orig.ident)
```

```{r fig.width=15, fig.height=8}
biomark@meta.data$clust.ID <- as.factor(biomark@meta.data$clust.ID)
DimPlot(biomark, group.by = "clust.ID", pt.size = 2, shape.by = "clust.ID")
```

```{r fig.width=15, fig.height = 5}
DimPlot(biomark, split.by = "clust.ID", pt.size = 2, shape.by = "orig.ident")
```

# Setup
## Load libraries required for Seuratv4
```{r message=FALSE, warning=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/")
knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
```

## Set global variables
```{r}
pop.id <- "CMP"
folder.appendix <- "m2"
fav.var <- 90
object.res <- ".0.5"
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
folder.id <- paste0(pop.id, folder.appendix)
projectName <- paste0(pop.id, "Subpop_vBiomark")
```

## Store session info
```{r}
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

*STOP* - testing showed that cells cluster better in seaborn when you include all probe data points. It was an intersting idea to cluster based only on the genes in the biomark probe set, but this doesn't seem to hold up in practice

```{r warning=FALSE, message=FALSE}
setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = c(folder.id))
data.object<-Read10X(data_file.list)
seurat.object<- CreateSeuratObject(counts = data.object, min.cells = 3, min.genes = 200, project = projectName)
remove(data.object)
probe.list <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/BiomarkRNAs.txt", header = FALSE)
probe.list <- unlist(as.list(probe.list), use.names = FALSE)
probe.list <- c(probe.list, "Kit", "Il3ra")
# Subset to biomark features
biomark.object <- subset(seurat.object, features = probe.list)
biomark.object[["percent.mt"]] <- PercentageFeatureSet(biomark.object, pattern = "^mt-")
print(paste("original object:", nrow(biomark.object@meta.data), "cells", sep = " "))
biomark.object <- subset(biomark.object, 
												subset = nFeature_RNA >=25 & 
													percent.mt <= 5
)
print(paste("new object:", nrow(biomark.object@meta.data), "cells", sep = " "))
biomark.object <- NormalizeData(biomark.object, normalization.method = "LogNormalize", scale.factor = 10000)
object.downsample <- subset(biomark.object, cells = sample(Cells(biomark.object), 300))
```

## overlay data from clustered seurat.object onto biomark.object
```{r}
seurat.object <- readRDS("~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/CMPSubpop_pctVar90.RDS")
xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "counts"))))
object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
# biomark.export <- xprsn.mtx
biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
write.table(biomark.export, file = paste0(projectName, "_Counts-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)

xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "data"))))
object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
# biomark.export <- xprsn.mtx
biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
write.table(biomark.export, file = paste0(projectName, "_Data-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)

xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "scale.data"))))
object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
# biomark.export <- xprsn.mtx
biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
write.table(biomark.export, file = paste0(projectName, "_ScaleData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)

```

```{r}
biomark.object <- readRDS("~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/CMPSubpop_pctVar90.RDS")
```

### 2021-10-05 MergeBiomarkWith10X/MergeBiomark167With10X.Rmd
Show that anchor integration is the best method for biomark/10x integration

@@
 
+# Updates
+
+## 20211005
+Compared to archive 1, this version imports Biomark Ct spreadsheets that do not include fluorophore data
+
 ```{r setup}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/MergeBiomarkWith10X/")
+knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
+```
+
+```{r}
 library(dplyr)
@@
 projectName <- "Mergebiomark167With10X"
-jackstraw.dim <- 40
 ```
@@
 ```{r}
-source("msAggr_AnalysisCode/read_10XGenomics_data.R")
-source("msAggr_AnalysisCode/PercentVariance.R")
-source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
 ```
@@
 
-```{r}
-color.palette <- c(
-	"coral",
-	"chartreuse4",
-	"goldenrod1",
-	"cadetblue1",
-	"burlywood",
-	"brown",
-	"brown1",
-	"blue",
-	"blue4",
-	"azure3",
-	"aquamarine",
-	"antiquewhite",
-	"cadetblue",
-	"gold3",
-	"black",
-	"darkgreen",
-	"deeppink",
-	"darkviolet",
-	"darkturquoise",
-	"darkslategray",
-	"darksalmon",
-	"darkorchid1",
-	"darkolivegreen2",
-	"forestgreen",
-	"dodgerblue",
-	"green",
-	"lightpink",
-	"lightcoral",
-	"khaki1",
-	"maroon",
-	"peru",
-	"lightseagreen",
-	"lightsalmon",
-	"plum",
-	"moccasin",
-	"tan",
-	"tan1", 
-	"red", 
-	"purple",
-	"khaki4",
-	"black", 
-	"plum4"
-)
-```
-
 # Create seurat object
@@
 
-Will need to include:
-cell list
-probe list
-expression matrix
+```{r}
+biomark.data <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/Biomark_CT.txt", sep = "\t", header = TRUE, row.names = 1)
+biomark.data[is.na(biomark.data)] <- 0
 
+# Limit to CMP data
+biomark.data <- biomark.data[grepl("CMP", rownames(biomark.data)) == TRUE,]
+dim(biomark.data)
+```
+Convert to pseudoexpression data
 ```{r}
-biomark.data <- read.table("biomark167_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
-biomark.data[is.na(biomark.data)] <- 0
-biomark.data.T <- data.table::transpose(biomark.data)
-colnames(biomark.data.T) <- rownames(biomark.data)
-rownames(biomark.data.T) <- colnames(biomark.data)
-# fix non-MGI gene names
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
-```
+# generate list of reference Ct averages, calculating average for ref genes where ref_gene Ct > 0
+ref_list <- c("Actb", "B2m", "Kit")
+biomark.data["refCt"] <- NA
 
-Pull in original cellID info
-
-```{r}
-biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
-biomark.order["clust.ID"] <- NA
-biomark.order[1:49, "clust.ID"] <- 0
-biomark.order[50:87, "clust.ID"] <- 17
-biomark.order[88:102, "clust.ID"] <- 10
-biomark.order[103:117, "clust.ID"] <- 0
-biomark.order[118:127, "clust.ID"] <- 3
-biomark.order[128:150, "clust.ID"] <- 0
-biomark.order[151:189, "clust.ID"] <- 11
-biomark.order[190:281, "clust.ID"] <- 0
-```
-
-```{r}
-# move fluorophor markers to metadata column, remove for expression dataset
-fluor.ids <- colnames(biomark.data[grep("pos", colnames(biomark.data))])
-fluor.table <- biomark.data[fluor.ids]
-biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) %in% fluor.ids),]
-```
-
-```{r warning=FALSE, message=FALSE}
-biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
-colnames(biomark.meta) <- c("wid", "pid")
-
-biomark.meta['nFeature_RNA'] <- NA
-for(cell in colnames(biomark.data.T)){
-	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
-	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
+for(i in 1:nrow(biomark.data)){
+	ref_cts <- c()
+	for(j in ref_list){
+		if(biomark.data[i, j] > 0){
+			ref_cts <- c(ref_cts, biomark.data[i, j])
+		}
+	}
+	biomark.data[i, "refCt"] <- mean(ref_cts)
 }
 
-# Match ids from original hierarchical clustering assignments to cells retained in Biomark167
-i1 <- match(row.names(biomark.meta), row.names(biomark.order))
-biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)
+# generate dCt col
+dbiomark.data <- biomark.data[, 1:ncol(biomark.data) -1]
+dbiomark.data[dbiomark.data == 0] <- NA
+dbiomark.data <- dbiomark.data - biomark.data[, "refCt"]
 
-# match normalized fluorescene 
-i1 <- match(row.names(biomark.meta), row.names(fluor.table))
-biomark.meta[names(fluor.table)[1:ncol(fluor.table)]] <- lapply(fluor.table[1:ncol(fluor.table)], `[`, i1)
+# create pseudo expression matrix
+pseudo.biomark <- 2^-dbiomark.data
 
@@
 ```{r}
-# N.B., can't find alternate name for Kng2, so remove
-biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
-biomark.data.T <- as.sparse(biomark.data.T)
-biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
+# fix non-MGI gene names
+colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'Cd117'] <- 'Kit'
+colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'Cd123'] <- 'Il3ra'
+colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'H2.Aa'] <- 'H2-Aa'
+colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'H2.Ab1'] <- 'H2-Ab1'
+pseudo.biomark <- subset(pseudo.biomark, select = -Kng2)
+```
+
+```{r warning=FALSE, message=FALSE}
+biomark.meta <- data.frame(stringr::str_split_fixed(rownames(pseudo.biomark), "_", 3)[,2:3], row.names = rownames(pseudo.biomark))
+colnames(biomark.meta) <- c("wid", "pid")
+
+pseudo.biomark.T <- data.table::transpose(pseudo.biomark)
+colnames(pseudo.biomark.T) <- rownames(pseudo.biomark)
+rownames(pseudo.biomark.T) <- colnames(pseudo.biomark)
+
+biomark.meta['nFeature_RNA'] <- NA
+for(cell in colnames(pseudo.biomark.T)){
+	ngenes <- length(which(!(pseudo.biomark.T[colnames(pseudo.biomark.T > 0) == cell])))
+	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
+}
+
+```
+
+```{r}
+pseudo.biomark.T[is.na(pseudo.biomark.T)] <- 0
+pseudo.biomark.T <- as.sparse(pseudo.biomark.T)
+biomark <- CreateSeuratObject(pseudo.biomark.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
 ```
@@
 ```{r warning=FALSE}
-setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
+setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
 data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
@@
 seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
-seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
-seurat.object@meta.data$orig.ident <- as.factor(seurat.object@meta.data$orig.ident)
-seurat.object@meta.data$pid <- as.factor(seurat.object@meta.data$pid)
-seurat.object@meta.data$wid <- as.factor(seurat.object@meta.data$wid)
+# seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
+# seurat.object@meta.data$orig.ident <- factor(seurat.object@meta.data$orig.ident)
+# seurat.object@meta.data$pid <- factor(seurat.object@meta.data$pid)
+# seurat.object@meta.data$wid <- factor(seurat.object@meta.data$wid)
 seurat.object
@@
 
+# Traditional normalize and scaling approaches
+trying sctransform and normalizeData/scaledata
 ```{r}
+# seurat.object <- SCTransform(seurat.object, vars.to.regress = c("wid"))
 seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
@@
 all.genes <- rownames(seurat.object)
-seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("orig.ident", "nFeature_RNA", "pid", "nCount_RNA"))
+seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("wid", "pid", "orig.ident"))
 ```
@@
 ```{r}
-# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
+saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
 ```
@@
 
-## Add cluster IDs from Seurat v1
+# scRNASeq integration	
+ neither sctransform nor normalizedata/scaledata gave what looked like reasonable results.
 
-Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
-Create column "clust.ID" and populate with 0's. Then import IDs for clusters
-
-```{r}
-clust3.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
-clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust17.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
-clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust10.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
-clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-clust11.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
-clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
-```
-
-## map cluster ids from Seuratv1
-```{r}
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
-seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
-seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
-seurat.object@meta.data$clust.ID <- as.factor(seurat.object@meta.data$clust.ID)
-```
-
-## Lets see which cells correspond to our old clusters
-
-```{r}
-DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
-```
-
-```{r fig.width=15, fig.height=5}
-DimPlot(seurat.object, split.by = "clust.ID", group.by = "orig.ident", pt.size = 2)
-```
-
-## Split object for anchor-based integration
-
-```{r}
-cmp.subset <- SplitObject(seurat.object, split.by = "orig.ident")$`10X`
-DimPlot(cmp.subset, group.by = "clust.ID")
-```
-
-## Follow integration protocol
-
-https://satijalab.org/seurat/archive/v3.0/integration.html
+https://satijalab.org/seurat/articles/integration_introduction.html
 
@@
 tech.list <- SplitObject(seurat.object, split.by = "orig.ident")
-
-for(i in 1:length(tech.list)){
-	tech.list[[i]] <- NormalizeData(tech.list[[i]], verbose = FALSE)
-	tech.list[[i]] <- FindVariableFeatures(tech.list[[i]], selection.method = "vst", nfeatures = 10, verbose = FALSE)
-}
+tech.list <- lapply(X = tech.list, FUN = function(x) {
+    x <- NormalizeData(x)
+    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 10)
+})
 ```
@@
 ```{r}
-# reference.list <- tech.list["10X"]
-tech.anchors <- FindIntegrationAnchors(object.list = tech.list, dims = 1:30, anchor.features = all.genes, k.anchor = 20, reduction = 'rpca', k.filter = NA)
+
+tech.anchors <- FindIntegrationAnchors(object.list = tech.list, anchor.features = all.genes, k.anchor = 20, k.filter = NA)
 # tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
-tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 5)
+tech.integrated <- IntegrateData(anchorset = tech.anchors, k.weight = 5)
 ```
@@
 Problem now is, how do we find whether Biomark overlays on 10X
-
-```{r fig.height = 5, fig.width=15}
-DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "pca")
-```
-
-```{r fig.height=10, fig.width=10}
-png(filename = "tech.integrated_biomark167Fluor.png", height = 1600, width = 1600)
-FeaturePlot(seurat.object, features = c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos"), split.by = "clust.ID", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 1)
-dev.off()
-
-```
 

### 2021-10-06 MergeBiomarkWith10X/MergeBiomark167With10X.Rmd
Merge 10x and biomark via seurat var95pct

@@
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 library(dplyr)
@@
 library(cowplot)
+library(clustree)
 ```
 
-# Set global variables
+## Set global variables
 ```{r}
+fav.var <- 95
+object.res <- ".0.5"
+ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
 projectName <- "Mergebiomark167With10X"
@@
 source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
 ```
-
 store session info
@@
 dim(biomark.data)
+```
+```{r}
+# fix non-MGI gene names
+colnames(biomark.data)[colnames(biomark.data) == 'Cd117'] <- 'Kit'
+colnames(biomark.data)[colnames(biomark.data) == 'Cd123'] <- 'Il3ra'
+colnames(biomark.data)[colnames(biomark.data) == 'H2.Aa'] <- 'H2-Aa'
+colnames(biomark.data)[colnames(biomark.data) == 'H2.Ab1'] <- 'H2-Ab1'
+biomark.data <- subset(biomark.data, select = -Kng2)
 ```
@@
 
-```
-
-```{r}
-# fix non-MGI gene names
-colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'Cd117'] <- 'Kit'
-colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'Cd123'] <- 'Il3ra'
-colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'H2.Aa'] <- 'H2-Aa'
-colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'H2.Ab1'] <- 'H2-Ab1'
-pseudo.biomark <- subset(pseudo.biomark, select = -Kng2)
 ```
@@
 ```{r}
-tech.list <- SplitObject(seurat.object, split.by = "orig.ident")
+tech.list <- c(cmp.object, biomark)
 tech.list <- lapply(X = tech.list, FUN = function(x) {
@@
 tech.integrated <- RunPCA(tech.integrated, npcs = 30)
-DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident")
+DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident", pt.size = 2)
 ```
@@
 
-Problem now is, how do we find whether Biomark overlays on 10X
+Number of PCs describing X% of variance
 
-## tech.integrated umap
 ```{r}
 tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
-paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
-paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
-paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
-ndims <- length(which(cumsum(tot.var) <= 90))
+paste0("Num pcs for 80% variance: ", length(which(cumsum(tot.var) <= 80)))
+paste0("Num pcs for 85% variance: ", length(which(cumsum(tot.var) <= 85)))
+paste0("Num pcs for 90% variance: ", length(which(cumsum(tot.var) <= 90)))
+paste0("Num pcs for 95% variance: ", length(which(cumsum(tot.var) <= 95)))
+
+```
+
+# ID clusters based on different variances
+## Describe 80% of variance
+
+### Neighborhood and umap
+```{r}
+tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+pct.var <- 80
+ndims <- length(which(cumsum(tot.var) <= pct.var))
+print(ndims)
+
+tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
+tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
+tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5)){
+	tech.integrated <- FindClusters(tech.integrated, resolution = x)
+}
 ```
@@
 ```{r}
-tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims, reduction = "pca")
-tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims, reduction = "pca")
+for (meta.col in colnames(tech.integrated@meta.data)){
+	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
+		plot(myplot)
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
+	}
+}
+```
+
+```{r}
+saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
+```
+
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
+```
+
+## Describe 85% of variance
+
+### Neighborhood and umap
+```{r}
+tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+pct.var <- 85
+ndims <- length(which(cumsum(tot.var) <= pct.var))
+print(ndims)
+
+tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
+tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
+tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5)){
+	tech.integrated <- FindClusters(tech.integrated, resolution = x)
+}
+```
+
+```{r}
+for (meta.col in colnames(tech.integrated@meta.data)){
+	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
+		plot(myplot)
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
+	}
+}
+```
+
+```{r}
+saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
+```
+
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
+```
+
+## Describe 90% of variance
+
+### Neighborhood and umap
+```{r}
+tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+pct.var <- 90
+ndims <- length(which(cumsum(tot.var) <= pct.var))
+print(ndims)
+
+tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
+tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
+tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
+```
+Plot UMAP
+
+```{r}
+for(x in c(0.5, 1, 1.5)){
+	tech.integrated <- FindClusters(tech.integrated, resolution = x)
+}
+```
+
+```{r}
+for (meta.col in colnames(tech.integrated@meta.data)){
+	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
+		plot(myplot)
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
+	}
+}
+```
+
+```{r}
+saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
+```
+
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
+dev.off()
+```
+
+## Describe 95% of variance
+
+### Neighborhood and umap
+```{r}
+tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+pct.var <- 95
+ndims <- length(which(cumsum(tot.var) <= pct.var))
+print(ndims)
+
+tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
+tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
+tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
+```
+Plot UMAP
+
+```{r warning=FALSE, results='hide'}
 for(x in c(0.5, 1, 1.5, 2, 2.5)){
@@
 for (meta.col in colnames(tech.integrated@meta.data)){
-	if(grepl(pattern = ("integrated_snn_res"), x = meta.col)==TRUE){
+	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
 		myplot <- DimPlot(tech.integrated, 
@@
 											reduction = "umap", 
-											cols = color.palette
-											) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("integrated_snn_res", "", meta.col) ))
+											pt.size = 1,
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
 		plot(myplot)
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+		myplot <- DimPlot(tech.integrated, 
+											group.by = meta.col,
+											reduction = "umap", 
+											pt.size = 1,
+											cols = color.palette) + 
+			facet_wrap(meta.col) + 
+			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
+
+		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		plot(myplot)
+		dev.off()
+
 	}
@@
 
-Switch ident
 ```{r}
-Idents(tech.integrated) <- "integrated_snn_res.0.5"
+saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
 ```
 
-```{r fig.width = 15, fig.height=20}
-DimPlot(tech.integrated, group.by = "clust.ID", pt.size = 2, split.by = "ident", ncol = 4, cols = color.palette, shape.by = "orig.ident") + ggmin::theme_powerpoint()
-```
-
-```{r fig.width=20}
-DimPlot(tech.integrated, group.by = "orig.ident", pt.size = 2, ncol = 2)
-```
-
-```{r fig.width=10, fig.height=20}
-DimPlot(tech.integrated, group.by = "orig.ident", pt.size = 2, ncol = 1, split.by = "clust.ID", cols = c("red", "black")) + ggmin::theme_powerpoint()
-```
-
-```{r}
-png("tech.integrated_biomark167UMAP.res.0.5.png", width = 1600, height = 1600)
-DimPlot(tech.integrated, split.by = "ident", group.by = "orig.ident", pt.size = 2, ncol = 5, ) + ggmin::theme_powerpoint()
+### Evaluate cluster stability
+```{r fig.height=5}
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
+	scale_color_continuous(low = 'red3', high = 'white')
+png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
+clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
+	scale_color_continuous(low = 'red3', high = 'white')
 dev.off()
@@
 
+## Create `FindAllMarkers()` lists for GSEA
 ```{r}
-png(filename = "tech.integrated_biomark167FeatureFluor.png", height = 1600, width =1600)
-FeaturePlot(tech.integrated, features = c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos"), reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 3, pt.size = 2)
-dev.off()
-```
+ident.list <- c("integrated_snn_res.0.5")
 
-```{r}
-for(i in c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos")){
-	my_title <- paste0("tech.integrated_biomark167Fluor_", i, ".png")
-	print(my_title)
-png(filename = my_title, height = 1000, width = 6200)
-my_plot <- FeaturePlot(tech.integrated, features = i, split.by = "ident", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 5, pt.size = 3)
-plot(my_plot)
-dev.off()
+for(tested.ident in ident.list){
+	Idents(tech.integrated) <- tested.ident
+	all.markers <- FindAllMarkers(tech.integrated)
+	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+									 file = paste0(projectName, " pctVar", pct.var, "_allres.xlsx"), 
+									 sheetName = tested.ident, 
+									 col.names = TRUE, 
+									 row.names = FALSE, 
+									 append = TRUE)
 }
@@
 
+## Map HGNC symbols
 ```{r}
-# for(i in unique(tech.integrated@meta.data$clust.ID)){
-	# my_title <- paste0("tech.integrated_biomark167clustID_", i, ".png")
-	print(my_title)
-png(filename = "tech.integrated_biomark167clustID.png", height = 1000, width = 6200)
-my_plot <- DimPlot(tech.integrated, pt.size = 6, group.by = "clust.ID", split.by = "ident", shape.by = "orig.ident") + ggmin::theme_powerpoint()
-plot(my_plot)
-dev.off()
-# }
-```
+Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
 
-# Biomark Only
-## Load biomark data
-
-```{r}
-biomark.data <- read.table("biomark167_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
-biomark.data[is.na(biomark.data)] <- 0
-biomark.data.T <- data.table::transpose(biomark.data)
-colnames(biomark.data.T) <- rownames(biomark.data)
-rownames(biomark.data.T) <- colnames(biomark.data)
-# fix non-MGI gene names
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
-rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
-```
-
-Pull in original cellID info
-
-```{r}
-biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
-```
-
-```{r}
-biomark.order["clust.ID"] <- NA
-biomark.order[1:49, "clust.ID"] <- 0
-biomark.order[50:87, "clust.ID"] <- 17
-biomark.order[88:102, "clust.ID"] <- 10
-biomark.order[103:117, "clust.ID"] <- 0
-biomark.order[118:127, "clust.ID"] <- 3
-biomark.order[128:150, "clust.ID"] <- 0
-biomark.order[151:189, "clust.ID"] <- 11
-biomark.order[190:281, "clust.ID"] <- 0
-```
-
-```{r warning=FALSE, message=FALSE}
-biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
-colnames(biomark.meta) <- c("wid", "pid")
-
-biomark.meta['nFeature_RNA'] <- NA
-for(cell in colnames(biomark.data.T)){
-	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
-	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
-}
-i1 <- match(row.names(biomark.meta), row.names(biomark.order))
-biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)
+HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
+head(object.res.allmarkers)
+object.res.allmarkers$HGNC <- HGNC
+tail(object.res.allmarkers)
+sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
+sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
+sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
+sig.res <- sig.res[]
 
@@
 ```{r}
-# N.B., can't find alternate name for Kng2, so remove
-biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
-biomark.data.T <- as.sparse(biomark.data.T)
-biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
+for(cluster in unique(sig.res$cluster)){
+	print(paste("writing cluster", cluster))
+	new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
+	new.table <- new.table[order(-new.table$avg_log2FC), ]
+	try(dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE))
+	try(write.table(new.table, file = paste0(projectName, "RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
+
+}
 ```
 
-```{r}
-biomark <- NormalizeData(biomark, normalization.method = "LogNormalize", scale.factor = 10000)
-biomark <- FindVariableFeatures(biomark, selection.method = "vst", nfeatures = 10)
-all.genes <- rownames(biomark)
-biomark <- ScaleData(biomark, features = all.genes, vars.to.regress = c("nFeature_RNA", "pid", "nCounts_RNA"))
-```
-
-### Save merged object
-```{r}
-# saveRDS(biomark, file = paste0(projectName, "_raw.RDS"))
-```
-## Run PCA
-```{r}
-biomark <- RunPCA(biomark, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
-```
-
-```{r}
-DimPlot(biomark, reduction = "pca", group.by = "orig.ident")
-VizDimLoadings(biomark, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
-
-```
-
-Calculate dimensionality
-```{r}
-ElbowPlot(biomark, ndims = 50)
-percent.variance(biomark@reductions$pca@stdev)
-```
-
-## Calculate percent variance
-```{r}
-tot.var <- percent.variance(biomark@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
-paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
-paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
-paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
-
-```
-
-## Add cluster IDs from Seurat v1
-
-Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
-Create column "clust.ID" and populate with 0's. Then import IDs for clusters
-
-## Lets see which cells correspond to our old clusters
-
-```{r}
-DimPlot(biomark, reduction = "pca", group.by = "orig.ident") + facet_wrap(~orig.ident)
-```
-
-```{r fig.width=15, fig.height=8}
-biomark@meta.data$clust.ID <- as.factor(biomark@meta.data$clust.ID)
-DimPlot(biomark, group.by = "clust.ID", pt.size = 2, shape.by = "clust.ID")
-```
-
-```{r fig.width=15, fig.height = 5}
-DimPlot(biomark, split.by = "clust.ID", pt.size = 2, shape.by = "orig.ident")
-```
-
-# Setup
+# Export 10X for Seaborn clustering
 ## Load libraries required for Seuratv4

### 2021-10-08 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Renamed RMD file to reflect use in analyzing all 3 populations

--- 2021-10-06 MergeBiomarkWith10X/MergeBiomark167With10X.Rmd
@@
 ---
-title: "Mergebiomark167With10X"
+title: "MergeBiomarkWith10X"
 output:
@@
 
+## 20211007
+Gonna crank through LSK and MEP, then go back to the fluorophore data
+
 ## 20211005
@@
 
+# Setup
 ```{r setup}
@@
 ```{r}
+pop.of.interest <- "CMP"
+cellranger.folder <- "m2"
 fav.var <- 95
@@
 ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
-projectName <- "Mergebiomark167With10X"
+projectName <- paste0("MergeBiomarkWith10X_", pop.of.interest)
 ```
@@
 ```{r }
-sink(paste0(projectName, ".txt"))
+sink(paste0(projectName, "_sessionInfo.txt"))
 sessionInfo()
@@
 
-# Limit to CMP data
-biomark.data <- biomark.data[grepl("CMP", rownames(biomark.data)) == TRUE,]
+# Limit to biomark pop subset
+biomark.data <- biomark.data[grepl(pop.of.interest, rownames(biomark.data)) == TRUE,]
 dim(biomark.data)
@@
 pseudo.biomark.T <- as.sparse(pseudo.biomark.T)
-biomark <- CreateSeuratObject(pseudo.biomark.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
+biomark.object <- CreateSeuratObject(pseudo.biomark.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
 ```
@@
 setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
-data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
+data_file.list <- read_10XGenomics_data(sample.list = paste0(pop.of.interest, cellranger.folder))
 object.data <-Read10X(data_file.list)
-cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
+sc.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
 ```
 
-Remove non-biomark genes from cmp.object
+Keep only genes in both objects
 
 ```{r}
-biomark.genes <- rownames(biomark@assays$RNA@data)
-cmp.object <- subset(cmp.object, features = biomark.genes)
-cmp.object@meta.data["wid"] <- '10X'
-cmp.object@meta.data["pid"] <- "p14"
-setdiff(rownames(biomark@assays$RNA@data), rownames(cmp.object@assays$RNA@data))
+shared.features <- intersect(rownames(biomark.object@assays$RNA@data), rownames(sc.object@assays$RNA@data))
+sc.object <- subset(sc.object, features = shared.features)
+biomark.object <- subset(biomark.object, features = shared.features)
+sc.object@meta.data["wid"] <- '10X'
+sc.object@meta.data["pid"] <- "p14"
+
+# Check that there's a clean overlap
+setdiff(rownames(biomark.object@assays$RNA@data), rownames(sc.object@assays$RNA@data))
 ```
@@
 ```{r}
-seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
+seurat.object <- merge(sc.object, biomark.object, project = 'mergebiomark')
+seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- '10X'
+seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == pop.of.interest] <- 'Biomark'
 # seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
@@
 ```{r}
-tech.list <- c(cmp.object, biomark)
+tech.list <- c(sc.object, biomark.object)
 tech.list <- lapply(X = tech.list, FUN = function(x) {
@@
 ```
+
 ## tech.integrated PCA
@@
 ```
-Plot UMAP
+find clusters
 
@@
 ```
-
-```{r}
-saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
-```
-
 ### Evaluate cluster stability
@@
 dev.off()
+```
+
+# Pick set of conditions to investigate further
+Now that have checked multiple pct.vars and object.res's, pick the favourite combination and do more detailed characerization of the. This includes:
+* Stats on each cluster
+* Find markers
+* Export lists for GSEA
+
+## Reload object 
+```{r}
+tech.integrated <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
+tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", object.res))
+```
+
+for each resolution, number/percentage of cells in each cluster?
+
+```{r}
+tot.cells <- nrow(tech.integrated@meta.data)
+for (meta.col in colnames(tech.integrated@meta.data)){
+	if(grepl(pattern = ("integrated_snn_res"), x = meta.col)==TRUE){
+		new.clusters <- sort(as.numeric(levels(tech.integrated@meta.data[[meta.col]])))
+		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
+		colnames(stats.df) <- c("num_cells", "pct_pop")
+		rownames(stats.df) <- new.clusters
+		meta.df <- tech.integrated@meta.data
+		for(row.id in rownames(stats.df)){
+			num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
+			pct.x <- as.integer(num.x / tot.cells *100)
+			# print(pct.x)
+			stats.df[row.id, "num_cells"] <- num.x
+			stats.df[row.id, "pct_pop"] <- pct.x
+		}
+		print(stats.df)
+	}
+}
+```
+```{r}
+tech.integrated <- readRDS("")
+```
+
+```{r}
+current_res <- 'integrated_snn_res.0.5'
+tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- "10X"
+tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest)] <- "biomark"
+tech.integrated@meta.data$orig.ident <- factor(tech.integrated@meta.data$orig.ident)
+
+cluster_ids <- sort(unique(tech.integrated@meta.data[,current_res]))
+counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
+rownames(counts_df) <- cluster_ids
+colnames(counts_df) <- c("10X", "biomark", "pct10X", "pctbiomark")
+tot.10X <- nrow(tech.integrated@meta.data[tech.integrated@meta.data$orig.ident == "10X", ])
+tot.biomark <- nrow(tech.integrated@meta.data[tech.integrated@meta.data$orig.ident == "biomark", ])
+for(id in cluster_ids){
+	cell_value <- nrow(tech.integrated@meta.data[(tech.integrated@meta.data[current_res] == id) & 
+																							 	(tech.integrated@meta.data$orig.ident == "10X"),])
+	counts_df[id, "10X"] = cell_value
+	counts_df[id, "pct10X"] = round(cell_value/tot.10X*100, 2)
+	cell_value <- nrow(tech.integrated@meta.data[(tech.integrated@meta.data[current_res] == id) & 
+																							 	(tech.integrated@meta.data$orig.ident == "biomark"),])
+	counts_df[id, "biomark"] = cell_value
+	counts_df[id, "pctbiomark"] = round(cell_value/tot.biomark*100, 2)
+}
+try(xlsx::write.xlsx(x = counts_df, 
+								 file = paste0(projectName, " pctVar", pct.var, "_stats.xlsx"), 
+								 sheetName = "cluster_stats", 
+								 col.names = TRUE, 
+								 row.names = TRUE, 
+								 append = TRUE)
+)
+counts_df
 ```
@@
 ```{r}
+tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", object.res))
+object.res.allmarkers <- FindAllMarkers(tech.integrated)
 Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
@@
 	new.table <- new.table[order(-new.table$avg_log2FC), ]
-	try(dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE))
-	try(write.table(new.table, file = paste0(projectName, "RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
+	try(dir.create(paste0("RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/"), showWarnings = FALSE))
+	try(write.table(new.table, file = paste0("RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/", projectName, "pctVar", pct.var, "res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
 

### 2021-10-08 MergeBiomarkWith10X/ExportSeuratSlotData.Rmd
Move export of seurat object slot data to new file

--- 2021-10-06 MergeBiomarkWith10X/MergeBiomark167With10X.Rmd
@@
 ---
-title: "Mergebiomark167With10X"
-output:
-  html_document:
-    df_print: paged
+title: "Export data from seurat object slots"
+output: html_notebook
 ---
-
-# Updates
-
-## 20211005
-Compared to archive 1, this version imports Biomark Ct spreadsheets that do not include fluorophore data
-
-```{r setup}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
-```
-
-```{r message=FALSE, warning=FALSE}
-library(dplyr)
-library(Seurat)
-library(patchwork)
-library(ggplot2)
-library(cowplot)
-library(clustree)
-```
-
-## Set global variables
-```{r}
-fav.var <- 95
-object.res <- ".0.5"
-ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
-projectName <- "Mergebiomark167With10X"
-```
-
-```{r}
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
-```
-store session info
-```{r }
-sink(paste0(projectName, ".txt"))
-sessionInfo()
-sink()
-```
-
-# Create seurat object
-
-## Load biomark data
-Try to read in Biomark Data
-
-```{r}
-biomark.data <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/Biomark_CT.txt", sep = "\t", header = TRUE, row.names = 1)
-biomark.data[is.na(biomark.data)] <- 0
-
-# Limit to CMP data
-biomark.data <- biomark.data[grepl("CMP", rownames(biomark.data)) == TRUE,]
-dim(biomark.data)
-```
-```{r}
-# fix non-MGI gene names
-colnames(biomark.data)[colnames(biomark.data) == 'Cd117'] <- 'Kit'
-colnames(biomark.data)[colnames(biomark.data) == 'Cd123'] <- 'Il3ra'
-colnames(biomark.data)[colnames(biomark.data) == 'H2.Aa'] <- 'H2-Aa'
-colnames(biomark.data)[colnames(biomark.data) == 'H2.Ab1'] <- 'H2-Ab1'
-biomark.data <- subset(biomark.data, select = -Kng2)
-```
-Convert to pseudoexpression data
-```{r}
-# generate list of reference Ct averages, calculating average for ref genes where ref_gene Ct > 0
-ref_list <- c("Actb", "B2m", "Kit")
-biomark.data["refCt"] <- NA
-
-for(i in 1:nrow(biomark.data)){
-	ref_cts <- c()
-	for(j in ref_list){
-		if(biomark.data[i, j] > 0){
-			ref_cts <- c(ref_cts, biomark.data[i, j])
-		}
-	}
-	biomark.data[i, "refCt"] <- mean(ref_cts)
-}
-
-# generate dCt col
-dbiomark.data <- biomark.data[, 1:ncol(biomark.data) -1]
-dbiomark.data[dbiomark.data == 0] <- NA
-dbiomark.data <- dbiomark.data - biomark.data[, "refCt"]
-
-# create pseudo expression matrix
-pseudo.biomark <- 2^-dbiomark.data
-
-```
-
-```{r warning=FALSE, message=FALSE}
-biomark.meta <- data.frame(stringr::str_split_fixed(rownames(pseudo.biomark), "_", 3)[,2:3], row.names = rownames(pseudo.biomark))
-colnames(biomark.meta) <- c("wid", "pid")
-
-pseudo.biomark.T <- data.table::transpose(pseudo.biomark)
-colnames(pseudo.biomark.T) <- rownames(pseudo.biomark)
-rownames(pseudo.biomark.T) <- colnames(pseudo.biomark)
-
-biomark.meta['nFeature_RNA'] <- NA
-for(cell in colnames(pseudo.biomark.T)){
-	ngenes <- length(which(!(pseudo.biomark.T[colnames(pseudo.biomark.T > 0) == cell])))
-	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
-}
-
-```
-
-```{r}
-pseudo.biomark.T[is.na(pseudo.biomark.T)] <- 0
-pseudo.biomark.T <- as.sparse(pseudo.biomark.T)
-biomark <- CreateSeuratObject(pseudo.biomark.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
-```
-
-## Load SC data
-
-```{r warning=FALSE}
-setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
-data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
-object.data <-Read10X(data_file.list)
-cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
-```
-
-Remove non-biomark genes from cmp.object
-
-```{r}
-biomark.genes <- rownames(biomark@assays$RNA@data)
-cmp.object <- subset(cmp.object, features = biomark.genes)
-cmp.object@meta.data["wid"] <- '10X'
-cmp.object@meta.data["pid"] <- "p14"
-setdiff(rownames(biomark@assays$RNA@data), rownames(cmp.object@assays$RNA@data))
-```
-
-Clean up to free memory
-
-```{r}
-remove(object.data, biomark.data, biomark.meta)
-```
-
-## merge objects
-```{r}
-seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
-# seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
-# seurat.object@meta.data$orig.ident <- factor(seurat.object@meta.data$orig.ident)
-# seurat.object@meta.data$pid <- factor(seurat.object@meta.data$pid)
-# seurat.object@meta.data$wid <- factor(seurat.object@meta.data$wid)
-seurat.object
-```
-
-# Traditional normalize and scaling approaches
-trying sctransform and normalizeData/scaledata
-```{r}
-# seurat.object <- SCTransform(seurat.object, vars.to.regress = c("wid"))
-seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
-seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
-```
-
-Scale data (linear transformation)
-
-```{r}
-all.genes <- rownames(seurat.object)
-seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("wid", "pid", "orig.ident"))
-```
-
-### Save merged object
-```{r}
-saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
-```
-## Run PCA
-```{r}
-seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
-```
-
-```{r}
-DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
-```
-
-```{r}
-VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
-
-```
-
-Calculate dimensionality
-```{r, figures-side, fig.show='hold', out.width="50%"}
-ElbowPlot(seurat.object, ndims = 50)
-percent.variance(seurat.object@reductions$pca@stdev)
-```
-
-## Calculate percent variance
-```{r}
-tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
-paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
-paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
-paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))
-
-```
-
-# scRNASeq integration	
-neither sctransform nor normalizedata/scaledata gave what looked like reasonable results.
-
-https://satijalab.org/seurat/articles/integration_introduction.html
-
-```{r}
-tech.list <- c(cmp.object, biomark)
-tech.list <- lapply(X = tech.list, FUN = function(x) {
-	x <- NormalizeData(x)
-	x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 10)
-})
-```
-
-## identify anchors
-```{r}
-
-tech.anchors <- FindIntegrationAnchors(object.list = tech.list, anchor.features = all.genes, k.anchor = 20, k.filter = NA)
-# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
-tech.integrated <- IntegrateData(anchorset = tech.anchors, k.weight = 5)
-```
-## tech.integrated PCA
-
-```{r}
-DefaultAssay(tech.integrated) <- "integrated"
-tech.integrated <- ScaleData(tech.integrated)
-tech.integrated <- RunPCA(tech.integrated, npcs = 30)
-DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident", pt.size = 2)
-```
-
-so it looks like anchor-based integration is much nicer! goign to stick with this approach.
-
-Number of PCs describing X% of variance
-
-```{r}
-tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-paste0("Num pcs for 80% variance: ", length(which(cumsum(tot.var) <= 80)))
-paste0("Num pcs for 85% variance: ", length(which(cumsum(tot.var) <= 85)))
-paste0("Num pcs for 90% variance: ", length(which(cumsum(tot.var) <= 90)))
-paste0("Num pcs for 95% variance: ", length(which(cumsum(tot.var) <= 95)))
-
-```
-
-# ID clusters based on different variances
-## Describe 80% of variance
-
-### Neighborhood and umap
-```{r}
-tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-pct.var <- 80
-ndims <- length(which(cumsum(tot.var) <= pct.var))
-print(ndims)
-
-tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
-tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
-tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
-```
-Plot UMAP
-
-```{r}
-for(x in c(0.5, 1, 1.5)){
-	tech.integrated <- FindClusters(tech.integrated, resolution = x)
-}
-```
-
-```{r}
-for (meta.col in colnames(tech.integrated@meta.data)){
-	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
-		plot(myplot)
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			facet_wrap(meta.col) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
-
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-
-	}
-}
-```
-
-```{r}
-saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
-```
-
-### Evaluate cluster stability
-```{r fig.height=5}
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
-	scale_color_continuous(low = 'red3', high = 'white')
-dev.off()
-```
-
-## Describe 85% of variance
-
-### Neighborhood and umap
-```{r}
-tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-pct.var <- 85
-ndims <- length(which(cumsum(tot.var) <= pct.var))
-print(ndims)
-
-tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
-tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
-tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
-```
-Plot UMAP
-
-```{r}
-for(x in c(0.5, 1, 1.5)){
-	tech.integrated <- FindClusters(tech.integrated, resolution = x)
-}
-```
-
-```{r}
-for (meta.col in colnames(tech.integrated@meta.data)){
-	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
-		plot(myplot)
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			facet_wrap(meta.col) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
-
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-
-	}
-}
-```
-
-```{r}
-saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
-```
-
-### Evaluate cluster stability
-```{r fig.height=5}
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
-	scale_color_continuous(low = 'red3', high = 'white')
-dev.off()
-```
-
-## Describe 90% of variance
-
-### Neighborhood and umap
-```{r}
-tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-pct.var <- 90
-ndims <- length(which(cumsum(tot.var) <= pct.var))
-print(ndims)
-
-tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
-tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
-tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
-```
-Plot UMAP
-
-```{r}
-for(x in c(0.5, 1, 1.5)){
-	tech.integrated <- FindClusters(tech.integrated, resolution = x)
-}
-```
-
-```{r}
-for (meta.col in colnames(tech.integrated@meta.data)){
-	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
-		plot(myplot)
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			facet_wrap(meta.col) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
-
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-
-	}
-}
-```
-
-```{r}
-saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
-```
-
-### Evaluate cluster stability
-```{r fig.height=5}
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
-	scale_color_continuous(low = 'red3', high = 'white')
-dev.off()
-```
-
-## Describe 95% of variance
-
-### Neighborhood and umap
-```{r}
-tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
-pct.var <- 95
-ndims <- length(which(cumsum(tot.var) <= pct.var))
-print(ndims)
-
-tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
-tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
-tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
-```
-Plot UMAP
-
-```{r warning=FALSE, results='hide'}
-for(x in c(0.5, 1, 1.5, 2, 2.5)){
-	tech.integrated <- FindClusters(tech.integrated, resolution = x)
-}
-```
-
-```{r}
-for (meta.col in colnames(tech.integrated@meta.data)){
-	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
-		plot(myplot)
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-		myplot <- DimPlot(tech.integrated, 
-											group.by = meta.col,
-											reduction = "umap", 
-											pt.size = 1,
-											cols = color.palette) + 
-			facet_wrap(meta.col) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
-
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
-		plot(myplot)
-		dev.off()
-
-	}
-}
-```
-
-```{r}
-saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
-```
-
-### Evaluate cluster stability
-```{r fig.height=5}
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
-	scale_color_continuous(low = 'red3', high = 'white')
-png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
-clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
-	scale_color_continuous(low = 'red3', high = 'white')
-dev.off()
-```
-
-## Create `FindAllMarkers()` lists for GSEA
-```{r}
-ident.list <- c("integrated_snn_res.0.5")
-
-for(tested.ident in ident.list){
-	Idents(tech.integrated) <- tested.ident
-	all.markers <- FindAllMarkers(tech.integrated)
-	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
-									 file = paste0(projectName, " pctVar", pct.var, "_allres.xlsx"), 
-									 sheetName = tested.ident, 
-									 col.names = TRUE, 
-									 row.names = FALSE, 
-									 append = TRUE)
-}
-```
-
-## Map HGNC symbols
-```{r}
-Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
-
-HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
-head(object.res.allmarkers)
-object.res.allmarkers$HGNC <- HGNC
-tail(object.res.allmarkers)
-sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
-sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
-sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
-sig.res <- sig.res[]
-
-```
-
-```{r}
-for(cluster in unique(sig.res$cluster)){
-	print(paste("writing cluster", cluster))
-	new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
-	new.table <- new.table[order(-new.table$avg_log2FC), ]
-	try(dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE))
-	try(write.table(new.table, file = paste0(projectName, "RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
-
-}
-```
 
@@
 ```
-
-```{r}
-biomark.object <- readRDS("~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/CMPSubpop_pctVar90.RDS")
-```

### 2021-10-08 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Move export of seurat object slot data to new file

@@
 ```
-
-# Export 10X for Seaborn clustering
-## Load libraries required for Seuratv4
-```{r message=FALSE, warning=FALSE}
-knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/")
-knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
-library(dplyr)
-library(Seurat)
-library(patchwork)
-library(ggplot2)
-library(clustree)
-```
-## Source functions
-```{r}
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
-source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
-```
-
-## Set global variables
-```{r}
-pop.id <- "CMP"
-folder.appendix <- "m2"
-fav.var <- 90
-object.res <- ".0.5"
-ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
-folder.id <- paste0(pop.id, folder.appendix)
-projectName <- paste0(pop.id, "Subpop_vBiomark")
-```
-
-## Store session info
-```{r}
-sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
-sink(sessionInfo.filename)
-sessionInfo()
-sink()
-```
-
-*STOP* - testing showed that cells cluster better in seaborn when you include all probe data points.
-It was an intersting idea to cluster based only on the genes in the biomark probe set, but this doesn't seem to hold up in practice
-
-```{r warning=FALSE, message=FALSE}
-setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
-data_file.list <- read_10XGenomics_data(sample.list = c(folder.id))
-data.object<-Read10X(data_file.list)
-seurat.object<- CreateSeuratObject(counts = data.object, min.cells = 3, min.genes = 200, project = projectName)
-remove(data.object)
-probe.list <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/BiomarkRNAs.txt", header = FALSE)
-probe.list <- unlist(as.list(probe.list), use.names = FALSE)
-probe.list <- c(probe.list, "Kit", "Il3ra")
-# Subset to biomark features
-biomark.object <- subset(seurat.object, features = probe.list)
-biomark.object[["percent.mt"]] <- PercentageFeatureSet(biomark.object, pattern = "^mt-")
-print(paste("original object:", nrow(biomark.object@meta.data), "cells", sep = " "))
-biomark.object <- subset(biomark.object, 
-												 subset = nFeature_RNA >=25 & 
-												 	percent.mt <= 5
-)
-print(paste("new object:", nrow(biomark.object@meta.data), "cells", sep = " "))
-biomark.object <- NormalizeData(biomark.object, normalization.method = "LogNormalize", scale.factor = 10000)
-object.downsample <- subset(biomark.object, cells = sample(Cells(biomark.object), 300))
-```
-
-## overlay data from clustered seurat.object onto biomark.object
-```{r}
-seurat.object <- readRDS("~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/CMPSubpop_pctVar90.RDS")
-xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "counts"))))
-object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
-# biomark.export <- xprsn.mtx
-biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
-colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
-colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
-write.table(biomark.export, file = paste0(projectName, "_Counts-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
-
-xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "data"))))
-object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
-# biomark.export <- xprsn.mtx
-biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
-colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
-colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
-write.table(biomark.export, file = paste0(projectName, "_Data-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
-
-xprsn.mtx <- t(as.data.frame(as.matrix(GetAssayData(object.downsample, slot = "scale.data"))))
-object.metadata <- seurat.object@meta.data[rownames(seurat.object@meta.data) %in% rownames(xprsn.mtx), ]
-# biomark.export <- xprsn.mtx
-biomark.export <- merge(xprsn.mtx, object.metadata[paste0("RNA_snn_res", object.res)], by = 0)
-colnames(biomark.export)[colnames(biomark.export) == "Kit"] <- "Cd117"
-colnames(biomark.export)[colnames(biomark.export) == "Il3ra"] <- "Cd123"
-write.table(biomark.export, file = paste0(projectName, "_ScaleData-BiomarkProbes.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)
-
-```
-
-```{r}
-biomark.object <- readRDS("~/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/CMPSubpop_pctVar90.RDS")
-```

### 2021-10-08 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Generate GSEA lists for Biomark and 10X merging

@@
 # Setup
-```{r setup}
+```{r setup, include=FALSE}
 knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
@@
 ```{r}
-pop.of.interest <- "CMP"
+pop.of.interest <- "LSK"
 cellranger.folder <- "m2"
@@
 ```
-
+source code
 ```{r}
@@
 pseudo.biomark.T <- as.sparse(pseudo.biomark.T)
-biomark.object <- CreateSeuratObject(pseudo.biomark.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
+biomark.object <- CreateSeuratObject(pseudo.biomark.T, project = "biomark", assay = "RNA", meta.data = biomark.meta)
 ```
@@
 seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- '10X'
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == pop.of.interest] <- 'Biomark'
+seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == pop.of.interest] <- 'biomark'
 # seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
@@
 ```{r}
-saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
+# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
 ```
@@
 ```{r}
-
 tech.anchors <- FindIntegrationAnchors(object.list = tech.list, anchor.features = all.genes, k.anchor = 20, k.filter = NA)
@@
 ```
+```{r}
+saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
+```
 
@@
 ```
-
 for each resolution, number/percentage of cells in each cluster?
-
-```{r}
-tot.cells <- nrow(tech.integrated@meta.data)
-for (meta.col in colnames(tech.integrated@meta.data)){
-	if(grepl(pattern = ("integrated_snn_res"), x = meta.col)==TRUE){
-		new.clusters <- sort(as.numeric(levels(tech.integrated@meta.data[[meta.col]])))
-		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
-		colnames(stats.df) <- c("num_cells", "pct_pop")
-		rownames(stats.df) <- new.clusters
-		meta.df <- tech.integrated@meta.data
-		for(row.id in rownames(stats.df)){
-			num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
-			pct.x <- as.integer(num.x / tot.cells *100)
-			# print(pct.x)
-			stats.df[row.id, "num_cells"] <- num.x
-			stats.df[row.id, "pct_pop"] <- pct.x
-		}
-		print(stats.df)
-	}
-}
-```
-```{r}
-tech.integrated <- readRDS("")
-```
-
 ```{r}
@@
 ```{r}
-ident.list <- c("integrated_snn_res.0.5")
+ident.list <- c(paste0("integrated_snn_res", object.res))
 
@@
 object.res.allmarkers$HGNC <- HGNC
-tail(object.res.allmarkers)
+# tail(object.res.allmarkers)
 sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
@@
 
-```
-
-```{r}
 for(cluster in unique(sig.res$cluster)){
@@
 	new.table <- new.table[order(-new.table$avg_log2FC), ]
-	try(dir.create(paste0("RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/"), showWarnings = FALSE))
-	try(write.table(new.table, file = paste0("RankList_pctVar", pct.var, "res", object.res, "_findAll_hgnc/", projectName, "pctVar", pct.var, "res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
+	try(dir.create(paste0("RankList_findAll_hgnc/"), showWarnings = FALSE))
+	try(write.table(new.table, file = paste0("RankList_findAll_hgnc/", projectName, "pctVar", pct.var, "res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
 

### 2021-10-14 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Completed script to combine 10X and biomark into tech.integrated seurat objects and export biomark flow data as ridgeplots

@@
 # Updates
+## 20211011
+Have successfully co-clustered LSK, CMP, and MEP scCells with respective biomark populations. Now have to determine if the fluorophore data supports this.  
+Will load the fluorophore data in as metadata, and try to plot the fluor metadata based on the sc/biomark clusters.
+not sure if the NAs in sc will be a problem, but can subset to the biomark object and try that way.
 
@@
 }
+```
+```{r}
 try(xlsx::write.xlsx(x = counts_df, 
@@
 	all.markers <- FindAllMarkers(tech.integrated)
-	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+	try(xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
 									 file = paste0(projectName, " pctVar", pct.var, "_allres.xlsx"), 
@@
 									 append = TRUE)
+			)
 }
@@
 ```
+
+# Merge index sorting data
+
+## load object
+```{r}
+tech.integrated <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
+tech.integrated <- SetIdent(tech.integrated, value = "integrated_snn_res.0.5")
+tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- "10X"
+tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest)] <- "biomark"
+tech.integrated@meta.data$orig.ident <- factor(tech.integrated@meta.data$orig.ident)
+```
+
+## load index sort data
+```{r}
+mfi <- read.csv(file = paste0(pop.of.interest, "_mfi_unNormalized.txt"), sep = "\t", row.names = 1, skipNul = TRUE)
+head(mfi)
+```
+
+sort_dict = {'010319':['p1', 'p3'], '121218': ['p2', 'p4'], '020119': ['p6', 'p7'], '090618': ['p12', 'p13']}
+normalize based on sort data  
+Plot to check distribution
+```{r}
+for(colname in colnames(mfi[,grepl("pos$", names(mfi))])){
+	h <- hist(mfi[, colname], plot = FALSE) 
+	h$density = h$counts/sum(h$counts)*100
+	plot(h,freq=FALSE, main = colname, xlab = "mfi", ylab = "% Cells")
+	}
+```
+So some distributions are gaussian, and some are poisson. Best bet is probaby just to normalize everyone.  
+> Min-max normalization: guarantees the same scale, but is thrown by outliers  
+> Z score normalization: does not guarantee the same scale, but handles outliers better
+
+Flow is full of outliers. Going to z-score normalize
+
+```{r}
+fluor.cols <- names(mfi[, grepl("pos$", names(mfi))])
+mfi.norm <- data.frame(matrix(nrow = 0, ncol = length(fluor.cols)))
+colnames(mfi.norm) <- fluor.cols
+for(sortdate in unique(mfi$sortdate)){
+	mfi.subset <- scale(mfi[mfi$sortdate == sortdate, fluor.cols])
+	mfi.norm <- rbind(mfi.norm, mfi.subset)
+}
+mfi.norm["pid"] <- do.call(rbind, strsplit(rownames(mfi.norm), '_'))[,3]
+```
+Add metadata column
+```{r}
+tech.integrated <- AddMetaData(tech.integrated, metadata = mfi.norm)
+```
+now plot some stuff
+```{r}
+biomark.integrated <- subset(tech.integrated, subset = orig.ident == "biomark")
+```
+```{r fig.height=12, fig.width=5}
+RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
+```
+```{r}
+png(filename = paste0(projectName, " pctVar", fav.var, "res.", object.res, "-fluor.png"), height = 2400, width = 800)
+RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
+dev.off()
+```

### 2021-10-15 RFunctions/metadata_overlap_stats.R
added function to calculate overlapping number of cells between metadata columns in seurat object

@@ -1,0 +1,52 @@
+#' metadata_overlap_stas
+#'
+#' @description Generates table describing percentage of one metadata col's ids in another metadata col's ids. Note that is currently hardcoded to require numeric data
+#'
+#' @param seurat.object object of class Seurat
+#' @param target.col.pattern metadata column pattern to match against
+#' @param xlsx_filename if not NULL, generate an excel file of the enrichment tables with this name
+#' @param test.col.pattern metadata column pattern to test for enrichment in coluns matching target.col.pattern
+#' @param print_to_screen whether or not to print enrichment dataframes to screen
+#' @param xlsx.sheet.count.start if exporting xlsx, sheet number to start counting at. Ignored if xlsx_filename = NULL
+#'
+#' @return
+#' @export
+#'
+#' @examples
+metadata_overlap_stats <- function(seurat.object, target.col.pattern, test.col.pattern, print_to_screen = TRUE, xlsx_filename = NULL, xlsx.sheet.count.start = 1){
+	xlsx.sheet.count <- xlsx.sheet.count.start
+	for(test.col in colnames(seurat.object@meta.data)){
+		if(grepl(pattern = test.col.pattern, x = test.col) == TRUE){
+			test.col.id <- levels(seurat.object@meta.data[, test.col])
+			for(target.col in colnames(seurat.object@meta.data)){
+				if(grepl(pattern = target.col.pattern, x = target.col) == TRUE){
+					print(paste("Comparing", test.col, "and", target.col))
+					target.col.id <- levels(seurat.object@meta.data[, target.col])
+					enrich.df <- data.frame(matrix(nrow = length(target.col.id), ncol = length(test.col.id)))
+					rownames(enrich.df) <- target.col.id
+					colnames(enrich.df) <- test.col.id
+					seurat.meta.df <- seurat.object@meta.data
+					for(col.id in colnames(enrich.df)){
+						tot.test.cells <- nrow(seurat.meta.df[seurat.meta.df[[test.col]] == col.id,])
+						for(row.id in rownames(enrich.df)){
+							num.x <- nrow(seurat.meta.df[seurat.meta.df[[target.col]] == row.id & seurat.meta.df[[test.col]] == col.id,])
+							pct.x <- as.integer((num.x / tot.test.cells)*100)
+							enrich.df[row.id, col.id] <- pct.x
+						}
+					}
[…]

### 2021-10-15 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
adding code to overlapy cluster ids from integrated 10X and biomark seurat object back onto original 10X seurat object

@@
 # Updates
+## 20211014
+Newest idea (C/O Dave) is to take the 10X cells from tech.integrated (the merged Biomark/SingeCell object) and find them on the sc.object.
+Then add the metadata from  `tech.integrated@metadata$integrated_snn_res`.
+Running `FindMarkers()` on this parameter might help define population types *without* being biased by the biomark probes.
+
 ## 20211011
@@
 ```{r}
-pop.of.interest <- "LSK"
+pop.of.interest <- "CMP"
 cellranger.folder <- "m2"
-fav.var <- 95
-object.res <- ".0.5"
+integrated.var <- 95
+integrated.res <- ".0.5"
 ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
 projectName <- paste0("MergeBiomarkWith10X_", pop.of.interest)
+sc.pop.var <- 90
+sc.pop.res <- ".0.5"
 ```
@@
 
-# Create seurat object
+# Create seurat objects
 
@@
 colnames(biomark.data)[colnames(biomark.data) == 'H2.Ab1'] <- 'H2-Ab1'
-biomark.data <- subset(biomark.data, select = -Kng2)
+if("Kng2" %in% biomark.data){biomark.data <- subset(biomark.data, select = -Kng2)}
 ```
@@
 shared.features <- intersect(rownames(biomark.object@assays$RNA@data), rownames(sc.object@assays$RNA@data))
-sc.object <- subset(sc.object, features = shared.features)
+sc.subset <- subset(sc.object, features = shared.features)
 biomark.object <- subset(biomark.object, features = shared.features)
-sc.object@meta.data["wid"] <- '10X'
-sc.object@meta.data["pid"] <- "p14"
+sc.subset@meta.data["wid"] <- '10X'
+sc.subset@meta.data["pid"] <- "p14"
 
 # Check that there's a clean overlap
-setdiff(rownames(biomark.object@assays$RNA@data), rownames(sc.object@assays$RNA@data))
+setdiff(rownames(biomark.object@assays$RNA@data), rownames(sc.subset@assays$RNA@data))
 ```
@@
 ```{r}
-seurat.object <- merge(sc.object, biomark.object, project = 'mergebiomark')
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- '10X'
-seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == pop.of.interest] <- 'biomark'
-# seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
-# seurat.object@meta.data$orig.ident <- factor(seurat.object@meta.data$orig.ident)
-# seurat.object@meta.data$pid <- factor(seurat.object@meta.data$pid)
-# seurat.object@meta.data$wid <- factor(seurat.object@meta.data$wid)
-seurat.object
+merged.object <- merge(sc.subset, biomark.object, project = 'mergebiomark')
+merged.object@meta.data$orig.ident[merged.object@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- '10X'
+merged.object@meta.data$orig.ident[merged.object@meta.data$orig.ident == pop.of.interest] <- 'biomark'
+# merged.object@meta.data[is.na(merged.object@meta.data)] <- 0
+# merged.object@meta.data$orig.ident <- factor(merged.object@meta.data$orig.ident)
+# merged.object@meta.data$pid <- factor(merged.object@meta.data$pid)
+# merged.object@meta.data$wid <- factor(merged.object@meta.data$wid)
+merged.object
 ```
@@
 ```{r}
-# seurat.object <- SCTransform(seurat.object, vars.to.regress = c("wid"))
-seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
-seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
+# merged.object <- SCTransform(merged.object, vars.to.regress = c("wid"))
+merged.object <- NormalizeData(merged.object, normalization.method = "LogNormalize", scale.factor = 10000)
+merged.object <- FindVariableFeatures(merged.object, selection.method = "vst", nfeatures = 10)
 ```
@@
 ```{r}
-all.genes <- rownames(seurat.object)
-seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("wid", "pid", "orig.ident"))
+all.genes <- rownames(merged.object)
+merged.object <- ScaleData(merged.object, features = all.genes, vars.to.regress = c("wid", "pid", "orig.ident"))
 ```
@@
 ```{r}
-# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
+# saveRDS(merged.object, file = paste0(projectName, "_raw.RDS"))
 ```
@@
 ```{r}
-seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
+merged.object <- RunPCA(merged.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
 ```
@@
 ```{r}
-DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
+DimPlot(merged.object, reduction = "pca", group.by = "orig.ident")
 ```
@@
 ```{r}
-VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
+VizDimLoadings(merged.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)
 
@@
 ```{r, figures-side, fig.show='hold', out.width="50%"}
-ElbowPlot(seurat.object, ndims = 50)
-percent.variance(seurat.object@reductions$pca@stdev)
+ElbowPlot(merged.object, ndims = 50)
+percent.variance(merged.object@reductions$pca@stdev)
 ```
@@
 ```{r}
-tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+tot.var <- percent.variance(merged.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
 paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
@@
 ```{r}
-tech.list <- c(sc.object, biomark.object)
+tech.list <- c(sc.subset, biomark.object)
 tech.list <- lapply(X = tech.list, FUN = function(x) {
@@
 # Pick set of conditions to investigate further
-Now that have checked multiple pct.vars and object.res's, pick the favourite combination and do more detailed characerization of the. This includes:
+Now that have checked multiple pct.vars and integrated.res's, pick the favourite combination and do more detailed characerization of the. This includes:
 * Stats on each cluster
@@
 ```{r}
-tech.integrated <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
-tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", object.res))
+tech.integrated <- readRDS(paste0(projectName, "_pctVar", integrated.var, ".RDS"))
+tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", integrated.res))
 ```
@@
 ```{r}
-ident.list <- c(paste0("integrated_snn_res", object.res))
+ident.list <- c(paste0("integrated_snn_res", integrated.res))
 
@@
 ```{r}
-tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", object.res))
-object.res.allmarkers <- FindAllMarkers(tech.integrated)
-Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
+tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", integrated.res))
+integrated.res.allmarkers <- FindAllMarkers(tech.integrated)
+Mouse2HumanTable <- Mouse2Human(integrated.res.allmarkers$gene)
 
-HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
-head(object.res.allmarkers)
-object.res.allmarkers$HGNC <- HGNC
-# tail(object.res.allmarkers)
-sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
+HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(integrated.res.allmarkers$gene, Mouse2HumanTable$MGI)])
+head(integrated.res.allmarkers)
+integrated.res.allmarkers$HGNC <- HGNC
+# tail(integrated.res.allmarkers)
+sig.res <- integrated.res.allmarkers[integrated.res.allmarkers$p_val_adj <= 0.05, ]
 sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
@@
 	try(dir.create(paste0("RankList_findAll_hgnc/"), showWarnings = FALSE))
-	try(write.table(new.table, file = paste0("RankList_findAll_hgnc/", projectName, "pctVar", pct.var, "res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
+	try(write.table(new.table, file = paste0("RankList_findAll_hgnc/", projectName, "pctVar", pct.var, "res", integrated.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))
 
@@
 ```{r}
-tech.integrated <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
+tech.integrated <- readRDS(paste0(projectName, "_pctVar", integrated.var, ".RDS"))
 tech.integrated <- SetIdent(tech.integrated, value = "integrated_snn_res.0.5")
@@
 ```{r}
-png(filename = paste0(projectName, " pctVar", fav.var, "res.", object.res, "-fluor.png"), height = 2400, width = 800)
+png(filename = paste0(projectName, " pctVar", integrated.var, "res.", integrated.res, "-fluor.png"), height = 2400, width = 800)
 RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
@@
 ```
+
+# Add tech.integrated cluster IDs to sc.object
+## Load analyzed sc.object from Indivdual_SubpopAnalysis
+Use previously determined optimal percent variance to describe (`sc.pop.var`) and resolution (`sc.pop.res`) from notebook "Individual_SubpopAnalysis" and clustree.
+Will increase resolution if can get improved coverage of integrated_snn clusters
+```{r}
+seurat.object <- readRDS(paste0("/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/", pop.of.interest, "Subpop_pctVar", sc.pop.var, ".RDS"))
+seurat.object
+```
+Get cluster IDs from merged SC and Biomark seurat object (tech.integrated) and add them back to the original sc.object.
+```{r}
+seurat.object <- AddMetaData(seurat.object, metadata = tech.integrated@meta.data[, ident.list])
+```
+Find the percent of integrated_snn_res.X cells for each RNA_snn_res.in the seurat.object
+Test: what percentage of each new clusterID matches one of the older clusters?
+
+```{r}
+test.col.pattern <- "integrated_snn_res"
+target.col.pattern <- "RNA_snn_res"
+xlsx_filename <- paste0(projectName, "_pctVar", pct.var, "_mergedEnrichmentStats.xlsx")
+
+metadata_overlap_stats(seurat.object = seurat.object, 
+											 target.col.pattern = target.col.pattern, 
+											 test.col.pattern = test.col.pattern, 
+											 print_to_screen = TRUE,
+											 xlsx_filename = xlsx_filename)
+```
+
+```{r}
+for(test.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = test.col.pattern, x = test.col) == TRUE){
+		test.col.id <- levels(seurat.object@meta.data[, test.col])
+		for(target.col in colnames(seurat.object@meta.data)){
+			if(grepl(pattern = target.col.pattern, x = target.col) == TRUE){
+				target.col.id <- levels(seurat.object@meta.data[, target.col])
+				enrich.df <- data.frame(matrix(nrow = length(target.col.id), ncol = length(test.col.id)))
+				rownames(enrich.df) <- target.col.id
+				colnames(enrich.df) <- test.col.id
+				seurat.meta.df <- seurat.object@meta.data
+				print(test.col)
+				for(col.id in colnames(enrich.df)){
+					tot.test.cells <- nrow(seurat.meta.df[seurat.meta.df[[test.col]] == col.id,])
+					for(row.id in rownames(enrich.df)){
+						num.x <- nrow(seurat.meta.df[seurat.meta.df[[target.col]] == row.id & seurat.meta.df[[test.col]] == col.id,])
+						pct.x <- as.integer(num.x / tot.test.cells*100)
+						enrich.df[row.id, col.id] <- num.x
+					}
+				}
+				if(!sum(rowSums(enrich.df)) == nrow(seurat.object@meta.data)){print("error! Not all cells have been assigned to a category!")}
+				colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0(test.col.pattern, "_", x))
+				rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0(target.col.pattern, "_", x))
+				print(enrich.df)
+			}
+		}
+	}
+}
+```
+
+```{r}
+for(test.col in colnames(seurat.object@meta.data)){
+	if(grepl(pattern = test.col.pattern, x = test.col)){
+		test.col.id <- sort(as.numeric(levels(seurat.object@meta.data[[test.col]])))
+		for (target.col in colnames(seurat.object@meta.data)){
+			if(grepl(pattern = target.col.pattern, x = target.col)){
+				match.col.id <- sort(as.numeric(levels(seurat.object@meta.data[[target.col]])))}
+				enrich.df <- data.frame(matrix(ncol = length(test.col.id), nrow = length(match.col.id)))
+				colnames(enrich.df) <- test.col.id
+				rownames(enrich.df) <- match.col.id
+				meta.df <- seurat.object@meta.data
+				for(row.id in rownames(enrich.df)){
+					tot.clus <- nrow(meta.df[meta.df[[target.col]] == row.id,])
+					for(col.id in colnames(enrich.df)){
+						num.x <- nrow(meta.df[(meta.df[[target.col]] == row.id) & (meta.df$clust.ID == col.id),])
+						pct.x <- as.integer(num.x / tot.clus *100)
+						# print(pct.x)
+						enrich.df[row.id, col.id] <- pct.x
+					}
+				}
+				colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0(test.col.pattern, "_", x))
+				rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0(target.col.pattern, "_", x))
+				print(sum(enrich.df))
+
+				if(!is.null(xlsx_filename)){
+					xlsx::write.xlsx(enrich.df, file = xlsx_filename, append = TRUE)}
+
+		}
+	}
+}
+```

### 2021-10-18 RFunctions/metadata_overlap_stats.R
Added function to calculate percentages of overlapping metadata columns for seurat objects

@@ -22,21 +22,26 @@ metadata_overlap_stats <- function(seurat.object, target.col.pattern, test.col.p
 				if(grepl(pattern = target.col.pattern, x = target.col) == TRUE){
 					print(paste("Comparing", test.col, "and", target.col))
 					target.col.id <- levels(seurat.object@meta.data[, target.col])
-					enrich.df <- data.frame(matrix(nrow = length(target.col.id), ncol = length(test.col.id)))
-					rownames(enrich.df) <- target.col.id
-					colnames(enrich.df) <- test.col.id
+					enrich.df <- data.frame(matrix(nrow = length(target.col.id) + 1, ncol = length(test.col.id) + 1))
+					rownames(enrich.df) <- c(target.col.id, "tot_cells_in_test_cluster")
+					colnames(enrich.df) <- c(test.col.id, "tot_cells_in_target_cluster")
 					seurat.meta.df <- seurat.object@meta.data
 					for(col.id in colnames(enrich.df)){
 						tot.test.cells <- nrow(seurat.meta.df[seurat.meta.df[[test.col]] == col.id,])
 						for(row.id in rownames(enrich.df)){
+							tot.target.cells <- nrow(seurat.meta.df[seurat.meta.df[[target.col]] == row.id, ])
 							num.x <- nrow(seurat.meta.df[seurat.meta.df[[target.col]] == row.id & seurat.meta.df[[test.col]] == col.id,])
 							pct.x <- as.integer((num.x / tot.test.cells)*100)
 							enrich.df[row.id, col.id] <- pct.x
+							enrich.df[row.id, "tot_cells_in_target_cluster"] <- tot.target.cells
+							enrich.df["tot_cells_in_test_cluster", col.id] <- tot.test.cells
 						}
 					}
 					# if(!sum(rowSums(enrich.df)) == nrow(seurat.object@meta.data)){print("error! Not all cells have been assigned to a category!")}
-					colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0(test.col, "_", x))
-					rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0(target.col, "_", x))
+					colnames(enrich.df)[1:ncol(enrich.df) - 1] <- sapply(colnames(enrich.df)[1:ncol(enrich.df) - 1], function(x) paste0(test.col, "_", x))
+					# print(length(rownames(enrich.df)))
[…]

### 2021-10-18 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Calculated and plotted overlapping tech.integrated cluster assignments onto original 10X Seurat object

@@
 # Setup
-```{r setup, include=FALSE}
+```{r setup}
 knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
@@
 integrated.res <- ".0.5"
-ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
+integrated.ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
 projectName <- paste0("MergeBiomarkWith10X_", pop.of.interest)
@@
 sc.pop.res <- ".0.5"
+sc.ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2")
 ```
@@
 											cols = color.palette) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
+			ggtitle(paste0(projectName, "_pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
 		plot(myplot)
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		png(filename = paste0(projectName, "_pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
 		plot(myplot)
@@
 			facet_wrap(meta.col) + 
-			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
+			ggtitle(paste0(projectName, "_pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))
 
-		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		png(filename = paste0(projectName, "_pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
 		plot(myplot)
@@
 try(xlsx::write.xlsx(x = counts_df, 
-								 file = paste0(projectName, " pctVar", pct.var, "_stats.xlsx"), 
+								 file = paste0(projectName, "_pctVar", pct.var, "_stats.xlsx"), 
 								 sheetName = "cluster_stats", 
@@
 ```{r}
-ident.list <- c(paste0("integrated_snn_res", integrated.res))
+integrated.ident.list <- c(paste0("integrated_snn_res", integrated.res))
 
-for(tested.ident in ident.list){
+for(tested.ident in integrated.ident.list){
 	Idents(tech.integrated) <- tested.ident
@@
 	try(xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
-									 file = paste0(projectName, " pctVar", pct.var, "_allres.xlsx"), 
+									 file = paste0(projectName, "_pctVar", pct.var, "_allres.xlsx"), 
 									 sheetName = tested.ident, 
@@
 ```{r}
-png(filename = paste0(projectName, " pctVar", integrated.var, "res.", integrated.res, "-fluor.png"), height = 2400, width = 800)
+png(filename = paste0(projectName, "_pctVar", integrated.var, "res.", integrated.res, "-fluor.png"), height = 2400, width = 800)
 RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
@@
 ```{r}
-seurat.object <- AddMetaData(seurat.object, metadata = tech.integrated@meta.data[, ident.list])
+seurat.object <- AddMetaData(seurat.object, metadata = tech.integrated@meta.data[, integrated.ident.list])
 ```
@@
 
+plot overlays
 ```{r}
-for(test.col in colnames(seurat.object@meta.data)){
-	if(grepl(pattern = test.col.pattern, x = test.col) == TRUE){
-		test.col.id <- levels(seurat.object@meta.data[, test.col])
-		for(target.col in colnames(seurat.object@meta.data)){
-			if(grepl(pattern = target.col.pattern, x = target.col) == TRUE){
-				target.col.id <- levels(seurat.object@meta.data[, target.col])
-				enrich.df <- data.frame(matrix(nrow = length(target.col.id), ncol = length(test.col.id)))
-				rownames(enrich.df) <- target.col.id
-				colnames(enrich.df) <- test.col.id
-				seurat.meta.df <- seurat.object@meta.data
-				print(test.col)
-				for(col.id in colnames(enrich.df)){
-					tot.test.cells <- nrow(seurat.meta.df[seurat.meta.df[[test.col]] == col.id,])
-					for(row.id in rownames(enrich.df)){
-						num.x <- nrow(seurat.meta.df[seurat.meta.df[[target.col]] == row.id & seurat.meta.df[[test.col]] == col.id,])
-						pct.x <- as.integer(num.x / tot.test.cells*100)
-						enrich.df[row.id, col.id] <- num.x
-					}
-				}
-				if(!sum(rowSums(enrich.df)) == nrow(seurat.object@meta.data)){print("error! Not all cells have been assigned to a category!")}
-				colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0(test.col.pattern, "_", x))
-				rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0(target.col.pattern, "_", x))
-				print(enrich.df)
-			}
-		}
+for(integrated.col in integrated.ident.list){
+	for(sc.col in sc.ident.list){
+		plot.title <- paste0(projectName, "_pctVar", pct.var, "-", integrated.col, " faceted by ", sc.col)
+		myplot <- DimPlot(seurat.object, 
+											group.by = integrated.col,
+											split.by = sc.col, 
+											reduction = "umap", 
+											pt.size = 1.5,
+											ncol = round(sqrt(length(levels(seurat.object@meta.data[[sc.col]])))),
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, "_pctVar", pct.var, "-", integrated.col, " faceted by ", sc.col))
+
+		png(filename = paste0(projectName, "_pctVar", pct.var, 
+													"-integratedRes.", gsub("integrated_snn_res.", "", integrated.col), 
+													"mergedEnrichment-Facet_scRes.", gsub("RNA_snn_res.", "", sc.col), 
+													".png"), 
+				height = 1600, width = 1600)
+		plot(myplot)
+		dev.off()
+
 	}
@@
 ```
-
-```{r}
-for(test.col in colnames(seurat.object@meta.data)){
-	if(grepl(pattern = test.col.pattern, x = test.col)){
-		test.col.id <- sort(as.numeric(levels(seurat.object@meta.data[[test.col]])))
-		for (target.col in colnames(seurat.object@meta.data)){
-			if(grepl(pattern = target.col.pattern, x = target.col)){
-				match.col.id <- sort(as.numeric(levels(seurat.object@meta.data[[target.col]])))}
-				enrich.df <- data.frame(matrix(ncol = length(test.col.id), nrow = length(match.col.id)))
-				colnames(enrich.df) <- test.col.id
-				rownames(enrich.df) <- match.col.id
-				meta.df <- seurat.object@meta.data
-				for(row.id in rownames(enrich.df)){
-					tot.clus <- nrow(meta.df[meta.df[[target.col]] == row.id,])
-					for(col.id in colnames(enrich.df)){
-						num.x <- nrow(meta.df[(meta.df[[target.col]] == row.id) & (meta.df$clust.ID == col.id),])
-						pct.x <- as.integer(num.x / tot.clus *100)
-						# print(pct.x)
-						enrich.df[row.id, col.id] <- pct.x
-					}
-				}
-				colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0(test.col.pattern, "_", x))
-				rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0(target.col.pattern, "_", x))
-				print(sum(enrich.df))
-
-				if(!is.null(xlsx_filename)){
-					xlsx::write.xlsx(enrich.df, file = xlsx_filename, append = TRUE)}
-
-		}
-	}
-}
-```

### 2021-10-20 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
completed mergeBiomarkWith10X umap and stats code for CMP

@@
 # Setup
-```{r setup}
+```{r setup, message=FALSE, warning=FALSE}
 knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
-```
-
-```{r message=FALSE, warning=FALSE}
 library(dplyr)
@@
 sc.pop.res <- ".0.5"
-sc.ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2")
+sc.ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
 ```
@@
 ```{r}
+source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/metadata_overlap_stats.R")
 source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
@@
 ```
-
 # Create seurat objects
@@
 ```{r}
-current_res <- 'integrated_snn_res.0.5'
+res.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
+for(current_res in res.list){
 	tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- "10X"
@@
 	}
-```
-```{r}
 	try(xlsx::write.xlsx(x = counts_df, 
 											 file = paste0(projectName, "_pctVar", pct.var, "_stats.xlsx"), 
-								 sheetName = "cluster_stats", 
+											 sheetName = current_res, 
 											 col.names = TRUE, 
@@
 	counts_df
+}
 ```
@@
 ```
+plot umaps
+```{r}
+for(integrated.col in integrated.ident.list){
+		plot.title <- paste0(projectName, "_pctVar", pct.var, "-mergedEnrichment_OrigIdent-Facet-integratedRes.", gsub("integrated_snn_res.", "", integrated.col))
+		myplot <- DimPlot(tech.integrated, 
+											group.by = "orig.ident",
+											split.by = integrated.col, 
+											reduction = "umap", 
+											pt.size = 1.5,
+											ncol = round(sqrt(length(levels(seurat.object@meta.data[[sc.col]])))),
+											cols = color.palette) + 
+			ggtitle(paste0(projectName, "_pctVar", pct.var, "-", integrated.col, " faceted by ", sc.col))
+		plot(myplot)
+
+		png(filename = paste0(projectName, "_pctVar", pct.var,
+													"-mergedEnrichment_OrigIdent-Facet-integratedRes.", gsub("integrated_snn_res.", "", integrated.col),
+													".png"),
+				height = 1600, width = 1600)
+		plot(myplot)
+		dev.off()
+
+}
+```
 
@@
 ```
+
+Save object	
+```{r}
+# saveRDS(seurat.object, file = paste0("/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/", pop.of.interest, "Subpop_pctVar", sc.pop.var, ".RDS"))
+```
+# DGE
+```{r}
+for(tested.ident in integrated.ident.list){
+	Idents(seurat.object) <- tested.ident
+	all.markers <- FindAllMarkers(seurat.object)
+	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+									 file = paste0(projectName, "_FindALLMarkers_pctVar", pct.var, "_allres.xlsx"), 
+									 sheetName = tested.ident, 
+									 col.names = TRUE, 
+									 row.names = FALSE, 
+									 append = TRUE)
+}
+```

### 2021-11-03 msAggr_seurat/msAggr_seurat.Rmd
creating single script to streamline individual pop analyses
Adding code to generate 3D model

@@
 title: "msAggr_seurat"
-output: 
-  html_notebook: 
-    fig_height: 6
+output:s
+  html_document:
+    df_print: paged
 ---
@@
 * SCTransform accounts better for sources of variability
+
+# Generating 3D models
+```{r}
+rds.file <- "msAggr_seurat_dim26.RDS"
+seurat.object <- readRDS(rds.file)
+ndims <- as.numeric(gsub("[^0-9]", "", stringr::str_split(rds.file, "_")[[1]][3]))
+```
+
+```{r}
+tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= 90))
+
+seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
+seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+
+```

### 2021-11-03 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd

@@
 library(clustree)
+# library(scatterplot3d)
+library(rgl)
 ```
@@
 plot umaps
+for(integrated.col in integrated.ident.list){
 ```{r}
-for(integrated.col in integrated.ident.list){
 		plot.title <- paste0(projectName, "_pctVar", pct.var, "-mergedEnrichment_OrigIdent-Facet-integratedRes.", gsub("integrated_snn_res.", "", integrated.col))
@@
 ```
+
+# Plot 3D
+Plot in 3d
+```{r}
+tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= integrated.var))
+
+tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
+tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
+tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims, n.components = 3)
+```
+Plot 3D UMAP
+
+```{r}
+my_object <- tech.integrated@reductions$umap@cell.embeddings
+# color.palette @ res.0.5 has 15 levels. Change color.palette[11] to a different color so can add the biomark cells as large black dots afterwards
+color.palette[11] <- "mediumpurple1"
+par3d("windowRect" = c(122, 44, 950, 929))
+plot3d(x=my_object[,1], y=my_object[,2], z=my_object[,3], 
+			 col = color.palette[tech.integrated@active.ident], 
+			 box=FALSE, axes=FALSE, xlab = "", ylab = "", zlab = "")
+rgl.lines(c(-10, 10), c(0, 0), c(0, 0), color = "black")
+rgl.lines(c(0, 0), c(-10, 10), c(0, 0), color = "red")
+rgl.lines(c(0, 0), c(0, 0), c(-10, 10), color = "green")
+
+# biomark.embeddings <- subset(tech.integrated, subset = orig.ident == "biomark")
+# biomark.embeddings <- biomark.embeddings@reductions$umap@cell.embeddings
+# plot3d(x = biomark.embeddings[,1], y = biomark.embeddings[,2], z = biomark.embeddings[,3],
+# 			 col = "black", size = 10,
+# 			 add = TRUE)
+rgl.viewpoint(zoom = .5)
+movie3d(spin3d(axis = c(1, 1, 1), rpm = 6), duration = 10, dir = getwd(), clean = TRUE, webshot = TRUE)
+```

### 2021-11-05 msAggr_seurat/msAggr_seurat.Rmd
Added code for 3D umap plotting
Added GO code for bubble plots

@@
 library(clustree)
+library(rgl)
 ```
@@
 			facet_wrap(meta.col) + 
-			ggtitle(paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col)))
+			ggtitle(paste0(projectName, "dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col)))
 
-		png(filename = paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		png(filename = paste0(projectName, "dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
 		plot(myplot)
@@
 ```
+Plot gene expression
+```{r}
+feature.list <- c("Meis1", "Gata2", "Gata1", "Irf8", "Elane", "Fli1", "Zfpm1")
 
+for(gene in feature.list){
+	png(filename = paste0(projectName, "dim", ndims, "-FeaturePlot-", gene, ".png"), height = 400, width = 400)
+	plot(FeaturePlot(seurat.object, features = gene))
+	dev.off()
+}
+
+```
+
+Save seurat.object RDS
 ```{r}
@@
 for(tested.ident in ident.list){
-	Idents(seurat.object) <- tested.ident
+	ident.res <- gsub("RNA_snn_res", "", tested.ident)
+	seurat.object <- SetIdent(seurat.object, value = tested.ident)
 	all.markers <- FindAllMarkers(seurat.object)
-	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
-									 file = paste0(projectName, "_FindALLMarkers_dim",ndims, "_allres.xlsx"), 
-									 sheetName = tested.ident, 
-									 col.names = TRUE, 
-									 row.names = FALSE, 
-									 append = TRUE)
-}
-```
-
-## Create `FindAllMarkers()` lists for GSEA
-```{r}
-object.res.allmarkers <- FindAllMarkers(seurat.object)
-```
+	# xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
+	# 								 file = paste0(projectName, "_FindALLMarkers_dim",ndims, ".xlsx"), 
+	# 								 sheetName = paste0("res", ident.res), 
+	# 								 col.names = TRUE, 
+	# 								 row.names = FALSE, 
+	# 								 append = TRUE)
 
 	## Map HGNC symbols
-```{r}
-Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)
+	Mouse2HumanTable <- Mouse2Human(all.markers$gene)
 
-HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
-head(object.res.allmarkers)
-object.res.allmarkers$HGNC <- HGNC
-tail(object.res.allmarkers)
-sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
+	HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(all.markers$gene, Mouse2HumanTable$MGI)])
+	all.markers$HGNC <- HGNC
+	sig.res <- all.markers[all.markers$p_val_adj <= 0.05, ]
 	sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
@@
 
-```
-
-```{r}
 	for(cluster in unique(sig.res$cluster)){
@@
 		new.table <- new.table[order(-new.table$avg_log2FC), ]
-	dir.create(paste0("RankList_res", object.res, "_findAll_hgnc/"), showWarnings = FALSE)
-	write.table(new.table, file = paste0("RankList_res", object.res, "_findAll_hgnc/res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		try(dir.create(paste0("RankList_res", ident.res, "_findAll_hgnc/"), showWarnings = FALSE))
+		write.table(new.table, file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+	}
+}
+```
 
-}
+## topGO with FindAllMarkers
+```{r}
+source("/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/GoEnrichmentAnalysis.R")
+GO_enrichment_function(my_folder = "RankList_res.0.5_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
 ```
@@
 seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
-seurat.object <- FindClusters(seurat.object, resolution = 0.5)
-seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
+# seurat.object <- FindClusters(seurat.object, resolution = 0.5)
+seurat.object <- RunUMAP(seurat.object, dims = 1: ndims, n.components = 3)
+```
+
+```{r}
+for(x in c(0.5, 1, 1.5, 2, 2.5)){
+	seurat.object <- FindClusters(seurat.object, resolution = x)
+}
+```
+Set object ident
+```{r}
+seurat.object <- SetIdent(seurat.object, value = "RNA_snn_res.0.5")
+```
+
+```{r}
+my_object <- seurat.object@reductions$umap@cell.embeddings
+# color.palette @ res.0.5 has 15 levels. Change color.palette[11] to a different color so can add the biomark cells as large black dots afterwards
+color.palette[11] <- "mediumpurple1"
+par3d("windowRect" = c(122, 44, 950, 929))
+bg3d("white", alpha = 0)
+plot3d(x=my_object[,1], y=my_object[,2], z=my_object[,3], 
+			 col = color.palette[seurat.object@meta.data$orig.ident], 
+			 box=FALSE, axes=FALSE, xlab = "", ylab = "", zlab = "")
+rgl.lines(c(-10, 10), c(0, 0), c(0, 0), color = "black")
+rgl.lines(c(0, 0), c(-10, 10), c(0, 0), color = "red")
+rgl.lines(c(0, 0), c(0, 0), c(-10, 10), color = "green")
+legend3d("topright", c(levels(seurat.object@meta.data$orig.ident)), col = color.palette, pch = 19, cex = 1)
+# biomark.embeddings <- subset(tech.integrated, subset = orig.ident == "biomark")
+# biomark.embeddings <- biomark.embeddings@reductions$umap@cell.embeddings
+# plot3d(x = biomark.embeddings[,1], y = biomark.embeddings[,2], z = biomark.embeddings[,3],
+# 			 col = "black", size = 10,
+# 			 add = TRUE)
+rgl.viewpoint(zoom = .5)
+movie3d(spin3d(axis = c(1, 1, 1), rpm = 6), duration = 10, dir = getwd(), clean = TRUE, webshot = FALSE, movie = paste0(projectName, "_dim", ndims, "umap_movie-orig.ident"))
+```
+
+```{r}
+my_object <- seurat.object@reductions$umap@cell.embeddings
+# color.palette @ res.0.5 has 15 levels. Change color.palette[11] to a different color so can add the biomark cells as large black dots afterwards
+color.palette[11] <- "mediumpurple1"
+par3d("windowRect" = c(122, 44, 950, 929))
+bg3d("white", alpha = 0)
+plot3d(x=my_object[,1], y=my_object[,2], z=my_object[,3], 
+			 col = color.palette[seurat.object@meta.data$RNA_snn_res.0.5], 
+			 box=FALSE, axes=FALSE, xlab = "", ylab = "", zlab = "")
+rgl.lines(c(-10, 10), c(0, 0), c(0, 0), color = "black")
+rgl.lines(c(0, 0), c(-10, 10), c(0, 0), color = "red")
+rgl.lines(c(0, 0), c(0, 0), c(-10, 10), color = "green")
+legend3d("topright", c(levels(seurat.object@meta.data$RNA_snn_res.0.5)), col = color.palette, pch = 19, cex = 1)
+rgl.viewpoint(zoom = .5)
+movie3d(spin3d(axis = c(1, 1, 1), rpm = 6), duration = 10, dir = getwd(), clean = TRUE, webshot = FALSE, movie = paste0(projectName, "_dim", ndims, "res.0.5-umap_movie"))
 

### 2021-11-05 RFunctions/metadata_stats.R
Additional functions for GO enrichment and 3D plots

@@ -1,0 +1,30 @@
+#' Metadata.Stats
+#' @description for each metadata column matching regex X, what's that number and percentage of cells in each?
+#'
+#' @param seurat.object object of class Seurat
+#' @param metadata.pattern grepl pattern to match for metadata columns
+#'
+#' @return
+#' @export
+#'
+#' @examples
+metadata_stats <- function(seurat.object, metadata.pattern){
+	tot.cells <- nrow(seurat.object@meta.data)
+	for (meta.col in colnames(cmp.object@meta.data)){
+		if(grepl(pattern = metadata.pattern, x = meta.col)==TRUE){
+			new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
+			stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
+			colnames(stats.df) <- c("num_cells", "pct_pop")
+			rownames(stats.df) <- new.clusters
+			meta.df <- seurat.object@meta.data
+			for(row.id in rownames(stats.df)){
+				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
+				pct.x <- as.integer(num.x / tot.cells *100)
+				# print(pct.x)
+				stats.df[row.id, "num_cells"] <- num.x
+				stats.df[row.id, "pct_pop"] <- pct.x
+			}
+			print(stats.df)
+		}
+	}
+}

### 2021-11-08 IndividualPops/Individual_SubpopAnalyses.Rmd

@@
 fav.var <- 90
+pct.var <- 90
 object.res <- ".1"
@@
 # ID clusters based on different variances
-## Describe 80% of variance with 12 dimensions
+## Describe 80% of variance
 
@@
 calculate `FindAllMarkers()` for different idents and save to new file  
+### [Note from devs](https://www.biostars.org/p/409790/):
+"FindMarkers will find markers between two different identity groups - you have to specify both identity groups. This is useful for comparing the differences between two specific groups.
+
+FindAllMarkers will find markers differentially expressed in each identity group by comparing it to all of the others - you don't have to manually define anything.
+Note that markers may bleed over between closely-related groups - they are not forced to be specific to only one group.
+This is what most people use (and likely what you want).
+
+FindConservedMarkers will find markers that are conserved between two groups - this can be useful if you want to find markers that are conserved between a treated and untreated condition for a specific cell type or group of cells.
+It means they are differentially expressed compared to other groups, but have similar expression between the two groups you're actually comparing."
 ```{r}
@@
 
-calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
+calculate `FindMarkers()` that distinguish each cluster 
 ```{r}

### 2021-11-08 MergeBiomarkWith10X/MergeBiomark281With10X.Rmd

---
title: "MergeBiomark281With10X"
output: html_notebook
---

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/MergeBiomarkWith10X/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
```

# Set global variables
```{r}
projectName <- "MergeBiomark281With10X"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")
source("msAggr_AnalysisCode/PercentVariance.R")
source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
```

store session info
```{r }
sink(paste0(projectName, ".txt"))
sessionInfo()
sink()
```

# Create seurat object

## Load biomark data
Try to read in Biomark Data

Will need to include:
cell list
probe list
expression matrix

```{r}
biomark.data <- read.table("Biomark281_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0
biomark.data.T <- data.table::transpose(biomark.data)
colnames(biomark.data.T) <- rownames(biomark.data)
rownames(biomark.data.T) <- colnames(biomark.data)
# fix non-MGI gene names
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
```

Pull in original cellID info

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r}
fluor.table <- biomark.data[grepl("pos", colnames(biomark.data))]
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
colnames(biomark.meta) <- c("wid", "pid")

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

i1 <- match(row.names(biomark.meta), row.names(fluor.table))
biomark.meta[names(fluor.table)[1:ncol(fluor.table)]] <- lapply(fluor.table[1:ncol(fluor.table)], `[`, i1)

```

```{r}
# N.B., can't find alternate name for Kng2, so remove
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
biomark.data.T <- as.sparse(biomark.data.T)
biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

## Load CMP data

```{r warning=FALSE}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
```

Remove non-biomark genes from cmp.object

```{r}
biomark.genes <- rownames(biomark@assays$RNA@data)
cmp.object <- subset(cmp.object, features = biomark.genes)
cmp.object@meta.data["wid"] <- '10X'
cmp.object@meta.data["pid"] <- "p14"
setdiff(rownames(biomark@assays$RNA@data), rownames(cmp.object@assays$RNA@data))
```

Clean up to free memory

```{r}
remove(object.data, biomark.data, biomark.meta)
```

## merge objects
```{r}
seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
seurat.object
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
```

Scale data (linear transformation)

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("orig.ident", "nFeature_RNA", "pid", "nCount_RNA"))
```

### Save merged object
```{r}
# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

```{r}
clust3.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

## map cluster ids from Seuratv1
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
seurat.object@meta.data$clust.ID <- as.factor(seurat.object@meta.data$clust.ID)
```

## Lets see which cells correspond to our old clusters

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r fig.width=15, fig.height=5}
DimPlot(seurat.object, split.by = "clust.ID", group.by = "orig.ident", pt.size = 2)
```

## Split object for anchor-based integration

```{r}
cmp.subset <- SplitObject(seurat.object, split.by = "orig.ident")$`10X`
DimPlot(cmp.subset, group.by = "clust.ID")
```

## Follow integration protocol

https://satijalab.org/seurat/archive/v3.0/integration.html

```{r}
tech.list <- SplitObject(seurat.object, split.by = "orig.ident")

for(i in 1:length(tech.list)){
	tech.list[[i]] <- NormalizeData(tech.list[[i]], verbose = FALSE)
	tech.list[[i]] <- FindVariableFeatures(tech.list[[i]], selection.method = "vst", nfeatures = 10, verbose = FALSE)
}
```

## identify anchors
```{r}
# reference.list <- tech.list["10X"]
tech.anchors <- FindIntegrationAnchors(object.list = tech.list, dims = 1:30, anchor.features = all.genes, k.anchor = 20, reduction = 'rpca', k.filter = NA)
# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 5)
```
## tech.integrated PCA

```{r}
DefaultAssay(tech.integrated) <- "integrated"
tech.integrated <- ScaleData(tech.integrated)
tech.integrated <- RunPCA(tech.integrated, npcs = 30)
DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident")
```

so it looks like anchor-based integration is much nicer! goign to stick with this approach.

Problem now is, how do we find whether Biomark overlays on 10X

```{r fig.height = 5, fig.width=15}
DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "pca")
```

```{r fig.height=10, fig.width=10}
png(filename = "tech.integrated_biomark281Fluor.png", height = 1600, width = 1600)
FeaturePlot(tech.integrated, features = c("CD34pos", "cKitpos", "CD9pos", "CD36pos", "CD41pos", "CD48pos", "CD123pos", "CD150pos", "Flk2pos"), split.by = "clust.ID", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 1)
dev.off()

```

## tech.integrated umap
```{r}
tech.integrated <- RunUMAP(tech.integrated, reduction = "pca", dims = 1:30)
```
## Plot UMAP
```{r fig.width=15, fig.height=5}
DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "umap")
```

# Biomark Only
## Load biomark data

```{r}
biomark.data <- read.table("Biomark281_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0
biomark.data.T <- data.table::transpose(biomark.data)
colnames(biomark.data.T) <- rownames(biomark.data)
rownames(biomark.data.T) <- colnames(biomark.data)
# fix non-MGI gene names
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
```

Pull in original cellID info

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
```

```{r}
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
colnames(biomark.meta) <- c("wid", "pid")

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

```

```{r}
# N.B., can't find alternate name for Kng2, so remove
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
biomark.data.T <- as.sparse(biomark.data.T)
biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

```{r}
biomark <- NormalizeData(biomark, normalization.method = "LogNormalize", scale.factor = 10000)
biomark <- FindVariableFeatures(biomark, selection.method = "vst", nfeatures = 10)
all.genes <- rownames(biomark)
biomark <- ScaleData(biomark, features = all.genes, vars.to.regress = c("nFeature_RNA", "pid", "nCounts_RNA"))
```

### Save merged object
```{r}
# saveRDS(biomark, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
biomark <- RunPCA(biomark, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(biomark, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(biomark, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r}
ElbowPlot(biomark, ndims = 50)
percent.variance(biomark@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(biomark@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

## Lets see which cells correspond to our old clusters

```{r}
DimPlot(biomark, reduction = "pca", group.by = "orig.ident") + facet_wrap(~orig.ident)
```

```{r fig.width=15, fig.height=8}
biomark@meta.data$clust.ID <- as.factor(biomark@meta.data$clust.ID)
DimPlot(biomark, group.by = "clust.ID", pt.size = 2, shape.by = "clust.ID")
```

```{r fig.width=15, fig.height = 5}
DimPlot(biomark, split.by = "clust.ID", pt.size = 2, shape.by = "orig.ident")
```

### 2021-11-08 msAggr_seurat/msAggr_seurat.Rmd

@@
 											cols = color.palette) + 
+			NoLegend() + 
 			ggtitle(paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col) ))
 		plot(myplot)
-		png(filename = paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
+		png(filename = paste0(projectName, " dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800, bg = "transparent")
 		plot(myplot)
@@
 
-		png(filename = paste0(projectName, "dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
+		png(filename = paste0(projectName, "dim", ndims, "res.", gsub("RNA_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800, bg = "transparent")
 		plot(myplot)
@@
 ```
+Plot all black cells
+```{r}
+png(filename = paste0(projectName, " dim", ndims, "-umap.png"), height = 800, width = 800, bg = "transparent")
+DimPlot(seurat.object, cols = c("black", "black", "black", "black"), group.by = "orig.ident") + NoLegend()
+dev.off()
+```
+
 Plot gene expression
@@
 # ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
-ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+# ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
+# ident.list <- c("RNA_snn_res.1")
+ident.list <- c("RNA_snn_res.0.5")
+cut_off <- 1.2
+load(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/mouseIDTable.rda")
 
@@
 	## Map HGNC symbols
-	Mouse2HumanTable <- Mouse2Human(all.markers$gene)
 
-	HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(all.markers$gene, Mouse2HumanTable$MGI)])
+	HGNC <- with(mouseIDTable, mouseIDTable$HGNC[match(all.markers$gene, mouseIDTable$MGI)])
 	all.markers$HGNC <- HGNC
@@
 		write.table(new.table, file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		write.table(new.table[new.table$avg_log2FC >= cut_off,], file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, "-co", cut_off, "up.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		write.table(new.table[new.table$avg_log2FC >= (cut_off)*-1,], file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, "-co", cut_off, "dn.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
 	}
@@
 GO_enrichment_function(my_folder = "RankList_res.0.5_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
+# GO_enrichment_function(my_folder = "RankList_res.1_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
 ```

### 2021-11-12 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd

--- 2021-10-20 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
@@
 integrated.var <- 95
-integrated.res <- ".0.5"
-integrated.ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
+integrated.res <- ".1"
+integrated.ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1", "integrated_snn_res.1.5")
 projectName <- paste0("MergeBiomarkWith10X_", pop.of.interest)
@@
 ```{r}
-DimPlot(merged.object, reduction = "pca", group.by = "orig.ident")
+DimPlot(merged.object, reduction = "pca", group.by = "orig.ident", cols = c("black", "red"), pt.size = 1)
+png(filename = paste0(projectName, "merged_PCA.png"), height = 800, width = 800)
+DimPlot(merged.object, reduction = "pca", group.by = "orig.ident", cols = c("black", "red"), pt.size = 2)
+dev.off()
 ```
@@
 tech.integrated <- readRDS(paste0(projectName, "_pctVar", integrated.var, ".RDS"))
-tech.integrated <- SetIdent(tech.integrated, value = "integrated_snn_res.0.5")
+tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", integrated.res))
 tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- "10X"
@@
 ```{r}
-png(filename = paste0(projectName, "_pctVar", integrated.var, "res.", integrated.res, "-fluor.png"), height = 2400, width = 800)
+png(filename = paste0(projectName, "_pctVar", integrated.var, "res", integrated.res, "-fluor.png"), height = 2400, width = 800)
 RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
@@
 											pt.size = 1.5,
-											ncol = round(sqrt(length(levels(seurat.object@meta.data[[sc.col]])))),
+											ncol = round(sqrt(length(levels(tech.integrated@meta.data[[sc.col]])))),
 											cols = color.palette) + 
@@
 ```
+
+# Plot 3D
+Plot in 3d
+```{r}
+tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
+ndims <- length(which(cumsum(tot.var) <= integrated.var))
+tech.integrated.3d <- tech.integrated
+tech.integrated.3d <- FindNeighbors(tech.integrated.3d, dims = 1:ndims)
+tech.integrated.3d <- FindClusters(tech.integrated.3d, resolution = as.numeric(integrated.res))
+tech.integrated.3d <- RunUMAP(tech.integrated.3d, dims = 1: ndims, n.components = 3)
+```
+Plot 3D UMAP
+
+```{r}
+tech.integrated.3d <- SetIdent(tech.integrated.3d, value = paste0("integrated_snn_res", integrated.res))
+my_object <- tech.integrated.3d@reductions$umap@cell.embeddings
+
+# color.palette @ res.0.5 has 15 levels. Change color.palette[11] to a different color so can add the biomark cells as large black dots afterwards
+color.palette[11] <- "mediumpurple1"
+par3d("windowRect" = c(122, 44, 950, 929))
+plot3d(x=my_object[,1], y=my_object[,2], z=my_object[,3], 
+			 col = color.palette[tech.integrated.3d@active.ident], 
+			 box=FALSE, axes=FALSE, xlab = "", ylab = "", zlab = "")
+rgl.lines(c(-10, 10), c(0, 0), c(0, 0), color = "black")
+rgl.lines(c(0, 0), c(-10, 10), c(0, 0), color = "red")
+rgl.lines(c(0, 0), c(0, 0), c(-10, 10), color = "green")
+
+biomark.embeddings <- subset(tech.integrated.3d, subset = orig.ident == "biomark")
+biomark.embeddings <- biomark.embeddings@reductions$umap@cell.embeddings
+plot3d(x = biomark.embeddings[,1], y = biomark.embeddings[,2], z = biomark.embeddings[,3],
+			 col = "black", size = 10,
+			 add = TRUE)
+rgl.viewpoint(zoom = .5)
+movie3d(spin3d(axis = c(1, 1, 1), rpm = 6), duration = 10, dir = getwd(), clean = TRUE, webshot = TRUE, movie = "MergeBiomarkWith10X_CMP_pctVar95res.1.5")
+```
+
+# Find Seurat v1 cells
+## Add cluster IDs from Seurat v1
+Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
+Create column "clust.ID" and populate with 0's. Then import IDs for clusters
+```{r}
+clust3.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
+clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+clust17.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
+clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+clust10.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
+clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+clust11.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
+clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
+```
+
+## map cluster ids from Seuratv1
+```{r}
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
+seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
+seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
+seurat.object@meta.data$clust.ID <- as.factor(seurat.object@meta.data$clust.ID)
+```
+## plot something
+```{r}
+DimPlot(tech.integrated, group.by = "clust.ID", split.by = "integrated_snn_res.1.5", cols = c("black", "green", "blue", "red", "orange"), ncol = 4)
+```
+
+```{r}
+test.col.pattern <- "RNA_snn_res.1"
+target.col.pattern <- "integrated_snn_res"
+xlsx_filename <- paste0(projectName, "_pctVar", pct.var, "_mergedEnrichmentStats_Sv4B.xlsx")
+
+metadata_overlap_stats(seurat.object = seurat.object, 
+											 target.col.pattern = target.col.pattern, 
+											 test.col.pattern = test.col.pattern, 
+											 print_to_screen = TRUE,
+											 xlsx_filename = xlsx_filename)
+
+```
+
+## map cluster ids from Seuratv1
+```{r}
+tech.integrated@meta.data$clust.ID[rownames(tech.integrated@meta.data) %in% clust3.cells] <- 3
+tech.integrated@meta.data$clust.ID[rownames(tech.integrated@meta.data) %in% clust17.cells] <- 17
+tech.integrated@meta.data$clust.ID[rownames(tech.integrated@meta.data) %in% clust10.cells] <- 10
+tech.integrated@meta.data$clust.ID[rownames(tech.integrated@meta.data) %in% clust11.cells] <- 11
+tech.integrated@meta.data[is.na(tech.integrated@meta.data)] <- 0
+tech.integrated@meta.data$clust.ID <- as.factor(tech.integrated@meta.data$clust.ID)
+```
+## plot something
+```{r}
+DimPlot(tech.integrated, group.by = "clust.ID", split.by = "integrated_snn_res.1.5", cols = c("black", "green", "blue", "red", "orange"), ncol = 11)
+```
+
+```{r}
+DimPlot(seurat.object, group.by = "clust.ID", cols = c("black", "green", "blue", "red", "orange"))
+DimPlot(seurat.object, group.by = "clust.ID", split.by = "RNA_snn_res.1", cols = c("black", "green", "blue", "red", "orange"), ncol = 5)
+
+png(filename = paste0(projectName, "_pctVar", pct.var, "-Sv1ClustID_on_Sv4.png"), height = 800, width = 800)
+DimPlot(seurat.object, group.by = "clust.ID", cols = c("black", "green", "blue", "red", "orange"), pt.size = 1)
+dev.off()
+png(filename = paste0(projectName, "_pctVar", pct.var, "-Sv1ClustID_on_Sv4-facetSv4.png"), height = 800, width = 800)
+DimPlot(seurat.object, group.by = "clust.ID", split.by = "RNA_snn_res.1", cols = c("black", "green", "blue", "red", "orange"), ncol = 5) + NoAxes() + NoLegend()
+dev.off()
+```
+
+## generate emphasis colorpalette
+```{r}
+emphasis.palette <- rep("black", times = 25)
+```
+
+```{r}
+emphasis.palette.Sv4 <- emphasis.palette
+emphasis.palette.Sv4[c(4, 6, 7, 8)] <- c("green", "orange", "red", "blue")
+DimPlot(seurat.object, group.by = "RNA_snn_res.1", split.by = "integrated_snn_res.1.5", cols = emphasis.palette.Sv4, ncol = 5)
+
+png(filename = paste0(projectName, "_pctVar", pct.var, "-Sv4ClustID_on_Sv4B-facetSv4B.png"), height = 800, width = 800)
+DimPlot(seurat.object, group.by = "RNA_snn_res.1", split.by = "integrated_snn_res.1.5", cols = emphasis.palette.Sv4, ncol = 5) + NoAxes()
+dev.off()
+```

### 2021-11-17 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Additional plots added

@@
 ```
+## Thought experiment
+```{r}
+dbiomark.T <- data.table::transpose(dbiomark.data)
+dbiomark.object <- CreateSeuratObject(dbiomark.T, project = "biomark", assay = "RNA", meta.data = biomark.meta)
+```
 
@@
 tech.integrated <- RunPCA(tech.integrated, npcs = 30)
-DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident", pt.size = 2)
+DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident", pt.size = 2, cols = c("red", "black"))
+
+png(filename = paste0(projectName, "anchored_PCA.png"), height = 800, width = 800, bg = "transparent")
+DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident", pt.size = 2, cols = c("red", "black"), combine = FALSE)
+dev.off()
 ```

### 2021-12-08 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Updated analyses

--- 2021-11-12 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
@@
 	try(xlsx::write.xlsx(x = counts_df, 
-											 file = paste0(projectName, "_pctVar", pct.var, "_stats.xlsx"), 
+											 file = paste0(projectName, "_pctVar", integrated.var, "_stats.xlsx"), 
 											 sheetName = current_res, 
@@
 for(integrated.col in integrated.ident.list){
-		plot.title <- paste0(projectName, "_pctVar", pct.var, "-mergedEnrichment_OrigIdent-Facet-integratedRes.", gsub("integrated_snn_res.", "", integrated.col))
+		plot.title <- paste0(projectName, "_pctVar", integrated.var, "-mergedEnrichment_OrigIdent-Facet-integratedRes.", gsub("integrated_snn_res.", "", integrated.col))
 		myplot <- DimPlot(tech.integrated, 
@@
 
-		png(filename = paste0(projectName, "_pctVar", pct.var,
+		png(filename = paste0(projectName, "_pctVar", integrated.var,
 													"-mergedEnrichment_OrigIdent-Facet-integratedRes.", gsub("integrated_snn_res.", "", integrated.col),
@@
 ```{r}
-DimPlot(tech.integrated, group.by = "clust.ID", split.by = "integrated_snn_res.1.5", cols = c("black", "green", "blue", "red", "orange"), ncol = 4)
+DimPlot(seurat.object, group.by = "clust.ID", split.by = "integrated_snn_res.1.5", cols = c("black", "green", "blue", "red", "orange"), ncol = 4)
 ```
@@
 ```
+
+```{r}
+png(filename = paste0("CMP_RNA_res.1.png"), height = 800, width = 800)
+DimPlot(seurat.object, group.by = "RNA_snn_res.1", cols = color.palette, pt.size = 1)
+dev.off()
+
+```
+
+```{r}
+
+```

### 2021-12-10 monocle/msAggr_monocle.Rmd
Added msAggr document

---
title: "msAggr_monocle"
output: html_document
---

# Setup
## Load libraries
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/Monocle/")
library(monocle3)
library(ggplot2)
library(dplyr)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
```
# Load datasets
Provide list of datasets
```{r}
data.list <- c("/Users/heustonef/Desktop/10XGenomicsData/cellRanger/LSKm2",
							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/CMPm2", 
							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/MEPm", 
							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/GMPm")
```

### 2021-12-10 monocle/CMP_monocle.Rmd
Add CMP markdown and import cellranger data as cds

---
title: "CMP_monocle"
output: html_notebook
---
# Setup
## Define notebook variables
```{r}
working.genome <- "mm10"
umi.cutoff <- 200
```

## Load libraries
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/Monocle/")
library(monocle3)
library(ggplot2)
library(dplyr)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
```
# Load datasets
## Provide list of datasets
```{r}
data.path <- c("/Users/heustonef/Desktop/10XGenomicsData/cellRanger/CMPm2")
cds <- load_cellranger_data(data.path)
# cds <- new_cell_data_set(file.path(data.path, "filtered_feature_bc_matrix", "matrix.mtx.gz"),
# 										cell_metadata = cell_metadata, 
# 										gene_metadata = read.delim(file.path(data.path, "filtered_feature_bc_matrix"), header = FALSE, sep = "\t")[, 1:2]
# 										)
```
Extract metadata and gene_short_names from seurat.object
```{r, eval=FALSE}
# Don't run
library(Seurat)
seurat.object <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")
cell_metadata <- seurat.object@meta.data

```

### 2021-12-10 monocle/IndividualPops_monocle.Rmd
Changed CMP_monocle to IndividualPops_monocle because duh everyone should be analyzed the same

---
title: "CMP_monocle"
output: html_notebook
---
# Setup
## Define notebook variables
```{r}
working.genome <- "mm10"
umi.cutoff <- 200
data.path <- "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/"
cell.pop <- "CMP"
```

## Load libraries
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/Monocle/")
library(monocle3)
library(ggplot2)
library(dplyr)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
```
# Load datasets
## Provide list of datasets
After messing with this a bit, I've decided the easiest thing to do is integrate the data from the seurat object. This object already has the metadata and gene_short_names that I will need, and based on monocle3 documentation may be the easiest way to make sure everything gets where it needs to be.
```{r, eval=FALSE}
# Don't run
library(Seurat)
data.path <- 
seurat.object <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")

cell_metadata <- seurat.object@meta.data

```

### 2021-12-13 monocle/IndividualPops_monocle.Rmd
Completed initial trajectory inference for CMP via IndividualPops script

@@
 cell.pop <- "CMP"
+seurat.var.pct <- 90
+n.neighbors <- 20
+
+# Setting a seed to ensure reproducible data
+seed.value <- 10
+```
+## Notebook variable definitions
+```{r}
+seurat.rds <- paste0(cell.pop, "Subpop_pctVar", seurat.var.pct, ".RDS")
+projectName <- paste0(cell.pop, "Subpop_monocle" )
+set.seed(seed.value)
 ```
@@
 ```
+## Store session info
+```{r}
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
+```
+
 # Load datasets
@@
 This object already has the metadata and gene_short_names that I will need, and based on monocle3 documentation may be the easiest way to make sure everything gets where it needs to be.
-```{r, eval=FALSE}
-# Don't run
+```{r}
 library(Seurat)
-data.path <- 
-seurat.object <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")
+seurat.object <- readRDS(file.path(data.path, seurat.rds))
+seurat.xprsn <- GetAssayData(seurat.object, slot = "counts", assay = "RNA")
+cell_metadata <- seurat.object@meta.data[colnames(seurat.xprsn),]
+gene_metadata <- data.frame(gene_short_name = rownames(seurat.xprsn))
+rownames(gene_metadata) <- rownames(seurat.xprsn)
+cds <- new_cell_data_set(expression_data = seurat.xprsn, 
+												 cell_metadata = cell_metadata[colnames(seurat.xprsn), ], 
+												 gene_metadata = gene_metadata)
+rm(seurat.object, seurat.xprsn, cell_metadata, gene_metadata)
+```
+Look at object
+```{r}
+cds
+```
+## Save CDS object
+```{r}
+saveRDS(cds, file = paste0(projectName, ".RDS"))
+```
+# Data analysis
+## Load saved object
+```{r}
+cds <- readRDS(paste0(projectName, ".RDS"))
+```
 
-cell_metadata <- seurat.object@meta.data
+## Preprocessing
+```{r}
+cds <- preprocess_cds(cds = cds, metcdshod = "PCA", num_dim = 50, verbose = TRUE)
+```
+```{r}
+plot_pc_variance_explained(cds) + ggtitle("Proportion of variance explained (note not stdev)")
+```
+## calculate proportion of variance explained
+```{r}
+pct.list <- c(80, 85, 90, 95)
+for (pct in pct.list){
+	print(paste("Num pcs for", pct, "variance:", length(which(cumsum(cds@preprocess_aux$prop_var_expl*100) <= pct))))
+}
+```
+
+## Clustering analysis
+### Cluster with UMAP.
+UMAP is based on uwot's `umap` function
+```{r}
+cds <- reduce_dimension(cds = cds, max_components = 2, reduction_method = "UMAP", preprocess_method = "PCA", umap.n_neighbors = n.neighbors)
+```
+plot umap. Can color cells based on any `colData(cds)` column (essentially metadata columns)
+```{r}
+plot_cells(cds = cds, reduction_method = "UMAP", color_cells_by = "RNA_snn_res.1", show_trajectory_graph = FALSE, cell_size = .7, group_label_size = 5) + scale_color_manual(values = color.palette)
+```
+### Cluster wiht tSNE
+Note that this is stored in a different slot than UMAP and will not overwrite those data
+```{r}
+cds <- reduce_dimension(cds = cds, max_components = 2, reduction_method = "tSNE", preprocess_method = "PCA")
+```
+Plot tSNE. Must specifiy slot to plot
+```{r}
+plot_cells(cds = cds, reduction_method = "tSNE", color_cells_by = "RNA_snn_res.1", show_trajectory_graph = FALSE, cell_size = .7, group_label_size = 5) + scale_color_manual(values = color.palette)
+```
+## Save CDS object
+```{r}
+saveRDS(cds, file = paste0(projectName, ".RDS"))
+```
+### Check for batch effects
+Function `align_cds` tries to remove batch effects using mutual nearest neighbor alignment (package `batchelor`, Aaron Lun). Will test this on a different object.  
+Can specify alignment variable using parameter `alignment_group`. Note that only 1 variable can be specified
+```{r}
+# cds.aligned <- align_cds(cds = cds, preprocess_method = "PCA", alignment_group = "nCount_RNA")
+```
+*Failboat. Requires too much memory on laptop.*
+### Perform clustering analysis
+Monocle uses "community detection" (Leiden method and Louvain method).  
+Note that setting `resolution = NULL` and `cluster_method = "leiden"`, then resoltuion is determined automatically (actually I think it's just set to 0.0001, based on code)...
+```{r}
+# read RDS
+cds <- readRDS(file = paset0(projectName, ".RDS"))
+```
+
+```{r}
+res.list <- c(1e-4, 5e-4, 1e-3, 2e-3)
+# reduction.method <- "UMAP"
+# cluster.method <- "leiden"
+
+for(res in res.list){
+	# Cluster cells at given resolution
+	cds <- cluster_cells(cds = cds, 
+											 reduction_method = "UMAP", 
+											 k = n.neighbors, 
+											 cluster_method = "leiden", 
+											 resolution = res, 
+											 num_iter = 10, 
+											 random_seed = seed.value,
+											 verbose = TRUE)
+	# Copy cluster information to metadata in colData
+	coldata.res <- paste0("umap_res_", res)
+	colData(cds)[coldata.res] <- clusters(cds)
+}
+for(colID in colnames(colData(cds))){
+	if(grepl("umap_res", colID)){
+		cat("cluster count for", colID, ":", length(levels(colData(cds)[,colID])), "\n")
+	}
+}
 
 ```
+## save umap object
+```{r}
+saveRDS(cds, file = paste0(projectName, "-UMAP.RDS"))
+```
+## read object
+```{r}
+cds <- readRDS(paste0(projectName, "-UMAP.RDS"))
+```
+
+```{r}
+for(colID in colnames(colData(cds))){
+	if(grepl("umap_res", colID)){
+		if(length(levels(colData(cds)[,colID])) <= length(color.palette)){
+			cluster_plot <- plot_cells(cds = cds, reduction_method = "UMAP", 
+																 color_cells_by = colID, show_trajectory_graph = FALSE, 
+																 cell_size = 1, group_label_size = 5, label_groups_by_cluster = FALSE) + 
+				scale_color_manual(values = color.palette) +
+				ggtitle(paste0(projectName, "-", colID))
+			plot(cluster_plot)
+		} else {
+			cluster_plot <- plot_cells(cds = cds, reduction_method = "UMAP", 
+																 color_cells_by = colID,  show_trajectory_graph = FALSE, 
+																 cell_size = 1, group_label_size = 5, label_groups_by_cluster = FALSE) + 
+				ggtitle(paste0(projectName,  "-", colID))
+			plot(cluster_plot)
+
+		}
+		# save plot
+		png(filename = paste0(projectName,  "-", colID, ".png"), height= 800, width = 800)
+		plot(cluster_plot)
+		dev.off()
+	}
+}
+```
+
+### Explore resolution
+```{r eval = FALSE}
+# library(igraph)
+# library(schex)
+# library(scater)
+# library(scran)
+# library(ggrepel)
+# # plot_hexbin_meta_shiny(cds, col = "RNA_snn_res.1", action = "majority", min_nbins = 20, max_nbins = 50, dimension_reduction = "UMAP", na.rm = TRUE) 
+```
+For CMP, let's start exploring with res = 0.001
+# Find Marker Genes
+randomly guessing to set reference_cells value at 1000 (could also provide a list of cells if, e.g., wanted to specifiy an originating population)
+```{r}
+marker_test_res <- top_markers(cds, group_cells_by = "umap_res_0.001", reference_cells = 1000, cores = 8)
+```
+```{r}
+marker_test_res
+```
+KEY:
+a data.frame where the rows are genes and the columns are  
+* gene_id vector of gene names
+* gene_short_name vector of gene short names
+* cell_group character vector of the cell group to which the cell belongs
+* marker_score numeric vector of marker scores as the fraction expressing scaled by the specificity. The value ranges from 0 to 1.
+* mean_expression numeric vector of mean normalized expression of the gene in the cell group
+* fraction_expressing numeric vector of fraction of cells expressing the gene within the cell group
+* specificity numeric vector of a measure of how specific the gene's expression is to the cell group based on the Jensen-Shannon divergence. The value ranges from 0 to 1.
+* pseudo_R2 numeric vector of pseudo R-squared values, a measure of how well the gene expression model fits the categorical data relative to the null model. The value ranges from 0 to 1.
+* marker_test_p_value numeric vector of likelihood ratio p-values
+* marker_test_q_value numeric vector of likelihood ratio q-values  
+```{r}
+top_pseudoR2_markers <- marker_test_res %>%
+	filter(fraction_expressing >= 0.10) %>%
+	group_by(cell_group) %>%
+	top_n(3, pseudo_R2)
+top_pseudoR2_markers <- unique(top_pseudoR2_markers %>% pull(gene_id))
+my_plot <- plot_genes_by_group(cds, top_pseudoR2_markers, group_cells_by = "umap_res_0.001", ordering_type = "maximal_on_diag", max.size = 10)
+plot(my_plot)
+png(filename = paste0(projectName, "-map_res_0.001", "-pseudoR2markers.png"), height = 1200, width = 800)
+plot(my_plot)
+dev.off()
+```
+
+# Auto Annotate with Garnett
+## Start with marker_test_res
+```{r}
+# Require that markers have at least JS specificty score > 0.5 and
+# be significant in the logistic test for identifying their cell type:
+garnett_markers <- marker_test_res %>%
+                        filter(marker_test_q_value < 0.01 & specificity >= 0.5) %>%
+                        group_by(cell_group) %>%
+                        top_n(5, marker_score)
+# Exclude genes that are good markers for more than one cell type:
+garnett_markers <- garnett_markers %>% 
+                        group_by(gene_short_name) %>%
+                        filter(n() == 1)
+# Generate garnett marker file
+generate_garnett_marker_file(garnett_markers, file = paste0(projectName, "-garnettMarkers.txt"))
+```
+
+__Note that this file can be manually modified to add/remove classifier markers
+__Abandoning this!__ 1) I'm getting distracted and 2) Will have to play with resolution to get marker genes for more than 3 clusters
+
+# Trajectory Inference
+First make sure correct resolution is represented in cluster slot
+```{r}
+cds <- cluster_cells(cds = cds, 
+										 reduction_method = "UMAP", 
+										 k = n.neighbors, 
+										 cluster_method = "leiden", 
+										 resolution = 0.001, 
+										 num_iter = 10, 
+										 random_seed = seed.value,
+										 verbose = TRUE)
+```
+Fit principal graph. Set `use_partition = TRUE` to not require all cells be connected on a single graph
+```{r}
+cds <- learn_graph(cds, use_partition = TRUE, verbose = TRUE)
+```
+## save result
+```{r}
+saveRDS(cds, file = paste0(projectName, "-umap_res0.001_psdtime.RDS"))
+```
+
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+											color_cells_by = "cluster", 
+											label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE,  
+											cell_size = 1, group_label_size = 8, graph_label_size = 5) + 
+					scale_color_manual(values = color.palette) +
+					ggtitle(paste0(projectName, "-umap_res0.001_psdtime"))
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_res0.001_psdtime.png"), height = 800, width = 800)
+plot(my_plot)
+dev.off()
+```
+
+Will need to complete msAggr to order cells. For example, will pick those CMP cells that cluster with LSK and "earliest" CMPs.

### 2021-12-14 monocle/IndividualPops_monocle.Rmd
added gene expression and trajectory overlay plots to script

@@
 Will need to complete msAggr to order cells. For example, will pick those CMP cells that cluster with LSK and "earliest" CMPs.
+
+## Plot by gene expression
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+            genes = c("Elane", "Fli1", "Gata1", "Gata2", "Klf1", "Meis1", "Zfpm1", "Dntt", "Cd150"), trajectory_graph_color = "red",
+            label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE,  
+            cell_size = 1, group_label_size = 8, graph_label_size = 5, trajectory_graph_segment_size = 1)
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_YFG_psdtime.png"), height = 1600, width = 2000, bg = "transparent")
+plot(my_plot)
+dev.off()
+```

### 2021-12-16 monocle/msAggr_monocle.Rmd
msAggr analysis through orderCells

@@
 # Setup
+## Define notebook variables
+```{r}
+working.genome <- "mm10"
+umi.cutoff <- 200
+data.path <- "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/msAggr_seurat/"
+cell.pop <- "msAggr"
+seurat.var.pct <- 90
+n.neighbors <- 20
+
+# Setting a seed to ensure reproducible data
+seed.value <- 10
+```
+## Notebook variable definitions
+```{r}
+seurat.rds <- "msAggr_seurat_dim26.RDS"
+projectName <- paste0(cell.pop, "_monocle" )
+set.seed(seed.value)
+```
+
 ## Load libraries
@@
 source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
+## Have to get rid of the earliest "black" as a color because am plotting trajectories in black
+color.palette <- gsub("black", "plum", color.palette)
 ```
+## Store session info
+```{r}
+sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
+sink(sessionInfo.filename)
+sessionInfo()
+sink()
+```
+
 # Load datasets
-Provide list of datasets
+## Provide list of datasets
+After messing with this a bit, I've decided the easiest thing to do is integrate the data from the seurat object.
+This object already has the metadata and gene_short_names that I will need, and based on monocle3 documentation may be the easiest way to make sure everything gets where it needs to be.
 ```{r}
-data.list <- c("/Users/heustonef/Desktop/10XGenomicsData/cellRanger/LSKm2",
-							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/CMPm2", 
-							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/MEPm", 
-							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/GMPm")
+library(Seurat)
+seurat.object <- readRDS(file.path(data.path, seurat.rds))
+seurat.xprsn <- GetAssayData(seurat.object, slot = "counts", assay = "RNA")
+cell_metadata <- seurat.object@meta.data[colnames(seurat.xprsn),]
+gene_metadata <- data.frame(gene_short_name = rownames(seurat.xprsn))
+rownames(gene_metadata) <- rownames(seurat.xprsn)
+cds <- new_cell_data_set(expression_data = seurat.xprsn, 
+												 cell_metadata = cell_metadata[colnames(seurat.xprsn), ], 
+												 gene_metadata = gene_metadata)
+rm(seurat.object, seurat.xprsn, cell_metadata, gene_metadata)
 ```
+Look at object
+```{r}
+cds
+```
+## Save CDS object
+```{r}
+saveRDS(cds, file = paste0(projectName, ".RDS"))
+```
+# Data analysis
+## Load saved object
+```{r}
+cds <- readRDS(paste0(projectName, ".RDS"))
+```
+
+## Preprocessing
+```{r}
+cds <- preprocess_cds(cds = cds, metcdshod = "PCA", num_dim = 50, verbose = TRUE)
+```
+```{r}
+plot_pc_variance_explained(cds) + ggtitle("Proportion of variance explained (note not stdev)")
+```
+## calculate proportion of variance explained
+```{r}
+pct.list <- c(80, 85, 90, 95)
+for (pct in pct.list){
+	print(paste("Num pcs for", pct, "variance:", length(which(cumsum(cds@preprocess_aux$prop_var_expl*100) <= pct))))
+}
+```
+
+## Clustering analysis
+### Cluster with UMAP.
+UMAP is based on uwot's `umap` function
+```{r}
+cds <- reduce_dimension(cds = cds, max_components = 2, reduction_method = "UMAP", preprocess_method = "PCA", umap.n_neighbors = n.neighbors)
+```
+plot umap. Can color cells based on any `colData(cds)` column (essentially metadata columns)
+```{r}
+plot_cells(cds = cds, reduction_method = "UMAP", color_cells_by = "RNA_snn_res.1", show_trajectory_graph = FALSE, cell_size = .7, group_label_size = 5) + scale_color_manual(values = color.palette)
+```
+### Cluster wiht tSNE
+Note that this is stored in a different slot than UMAP and will not overwrite those data
+```{r}
+cds <- reduce_dimension(cds = cds, max_components = 2, reduction_method = "tSNE", preprocess_method = "PCA")
+```
+Plot tSNE. Must specifiy slot to plot
+```{r}
+plot_cells(cds = cds, reduction_method = "tSNE", color_cells_by = "RNA_snn_res.1", show_trajectory_graph = FALSE, cell_size = .7, group_label_size = 5) + scale_color_manual(values = color.palette)
+```
+## Save CDS object
+```{r}
+saveRDS(cds, file = paste0(projectName, ".RDS"))
+```
+### Check for batch effects
+Function `align_cds` tries to remove batch effects using mutual nearest neighbor alignment (package `batchelor`, Aaron Lun). Will test this on a different object.  
+Can specify alignment variable using parameter `alignment_group`. Note that only 1 variable can be specified
+```{r}
+# cds.aligned <- align_cds(cds = cds, preprocess_method = "PCA", alignment_group = "nCount_RNA")
+```
+*Failboat. Requires too much memory on laptop.*
+### Perform clustering analysis
+Monocle uses "community detection" (Leiden method and Louvain method).  
+Note that setting `resolution = NULL` and `cluster_method = "leiden"`, then resoltuion is determined automatically (actually I think it's just set to 0.0001, based on code)...
+```{r}
+# read RDS
+cds <- readRDS(file = paset0(projectName, ".RDS"))
+```
+
+```{r}
+res.list <- c(1e-4, 5e-4, 1e-3, 2e-3)
+# reduction.method <- "UMAP"
+# cluster.method <- "leiden"
+
+for(res in res.list){
+	# Cluster cells at given resolution
+	cds <- cluster_cells(cds = cds, 
+											 reduction_method = "UMAP", 
+											 k = n.neighbors, 
+											 cluster_method = "leiden", 
+											 resolution = res, 
+											 num_iter = 10, 
+											 random_seed = seed.value,
+											 verbose = TRUE)
+	# Copy cluster information to metadata in colData
+	coldata.res <- paste0("umap_res_", res)
+	colData(cds)[coldata.res] <- clusters(cds)
+}
+for(colID in colnames(colData(cds))){
+	if(grepl("umap_res", colID)){
+		cat("cluster count for", colID, ":", length(levels(colData(cds)[,colID])), "\n")
+	}
+}
+
+```
+## save umap object
+```{r}
+saveRDS(cds, file = paste0(projectName, "-UMAP.RDS"))
+```
+## read object
+```{r}
+cds <- readRDS(paste0(projectName, "-UMAP.RDS"))
+```
+
+```{r}
+plotting_cols <- c("umap_res", "orig.ident", "RNA_snn" )
+for(colID in colnames(colData(cds))){
+	if(grepl(paste(plotting_cols, collapse = "|"), colID)){
+		if(length(levels(colData(cds)[,colID])) <= length(color.palette)){
+			cluster_plot <- plot_cells(cds = cds, reduction_method = "UMAP", 
+																 color_cells_by = colID, show_trajectory_graph = FALSE, 
+																 cell_size = 1, group_label_size = 5, label_groups_by_cluster = FALSE) + 
+				scale_color_manual(values = color.palette) +
+				ggtitle(paste0(projectName, "-", colID))
+			plot(cluster_plot)
+		} else {
+			cluster_plot <- plot_cells(cds = cds, reduction_method = "UMAP", 
+																 color_cells_by = colID,  show_trajectory_graph = FALSE, 
+																 cell_size = 1, group_label_size = 5, label_groups_by_cluster = FALSE) + 
+				ggtitle(paste0(projectName,  "-", colID))
+			plot(cluster_plot)
+
+		}
+		# save plot
+		png(filename = paste0(projectName,  "-", colID, ".png"), height= 1600, width = 1600)
+		plot(cluster_plot)
+		dev.off()
+	}
+}
+```
+
+### Explore resolution
+```{r eval = FALSE}
+# library(igraph)
+# library(schex)
+# library(scater)
+# library(scran)
+# library(ggrepel)
+# # plot_hexbin_meta_shiny(cds, col = "RNA_snn_res.1", action = "majority", min_nbins = 20, max_nbins = 50, dimension_reduction = "UMAP", na.rm = TRUE) 
+```
+For CMP, let's start exploring with res = 0.001
+# Find Marker Genes
+randomly guessing to set reference_cells value at 1000 (could also provide a list of cells if, e.g., wanted to specifiy an originating population)
+```{r}
+marker_test_res <- top_markers(cds, group_cells_by = "umap_res_5e-04", reference_cells = 1000, cores = 7)
+```
+```{r}
+marker_test_res
+```
+KEY:
+a data.frame where the rows are genes and the columns are  
+* gene_id vector of gene names
+* gene_short_name vector of gene short names
+* cell_group character vector of the cell group to which the cell belongs
+* marker_score numeric vector of marker scores as the fraction expressing scaled by the specificity. The value ranges from 0 to 1.
+* mean_expression numeric vector of mean normalized expression of the gene in the cell group
+* fraction_expressing numeric vector of fraction of cells expressing the gene within the cell group
+* specificity numeric vector of a measure of how specific the gene's expression is to the cell group based on the Jensen-Shannon divergence. The value ranges from 0 to 1.
+* pseudo_R2 numeric vector of pseudo R-squared values, a measure of how well the gene expression model fits the categorical data relative to the null model. The value ranges from 0 to 1.
+* marker_test_p_value numeric vector of likelihood ratio p-values
+* marker_test_q_value numeric vector of likelihood ratio q-values  
+```{r}
+top_pseudoR2_markers <- marker_test_res %>%
+	filter(fraction_expressing >= 0.10) %>%
+	group_by(cell_group) %>%
+	top_n(3, pseudo_R2)
+top_pseudoR2_markers <- unique(top_pseudoR2_markers %>% pull(gene_id))
+my_plot <- plot_genes_by_group(cds, top_pseudoR2_markers, group_cells_by = "umap_res_0.001", ordering_type = "maximal_on_diag", max.size = 10)
+plot(my_plot)
+png(filename = paste0(projectName, "-map_res_0.001", "-pseudoR2markers.png"), height = 1200, width = 800)
+plot(my_plot)
+dev.off()
+```
+
+# Auto Annotate with Garnett
+## Start with marker_test_res
+```{r}
+# Require that markers have at least JS specificty score > 0.5 and
+# be significant in the logistic test for identifying their cell type:
+garnett_markers <- marker_test_res %>%
+                        filter(marker_test_q_value < 0.01 & specificity >= 0.5) %>%
+                        group_by(cell_group) %>%
+                        top_n(5, marker_score)
+# Exclude genes that are good markers for more than one cell type:
+garnett_markers <- garnett_markers %>% 
+                        group_by(gene_short_name) %>%
+                        filter(n() == 1)
+# Generate garnett marker file
+generate_garnett_marker_file(garnett_markers, file = paste0(projectName, "-garnettMarkers.txt"))
+```
+
+__Note that this file can be manually modified to add/remove classifier markers
+__Abandoning this!__ 1) I'm getting distracted and 2) Will have to play with resolution to get marker genes for more than 3 clusters
+
+# Trajectory Inference
+First make sure correct resolution is represented in cluster slot
+```{r}
+cds <- cluster_cells(cds = cds, 
+										 reduction_method = "UMAP", 
+										 k = n.neighbors, 
+										 cluster_method = "leiden", 
+										 resolution = 5e-4, 
+										 num_iter = 10, 
+										 random_seed = seed.value,
+										 verbose = TRUE)
+```
+Fit principal graph. Set `use_partition = TRUE` to not require all cells be connected on a single graph
+```{r}
+cds <- learn_graph(cds, use_partition = TRUE, verbose = TRUE)
+```
+## save result
+```{r}
+saveRDS(cds, file = paste0(projectName, "-umap_res5e-4_psdtime.RDS"))
+```
+
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+											color_cells_by = "cluster", 
+											label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE,  
+											cell_size = 1, group_label_size = 8, graph_label_size = 5) + 
+					scale_color_manual(values = color.palette) +
+					ggtitle(paste0(projectName, "-umap_res5e-4_psdtime"))
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_res5e-4_psdtime.png"), height = 1600, width = 1600, bg = "transparent")
+plot(my_plot)
+dev.off()
+```
+
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+											color_cells_by = "cluster", 
+											label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE, 
+											cell_size = 1, group_label_size = 8, graph_label_size = 5) + 
+					scale_color_manual(values = color.palette) +
+					ggtitle(paste0(projectName, "-umap_res5e-4_psdtime"))
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_res5e-4_psdtimeNoBranchPts.png"), height = 1600, width = 1600, bg = "transparent")
+plot(my_plot)
+dev.off()
+```
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+											color_cells_by = "orig.ident", 
+											label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE,  label_cell_groups = FALSE,
+											cell_size = 1, group_label_size = 8, graph_label_size = 5, trajectory_graph_color = "black", trajectory_graph_segment_size = 1.5) + 
+					scale_color_manual(values = c("red", "cadetblue", "orange", "green")) +
+					ggtitle(paste0(projectName, "-umap_res5e-4_orig.ident-psdtime"))
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_res5e-4_orig.ident-psdtime.png"), height = 1600, width = 1600, bg = "transparent")
+plot(my_plot)
+dev.off()
+
+```
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+											color_cells_by = "partition", label_cell_groups = FALSE,
+											label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE,  
+											cell_size = 1, group_label_size = 8, graph_label_size = 5) + 
+					scale_color_manual(values = color.palette) +
+					ggtitle(paste0(projectName, "-umap_res5e-4_psdtime_partition"))
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_res5e-4_psdtime-partition.png"), height = 1600, width = 1600, bg = "transparent")
+plot(my_plot)
+dev.off()
+```
+
+Will need to complete msAggr to order cells. For example, will pick those CMP cells that cluster with LSK and "earliest" CMPs.
+
+## Plot by gene expression
+```{r}
+my_plot <- plot_cells(cds, reduction_method = "UMAP", 
+											genes = c("Elane", "Fli1", "Gata1", "Gata2", "Klf1", "Meis1", "Zfpm1", "Dntt", "Cd150", "Flk2", "Cd48", "Kit", "Ly6a", "Ly6e"), trajectory_graph_color = "red",
+											label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE,  
+											cell_size = 1, group_label_size = 8, graph_label_size = 5, trajectory_graph_segment_size = 1)
+plot(my_plot)
+
+png(filename = paste0(projectName, "-umap_YFG_psdtime.png"), height = 3200, width = 4000, bg = "transparent")
+plot(my_plot)
+dev.off()
+```
+## Order cells
+Liking roots 11, 3, and 1, at the intersection of clusters 16/19/11. I like this because it's far away from the CMP and other branch points.
+```{r}
+cds <- order_cells(cds, verbose = TRUE)
+```
+```{r}
+plot_cells(cds,
+           color_cells_by = "pseudotime",
+           label_cell_groups=FALSE,
+           label_leaves=FALSE,
+           label_branch_points=FALSE,
+           graph_label_size=1.5)
+```

### 2022-01-13 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Minor additions for figures

@@
 ```{r}
-biomark.integrated <- subset(tech.integrated, subset = orig.ident == "biomark")
+biomark.integrated <- subset(tech.integrated, subset = orig.ident == "CMP")
 ```
@@
 ```
+```{r fig.height=12, fig.width=5}
+RidgePlot(biomark.integrated, features = c("Cd34pos", "Cd41pos", "Cd9pos", "Cd48pos"), cols = color.palette, ncol = 1, same.y.lims = TRUE)
+```
 ```{r}
 png(filename = paste0(projectName, "_pctVar", integrated.var, "res", integrated.res, "-fluor.png"), height = 2400, width = 800)
-RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
+RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1, same.y.lims = TRUE)
 dev.off()
 ```
+plot ridgeplot subsets
+```{r}
+featurelist = c("Cd34pos", "Cd41pos", "Cd9pos", "Cd48pos")
+for(fluor in featurelist){
+	print(fluor)
+	myplot = RidgePlot(biomark.integrated, features = fluor, cols = color.palette, idents = c(4, 5, 7,13), same.y.lims = TRUE, log = TRUE) + xlim(c(-3.5,6))
+	png(filename = paste0(projectName, "_pctVar", integrated.var, "res", integrated.res, "_", fluor, "-idents4.5.7.13.png"))
+	plot(myplot)
+	dev.off()
+}
+```
+
 plot umaps
@@
 ```{r}
+tech.object <- tech.integrated 
+```
+```{r}
+tech.object@meta.data["collapsed"] <- tech.object@meta.data$integrated_snn_res.1
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(12)] <- 11
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(13)] <- 11
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(14)] <- 11
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(15)] <- 11
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(16)] <- 11
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(17)] <- 11
+tech.object@meta.data$collapsed[tech.object@meta.data$collapsed == c(18)] <- 11
+tech.object@meta.data$collapsed <- factor(tech.object@meta.data$collapsed)
+```
+```{r}
+features <- c(
+	# "Kit", 
+	# "Cd34", 
+	"Cd48", 
+	"Dntt", 
+	"Irf8",
+	"Lgals3",
+	"Ly86",
+	# "Mpl",
+	"Mpo",
+	"Mecom",
+	"Pf4",
+	# "Itga2b",
+	# "Itgb3",
+	"Gp1bb", 
+	"Vwf",
+	"Cd9", 
+	"Gata2", 
+	"Gata1", 
+	"Klf1",
+	# "Mt2",
+	"Birc5", 
+	"Ube2c"
+	)
+```
+```{r}
+DoHeatmap(subset(tech.object, downsample = 100), features = features, group.by = "collapsed", group.colors = color.palette)
+```
 
+```{r}
+colnames(tech.integrated@meta.data)
 ```
+
+```{r}
+seurat.object<- readRDS("../IndividualPops/CMPSubpop_pctVar90.RDS")
+seurat.object$collapsed <- tech.object$collapsed
+```
+```{r}
+DimPlot(seurat.object, group.by = "RNA_snn_res.0.5", cols = color.palette)
+```
+```{r}
+# msaggr <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")
+# mscmp <- subset(msaggr, orig.ident == "CMPm2")
+# mscmp$collapsed <- tech.integrated$integrated_snn_res.0.5
+test.col.pattern <- "collapsed"
+target.col.pattern <- "RNA_snn_res."
+xlsx_filename <- paste0("msAggr",projectName, "_mergedEnrichmentStats.xlsx")
+
+metadata_overlap_stats(seurat.object = mscmp, 
+											 target.col.pattern = target.col.pattern, 
+											 test.col.pattern = test.col.pattern, 
+											 print_to_screen = TRUE,
+											 xlsx_filename = xlsx_filename)
+```
+
+```{r}
+msaggr <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")
+```
+```{r}
+source("ScanpyPalette.R")
+```
+
+```{r}
+msaggr$collapsed <- tech.integrated$integrated_snn_res.0.5
+emphasis.palette <- rep("#7F7E7F", 16)
+emphasis.palette[5] <- "red"
+emphasis.palette[6] <- "blue"
+emphasis.palette[8] <- "orange"
+emphasis.palette[14] <- "green"
+png(filename = "Integrated_clusters.png", height = 800, width = 800)
+plot(DimPlot(msaggr, group.by = "collapsed", cols = emphasis.palette, na.value = "#7F7E7F", pt.size = 1))
+dev.off()
+```

### 2022-01-13 msAggr_seurat/msAggr_seurat.Rmd
Minor additions for figures

@@
 jackstraw.dim <- 40
+
 ```
@@
 ```{r}
-object.res <- ".0.5"
+object.res <- ".1"
 Idents(seurat.object) <- paste0("RNA_snn_res", object.res)
@@
 
-## DGE w/ resolution = 0.5
+## DGE 
 Strt with comparing all clusters against all other clusters and write out cluster info
@@
 # ident.list <- colnames(seurat.object@meta.data)[grepl("^RNA_snn", colnames(seurat.object@meta.data))]
+ident.list <- c("RNA_snn_res.1")
+# ident.list <- c("RNA_snn_res.0.5")
+cut_off <- 1.2
 # ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1")
-# ident.list <- c("RNA_snn_res.1")
-ident.list <- c("RNA_snn_res.0.5")
-cut_off <- 1.2
+
 load(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/mouseIDTable.rda")
 
-for(tested.ident in ident.list){
-	ident.res <- gsub("RNA_snn_res", "", tested.ident)
-	seurat.object <- SetIdent(seurat.object, value = tested.ident)
-	all.markers <- FindAllMarkers(seurat.object)
-	# xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
-	# 								 file = paste0(projectName, "_FindALLMarkers_dim",ndims, ".xlsx"), 
-	# 								 sheetName = paste0("res", ident.res), 
-	# 								 col.names = TRUE, 
-	# 								 row.names = FALSE, 
-	# 								 append = TRUE)
+# for(tested.ident in ident.list){
+# 	ident.res <- gsub("RNA_snn_res", "", tested.ident)
+# 	seurat.object <- SetIdent(seurat.object, value = tested.ident)
+# 	all.markers <- FindAllMarkers(seurat.object)
+all.markers <- data.frame(all.markers %>% group_by(cluster))
+top100 <- data.frame(all.markers %>% group_by(cluster) %>% top_n(n = 100, wt = absValue))
+	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")],
+									 file = paste0(projectName, "_FindALLMarkers_dim",ndims, ".xlsx"),
+									 sheetName = paste0("res", ident.res),
+									 col.names = TRUE,
+									 row.names = FALSE,
+									 append = TRUE)
+		xlsx::write.xlsx(x = data.frame(top100[,c("avg_log2FC", "p_val_adj", "cluster", "gene")]),
+									 file = paste0(projectName, "_FindALLMarkers_dim",ndims, "-top100.xlsx"),
+									 sheetName = 'top100',
+									 col.names = TRUE,
+									 row.names = FALSE,
+									 append = FALSE)
 
@@
 		write.table(new.table, file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
-		write.table(new.table[new.table$avg_log2FC >= cut_off,], file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, "-co", cut_off, "up.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
-		write.table(new.table[new.table$avg_log2FC >= (cut_off)*-1,], file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, "-co", cut_off, "dn.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		# write.table(new.table[new.table$avg_log2FC >= cut_off,], file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, "-co", cut_off, "up.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		# write.table(new.table[new.table$avg_log2FC >= (cut_off)*-1,], file = paste0("RankList_res", ident.res, "_findAll_hgnc/res", ident.res, "_cluster", cluster, "-co", cut_off, "dn.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
 		}
+
+		#top 100		
+	HGNC <- with(mouseIDTable, mouseIDTable$HGNC[match(top100$gene, mouseIDTable$MGI)])
+	top100$HGNC <- HGNC
+	sig.res <- top100[top100$p_val_adj <= 0.05, ]
+	sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
+	sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
+	sig.res <- sig.res[]
+
+	for(cluster in unique(sig.res$cluster)){
+		print(paste("writing cluster", cluster))
+		new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
+		new.table <- new.table[order(-new.table$avg_log2FC), ]
+		try(dir.create(paste0("RankList_res", ident.res, "_findAll_hgnc_top100/"), showWarnings = FALSE))
+		write.table(new.table, file = paste0("RankList_res", ident.res, "_findAll_hgnc_top100/res", ident.res, "_cluster", cluster, "-top100.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		# write.table(new.table[new.table$avg_log2FC >= cut_off,], file = paste0("RankList_res", ident.res, "_findAll_hgnc_top100/res", ident.res, "_cluster", cluster, "-co", cut_off, "top100up.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
+		# write.table(new.table[new.table$avg_log2FC >= (cut_off)*-1,], file = paste0("RankList_res", ident.res, "_findAll_hgnc_top100/res", ident.res, "_cluster", cluster, "-co", cut_off, "top100dn.rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )
 }
@@
 source("/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/GoEnrichmentAnalysis.R")
-GO_enrichment_function(my_folder = "RankList_res.0.5_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
-# GO_enrichment_function(my_folder = "RankList_res.1_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
+# GO_enrichment_function(my_folder = "RankList_res.0.5_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
+GO_enrichment_function(my_folder = "RankList_res.1_findAll_hgnc/", file.pattern = ".rnk", id.type = "symbol")
+GO_enrichment_function(my_folder = "backups/RankList_res.1_findAll_hgnc/", file.pattern = "up.rnk", id.type = "symbol")
+```
+
+write table of cluster IDs
+```{r}
+write.table(seurat.object@meta.data, file = "msAggr_SeuratMetaData.txt", row.names = row.names(seurat.object@meta.data), sep = "\t", col.names = TRUE, quote = FALSE)
 ```

### 2022-01-18 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Minor change to document while in wrong brach

@@
 ```
+
 ## Thought experiment
